### 2019-11-20 appendix/appendix-institutional-design-climate-policy.Rmd
appendix v0.1

---
title: |
       | Supplementary Materials for "Institutional Design and Elite Support for Climate Policies: Evidence from Latin American Countries"
author: 
- "Danilo Freire[^freire]"
- "Umberto Mignozzetti[^mignozzetti]"
- "David Skarbek[^skarbek]"
date: November 20, 2019
fontfamily: libertine
fontawesome: yes
fontsize: 12pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    toc: true
    keep_tex: no
    template: template.latex
---

[^freire]: Postdoctoral Research Associate, The Political Theory Project, Brown University, Providence, RI 02912, USA, <danilofreire@brown.edu>, <http://danilofreire.github.io>. Corresponding author.

[^mignozzetti]: School of International Relations, Fundação Getulio Vargas, São Paulo, SP, Brazil and Wilf Family Department of Politics, NYU, NY, USA, [umberto.mig@nyu.edu](mailto:umberto.mig@nyu.edu), <http://umbertomig.com>.

[^skarbek]: The Department of Political Science and the Political Theory Project, Brown University, Providence, RI, USA, <davidskarbek@brown.edu>, <http://davidskarbek.com>.

```{r setup, echo=FALSE}
plotAMCE <- 
    function (x, main = "", xlab = "Change in E[Y]", 
              ci = 0.95, colors = NULL, 
              xlim = NULL, breaks = NULL, 
              labels = NULL, attribute.names = NULL, 
              level.names = NULL, label.baseline = TRUE, 
              text.size = 11, 
              text.color = "black", 
              point.size = 0.5, dodge.size = 0.9, 
              plot.theme = NULL, plot.display = "all", 
              facet.names = NULL, 
              facet.levels = NULL, 
              group.order = NULL, font.family = NULL, 
              tblfy = F, ...) {
        amce_obj <- x
        ylim <- xlim
        pe <- NULL
        se <- NULL
        group <- NULL
        lower <- NULL
        upper <- NULL
        var <- NULL
        printvar <- NULL
        facet <- NULL
        raw_attributes <- names(amce_obj$estimates)
        raw_levels <- lapply(amce_obj$estimates, colnames)
        for (effect in names(raw_levels)) {
            effect_elements <- strsplit(effect, ":")[[1]]
            baseline_interactions <- c()
            for (elem in effect_elements) {
                base_coef <- paste(c(elem, amce_obj$baselines[[elem]]), 
                                   collapse = "")
                baseline_interactions <- c(baseline_interactions, 
                                           base_coef)
            }
            interaction_str <- paste(baseline_interactions, sep = "", 
                                     collapse = ":")
            raw_levels[[effect]] <- c(interaction_str, raw_levels[[effect]])
        }
        if (ci < 1 & ci > 0) {
            zscr <- qnorm(1 - ((1 - ci)/2))
        }
        else {
            cat("Invalid confidence interval -- Defaulting to 95%")
            zscr <- qnorm(1 - ((1 - 0.95)/2))
        }
        if (!is.null(attribute.names)) {
            attribute.names <- unique(attribute.names)
            if (length(attribute.names) != length(raw_attributes)) {
                cat(paste("Error: The number of unique elements in attribute.names ", 
                          length(attribute.names), " does not match the attributes in amce object for which estimates were obtained: ", 
                          paste(raw_attributes, collapse = ", "), "\n", 
                          sep = ""))
                cat("Defaulting attribute.names to attribute names in AMCE object\n")
                attribute.names <- NULL
            }
        }
        if (!is.null(level.names)) {
            names(level.names) <- clean.names(names(level.names))
            for (name in names(level.names)) {
                if (name %in% names(raw_levels)) {
                    if (length(level.names[[name]]) != length(raw_levels[[name]])) {
                        cat(paste("Error: level.names lengths do not match levels for attribute ", 
                                  name, "\n", sep = ""))
                        cat(paste("Defaulting level.names for attribute ", 
                                  name, " to level names in AMCE object", "\n", 
                                  sep = ""))
                        level.names[[name]] <- NULL
                    }
                }
                else {
                    cat(paste("Error: level.names entry ", name, 
                              " not in AMCE object. Removing level.names for attribute.", 
                              "\n", sep = ""))
                    level.names[[name]] <- NULL
                }
            }
        }
        if (is.null(attribute.names)) {
            attribute.names <- c()
            for (attr in names(amce_obj$estimates)) {
                attr_split <- strsplit(attr, ":")[[1]]
                attr_lookup <- paste(unlist(sapply(attr_split, function(x) amce_obj$user.names[x])), 
                                     collapse = ":")
                attribute.names <- c(attribute.names, attr_lookup)
            }
        }
        if (is.null(level.names)) 
            level.names <- list()
        if (any(!names(raw_levels) %in% names(level.names))) {
            for (attr in names(raw_levels)[!names(raw_levels) %in% 
                                           names(level.names)]) {
                attr_split <- strsplit(raw_levels[[attr]], ":")
                level.names[[attr]] <- unlist(lapply(attr_split, 
                                                     function(x) paste(sapply(x, function(y) amce_obj$user.levels[y]), 
                                                                       collapse = ":")))
            }
        }
        plot.display.opts <- c("all", "unconditional", "interaction")
        if (!is.element(plot.display, plot.display.opts)) {
            stop(paste(c("Error-- plot.display must be once of: ", 
                         paste(plot.display.opts, collapse = ", ")), collapse = " "))
        }
        if (!is.null(facet.names)) {
            facet.names <- clean.names(facet.names)
        }
        else if (!is.null(facet.levels)) {
            facet.names <- clean.names(names(facet.levels))
        }
        if (!is.null(facet.names)) {
            facet.names.check <- c()
            for (facet.name in facet.names) {
                if (grepl(":", facet.name)) 
                    stop("Error-- cannot facet by interaction in current version.")
                if (!facet.name %in% names(amce_obj$estimates) & 
                    !facet.name %in% names(amce_obj$cond.estimates)) {
                    stop(paste(c("Error-- cannot find facet name", 
                                 facet.name, "in AMCE object output."), collapse = " "))
                }
                else {
                    facet.names.check <- c(facet.names.check, facet.name)
                }
            }
            facet.names <- facet.names.check
        }
        if ((is.null(facet.names)) & (length(amce_obj$respondent.varying) > 
                                      0) & (plot.display != "unconditional")) {
            facet.names <- amce_obj$respondent.varying
        }
        if (is.null(facet.names) & plot.display == "interaction") {
            warning("Warning: no facet name or respondent varying characteristic provided to calculate conditional estimates. Will display unconditional only")
            plot.display <- "unconditional"
        }
        if (plot.display == "unconditional" & !is.null(facet.names)) {
            warning("Warning-- plot display is set to unconditional, facet names will be ignored")
            facet.names <- NULL
            facet.levels <- NULL
        }
        if (!is.null(facet.levels)) {
            names(facet.levels) <- clean.names(names(facet.levels))
            for (facet.name in names(facet.levels)) {
                if (facet.name %in% names(amce_obj$baselines)) {
                    facet.levels[[facet.name]] <- clean.names(facet.levels[[facet.name]])
                    if (facet.name %in% names(amce_obj$estimates) && 
                        is.element(amce_obj$baselines[[facet.name]], 
                                   facet.levels[[facet.name]])) {
                        stop(paste(c("Error: Facet level \"", as.character(amce_obj$baselines[[facet.name]]), 
                                     "\" is the baseline level of a profile varying attribute. Please provide alternative facet level or use defaults."), 
                                   collapse = ""))
                    }
                    if (is.null(names(facet.levels[[facet.name]]))) {
                        fac.levs <- sapply(facet.levels[[facet.name]], 
                                           function(x) paste(facet.name, x, sep = ""))
                        names(facet.levels[[facet.name]]) <- sapply(fac.levs, 
                                                                    USE.NAMES = F, function(x) amce_obj$user.levels[[x]])
                    }
                }
                else if (is.null(names(facet.levels[[facet.name]]))) {
                    names(facet.levels[[facet.name]]) <- as.character(facet.levels[[facet.name]])
                }
            }
        }
        if (is.null(facet.levels)) 
            facet.levels <- list()
        if (any(!facet.names %in% names(facet.levels))) {
            for (facet.name in facet.names[!facet.names %in% names(facet.levels)]) {
                if (facet.name %in% names(amce_obj$baselines)) {
                    if (facet.name %in% names(amce_obj$estimates)) {
                        fac.levs <- colnames(amce_obj$estimates[[facet.name]])
                    }
                    else {
                        fac.levs <- colnames(amce_obj$cond.estimates[[facet.name]])
                    }
                    facet.levels[[facet.name]] <- sub(facet.name, 
                                                      "", fac.levs)
                    facet.levels[[facet.name]] <- c(amce_obj$baselines[[facet.name]], 
                                                    facet.levels[[facet.name]])
                    fac.levs <- c(paste(facet.name, amce_obj$baselines[[facet.name]], 
                                        sep = ""), fac.levs)
                    names(facet.levels[[facet.name]]) <- sapply(fac.levs, 
                                                                USE.NAMES = F, function(x) amce_obj$user.levels[[x]])
                }
                else if (facet.name %in% names(amce_obj$continuous)) {
                    facet.levels[[facet.name]] <- amce_obj$continuous[[facet.name]]
                }
            }
        }
        covariate.values <- list()
        for (var in names(facet.levels)) {
            if (var %in% amce_obj$respondent.varying) {
                covariate.values[[var]] <- facet.levels[[var]]
            }
        }
        d <- data.frame(pe = c(), se = c(), upper = c(), lower = c(), 
                        var = c(), printvar = c(), group = c(), facet = c())
        if (plot.display != "interaction") {
            if (plot.display == "all") {
                uncond.facet.name <- "Unconditional"
            }
            else {
                uncond.facet.name <- NA
            }
            if (plot.display == "all" && !is.null(facet.names)) {
                attr_remove <- c()
                for (facet.name in facet.names[!is.element(facet.names, 
                                                           amce_obj$respondent.varying)]) {
                    attr_remove1 <- raw_attributes[grepl(":", raw_attributes)]
                    attr_remove1 <- attr_remove1[grepl(facet.name, 
                                                       attr_remove1)]
                    attr_remove <- c(attr_remove, attr_remove1)
                }
                raw_attributes <- raw_attributes[!is.element(raw_attributes, 
                                                             attr_remove)]
            }
            for (i in 1:length(raw_attributes)) {
                attr_name <- raw_attributes[i]
                print_attr_name <- attribute.names[which(names(amce_obj$estimates) == 
                                                             raw_attributes[i])]
                d_head <- data.frame(pe = NA, se = NA, upper = NA, 
                                     lower = NA, var = attr_name, printvar = paste(print_attr_name, 
                                                                                   ":", sep = ""), group = "<NA>", facet = uncond.facet.name)
                d <- rbind(d, d_head)
                for (j in 1:length(raw_levels[[attr_name]])) {
                    level_name <- raw_levels[[attr_name]][j]
                    print_level_name <- level.names[[attr_name]][j]
                    if (j == 1) {
                        if (label.baseline) {
                            print_level_name <- paste("(Baseline = ", 
                                                      print_level_name, ")", sep = "")
                        }
                        d_lev <- data.frame(pe = NA, se = NA, upper = NA, 
                                            lower = NA, var = level_name, printvar = paste("   ", 
                                                                                           print_level_name, sep = ""), group = print_attr_name, 
                                            facet = uncond.facet.name)
                    }
                    else {
                        val_pe <- amce_obj$estimates[[attr_name]][1, 
                                                                  level_name]
                        val_se <- amce_obj$estimates[[attr_name]][2, 
                                                                  level_name]
                        upper_bnd <- val_pe + zscr * val_se
                        lower_bnd <- val_pe - zscr * val_se
                        d_lev <- data.frame(pe = val_pe, se = val_se, 
                                            upper = upper_bnd, lower = lower_bnd, var = level_name, 
                                            printvar = paste("   ", print_level_name, 
                                                             sep = ""), group = print_attr_name, facet = uncond.facet.name)
                    }
                    d <- rbind(d, d_lev)
                }
            }
        }
        if (plot.display != "unconditional" & !is.null(facet.names)) {
            for (facet.name in facet.names) {
                print_facet_name <- amce_obj$user.names[[facet.name]]
                all_req_vars <- attr(terms(amce_obj$formula), "term.labels")
                all_mod <- unlist(sapply(all_req_vars, function(x) {
                    y <- strsplit(x, ":")[[1]]
                    if (any(y == facet.name)) 
                        x
                }))
                all_mod <- unlist(sapply(all_mod, function(x) {
                    subs <- strsplit(x, ":")[[1]]
                    subs <- subs[is.element(subs, names(amce_obj$estimates))]
                    subs <- subs[subs != facet.name]
                    if (length(subs) > 0) 
                        paste(subs, collapse = ":")
                }))
                if (length(all_mod) == 0) {
                    stop(paste(c("Error: Facet variable", facet.name, 
                                 "not interacted with profile attributes"), 
                               collapse = " "))
                }
                all_mod <- unique(all_mod)
                if (is.element(facet.name, names(amce_obj$estimates))) {
                    facet.start <- 2
                }
                else {
                    facet.start <- 1
                }
                for (k in facet.start:length(facet.levels[[facet.name]])) {
                    facet_lev <- facet.levels[[facet.name]][k]
                    if (is.element(facet.name, names(amce_obj$estimates))) {
                        print_facet_level <- paste(c("ACIE", paste(c(print_facet_name, 
                                                                     names(facet.levels[[facet.name]])[k]), collapse = " = ")), 
                                                   collapse = "\n")
                    }
                    else {
                        print_facet_level <- paste(c("Conditional on", 
                                                     paste(c(print_facet_name, names(facet.levels[[facet.name]])[k]), 
                                                           collapse = " = ")), collapse = "\n")
                    }
                    for (mod_var in all_mod) {
                        print_attr_name <- attribute.names[which(names(amce_obj$estimates) == 
                                                                     mod_var)]
                        d_head <- data.frame(pe = NA, se = NA, upper = NA, 
                                             lower = NA, var = mod_var, printvar = paste(print_attr_name, 
                                                                                         ":", sep = ""), group = "<NA>", facet = print_facet_level)
                        d <- rbind(d, d_head)
                        if (facet.name %in% names(amce_obj$estimates)) {
                            inter_coef <- paste(sort(c(mod_var, facet.name)), 
                                                collapse = ":")
                            estimate.source <- amce_obj$estimates[[inter_coef]]
                            estimate.source <- estimate.source[, grep(paste0(facet.name, 
                                                                             facet_lev), colnames(estimate.source))]
                        }
                        else {
                            estimate.source <- get.conditional.effects(amce_obj, 
                                                                       covariate.values, facet.name, facet_lev, 
                                                                       mod_var)
                        }
                        mod_vars <- strsplit(mod_var, ":")[[1]]
                        for (p in 1:length(raw_levels[[mod_var]])) {
                            mod_coef <- raw_levels[[mod_var]][p]
                            mod_coefs <- strsplit(mod_coef, ":")[[1]]
                            for (lev in 1:length(mod_coefs)) {
                                mod_lev <- sub(mod_vars[lev], "", mod_coefs[lev])
                            }
                            print_level_name <- level.names[[mod_var]][p]
                            if (p == 1) {
                                if (label.baseline) {
                                    print_level_name <- paste("(Baseline = ", 
                                                              print_level_name, ")", sep = "")
                                }
                                d_lev <- data.frame(pe = NA, se = NA, upper = NA, 
                                                    lower = NA, var = mod_coef, printvar = paste("   ", 
                                                                                                 print_level_name, sep = ""), group = print_attr_name, 
                                                    facet = print_facet_level)
                            }
                            else {
                                val_pe <- estimate.source[1, p - 1]
                                if (!is.na(val_pe)) {
                                    val_se <- estimate.source[2, p - 1]
                                    upper_bnd <- val_pe + zscr * val_se
                                    lower_bnd <- val_pe - zscr * val_se
                                }
                                else {
                                    val_se <- upper_bnd <- lower_bnd <- NA
                                }
                                d_lev <- data.frame(pe = val_pe, se = val_se, 
                                                    upper = upper_bnd, lower = lower_bnd, 
                                                    var = mod_coef, printvar = paste("   ", 
                                                                                     print_level_name, sep = ""), group = print_attr_name, 
                                                    facet = print_facet_level)
                            }
                            d <- rbind(d, d_lev)
                        }
                    }
                }
            }
        }
        else {
            d <- d[, -which(colnames(d) == "facet")]
        }
        if (is.null(ylim)) {
            max_upper <- max(d$upper, na.rm = T) + 0.05
            min_lower <- min(d$lower, na.rm = T) - 0.05
            ylim <- c(min_lower, max_upper)
            d[is.na(d)] <- max_upper + 100
        }
        else {
            d[is.na(d)] <- max(ylim) + 100
        }
        d$group[d$group == "<NA>"] <- NA
        if (!is.null(facet.names)) 
            d$facet[d$facet == "<NA>"] <- NA
        d$var <- factor(d$var, levels = unique(d$var)[length(d$var):1])
        if (!is.null(facet.names)) {
            d$facet <- factor(d$facet, levels = unique(d$facet))
        }
        if (!is.null(group.order)) {
            n.row <- length(unique(as.character(d$var)))
            order.var <- vector("character", length = n.row)
            i <- 1
            while (i < n.row) {
                for (j in group.order) {
                    order.var[i] <- unique(as.character(d$var[d$var == 
                                                                  gsub(" ", "", j)]))
                    i <- i + 1
                    temp.d <- d
                    temp.d$group <- gsub(" ", "", temp.d$group)
                    temp.d <- subset(temp.d, group == gsub(" ", "", 
                                                           j))
                    temp.var <- unique((as.character(temp.d$var)))
                    order.var[i:(i + length(temp.var) - 1)] <- temp.var
                    i <- i + length(temp.var)
                }
            }
            order.var <- rev(order.var)
            order.df <- data.frame(order.var, 1:length(order.var))
            colnames(order.df) <- c("var", "order")
            d$var <- factor(d$var, levels = order.var)
            d <- merge(d, order.df, by.x = "var", by.y = "var", suffixes = c("", 
                                                                             ""))
        }
        levs <- as.character(d$printvar)
        levs <- gsub('organizations', 'organisations', levs)
        d$printvar <- factor(levs, levels = levs)
        if(tblfy) {
            return(d)
        }
        p = ggplot(d, aes(y = pe, x = var, colour = group))
        p = p + coord_flip(ylim = ylim)
        p = p + geom_hline(yintercept = 0, size = 0.5, colour = "black", 
                           linetype = "dotted")
        p = p + geom_pointrange(aes(ymin = lower, ymax = upper), 
                                position = position_dodge(width = dodge.size), size = point.size)
        if (!is.null(facet.names)) {
            p = p + facet_wrap(~facet)
        }
        if (is.null(breaks) & is.null(labels)) {
            p = p + scale_y_continuous(name = xlab)
        }
        else if (is.null(breaks) & !is.null(labels)) {
            p = p + scale_y_continuous(name = xlab, labels = labels)
        }
        else if (!is.null(breaks) & is.null(labels)) {
            p = p + scale_y_continuous(name = xlab, breaks = breaks)
        }
        else if (!is.null(breaks) & !is.null(labels)) {
            p = p + scale_y_continuous(name = xlab, breaks = breaks, 
                                       labels = labels)
        }
        if (!is.null(group.order)) {
            fix.xlabs.df <- d[!duplicated(d$var), ]
            fix.xlabs <- fix.xlabs.df[order(-fix.xlabs.df$order), 
                                      ]$printvar
        }
        else {
            fix.xlabs <- as.character(d$printvar)[!duplicated(d$var)]
        }
        p = p + scale_x_discrete(name = "", labels = fix.xlabs[length(fix.xlabs):1])
        if (!is.null(main)) {
            if (main != "") {
                p = p + ggtitle(main)
            }
        }
        if (is.null(colors)) {
            p = p + scale_colour_discrete(" ")
        }
        else if (is.vector(colors)) {
            cPal <- rep(colors, ceiling(length(unique(d$group))/length(colors)))
            p = p + scale_colour_manual(values = cPal)
        }
        else {
            cat("Error: 'colors' must be a vector. Using default colors\n")
            p = p + scale_colour_discrete(" ")
        }
        if (is.null(plot.theme)) {
            theme_bw1 <- function(base_size = text.size, base_family = "") {
                theme_grey(base_size = base_size, base_family = base_family) %+replace% 
                    theme(axis.text.x = element_text(size = base_size * 
                                                         0.9, colour = text.color, hjust = 0.5, vjust = 1), 
                          axis.text.y = element_text(size = base_size, 
                                                     colour = text.color, hjust = 0, vjust = 0.5, 
                                                     family = font.family), axis.ticks = element_line(colour = "grey50"), 
                          axis.title.y = element_text(size = base_size, 
                                                      angle = 90, vjust = 0.01, hjust = 0.1, family = font.family), 
                          plot.title = element_text(face = "bold", family = font.family), 
                          legend.position = "none")
            }
            p = p + theme_bw1()
            print(p)
        }
        else if (is.null(class(plot.theme))) {
            cat("Error: 'plot.theme' is not a valid ggplot theme object. Using default theme\n")
            theme_bw1 <- function(base_size = text.size, base_family = "") {
                theme_grey(base_size = base_size, base_family = base_family) %+replace% 
                    theme(axis.text.x = element_text(size = base_size * 
                                                         0.9, colour = text.color, hjust = 0.5, vjust = 1), 
                          axis.text.y = element_text(size = base_size, 
                                                     colour = text.color, hjust = 0, vjust = 0.5, 
                                                     family = font.family), axis.ticks = element_line(colour = "grey50"), 
                          axis.title.y = element_text(size = base_size, 
                                                      angle = 90, vjust = 0.01, hjust = 0.1, family = font.family), 
                          plot.title = element_text(face = "bold", family = font.family), 
                          legend.position = "none")
            }
            p = p + theme_bw1()
            print(p)
        }
        else if (class(plot.theme)[1] != "theme") {
            cat("Error: 'plot.theme' is not a valid ggplot theme object. Using default theme\n")
            theme_bw1 <- function(base_size = text.size, base_family = "") {
                theme_grey(base_size = base_size, base_family = base_family) %+replace% 
                    theme(axis.text.x = element_text(size = base_size * 
                                                         0.9, colour = text.color, hjust = 0.5, vjust = 1), 
                          axis.text.y = element_text(size = base_size, 
                                                     colour = text.color, hjust = 0, vjust = 0.5, 
                                                     family = font.family), axis.ticks = element_line(colour = "grey50"), 
                          axis.title.y = element_text(size = base_size, 
                                                      angle = 90, vjust = 0.01, hjust = 0.1, family = font.family), 
                          plot.title = element_text(face = "bold", family = font.family), 
                          legend.position = "none")
            }
            p = p + theme_bw1()
            print(p)
        }
        else {
            p = p + plot.theme
            print(p)
        }
        if (length(covariate.values) > 1) {
            resp.message <- c("Note:")
            for (this.var in names(covariate.values)) {
                resp.message <- paste(c(resp.message, " For AMCE and ACIE conditional on ", 
                                        this.var, ", "), collapse = "")
                other.vars <- names(covariate.values)[names(covariate.values) != 
                                                          this.var]
                other.levels <- c()
                for (var in other.vars) {
                    other.levels <- c(other.levels, paste(c(var, 
                                                            " will be held at level \"", names(covariate.values[[var]])[1], 
                                                            "\""), collapse = ""))
                }
                other.levels <- paste(other.levels, collapse = ", and ")
                resp.message <- c(resp.message, other.levels, ".")
                resp.message <- paste(resp.message, collapse = "")
            }
            cat(resp.message, "\n")
        }
    }

```

# Latin American Elite Survey \label{sec:latam}

\setlength{\parindent}{1cm}
\setlength{\parskip}{0pt}

From October 1 to December 5, 2018, we ran an elite survey with respondents from ten Latin American countries: Argentina, Bolivia, Brazil, Chile, Colombia, Costa Rica, Ecuador, Mexico, Panama, and Peru. The survey proceeded in two steps.

First, we collected information on Latin American elites. Our target sample included:

1. Executive members
2. Legislative members
3. Civil society
4. Academics

For each profile, in each country, we aimed at surveying:

1. 10 Executive members
2. 10 Legislative members
3. 15 Civil society
4. 15 Academics

To do so, we built a dataset of prospective respondents that was ten times larger than our sample goal of 500 elite members. We constructed the dataset by gathering information from government websites, local NGOs, newspapers, and university departments.

After finishing this dataset, we started the surveying stage. From November 12 to December 5, we ran the survey by contacting the elite members by phone or by email. We had two teams of enumerators, one based in São Paulo and another based in Rio de Janeiro, Brazil. The teams were comprised of Portuguese and Spanish native speakers. Although our initial expectation was to collect 500 survey responses, we completed 654 interviews.

# Conjoint Experiment Analysis: Identification and Estimation \label{sec:conjoint}

A conjoint experiment is a statistical technique that allows individuals to express their preferences on multiple attributes of a single topic [@hainmueller2014causal; @bansak2016economic]. Individuals are presented with two hypothetical scenarios, A and B, each containing a series of characteristics a researcher wants to evaluate. The individual chooses one of them. As the attributes are randomised and individuals choose between different pairs of hypothetical scenarios, we can estimate how individuals value each of the conjoined elements.

In our research, we ask what features elite members would like to see implemented in a climate change agreement. We ask their opinion on 6 questions: 1) who makes the rules; 2) who enforces the laws; 3) what punishment should be used against lawbreakers; 4) how are repeated violations punished; 5) how are costs distributed, and 6) how often the agreement will be renegotiated. Each of the questions has four to five different attributes. For the first two questions, we ask individuals if they would rather have the community, local governments, federal governments, or international organizations to create or enforce rules. As for punishments, we have options ranging from do nothing to fines and incarceration. We ask whether the costs of climate change agreements should be paid mostly or exclusively by developed countries, by polluting countries regardless of their wealth, or if costs should be allocated according to the history of emissions of a given nation. Individuals also state their preferences for short-term agreements, a proxy we use for flexibility, or for long-term ones, our proxy for the stability of rules.

Conjoint experiments have many advantages. First, as each individual chooses between many pairs of possible climate agreements---seven in our case---we can drastically increase our statistical power without incurring in further financial costs. Thus, the design maximizes our research budget. Second, individuals rarely decide considering one attribute at a time, as presented in other types of survey experiments. In that regard, our conjoint analysis mirrors how people naturally make their choices, that is, by simultaneously taking several characteristics into consideration. Finally, as the experiments consist of a simple choice between two bundles, conjoint analyses are easy to understand and to implement. This research design reduces the cognitive load of interviewees, so we can expect more accurate responses from our sample.

# Conjoint Experiment Dimensions \label{sec:conjdim}

In Table 1 of the main paper, we present the attributes and values for climate mitigation in our conjoint experiment. Here, we discuss the theoretical underpinnings of each of the dimensions and attributes evoked.

## Attribute One: *Who Makes the Rules*

Climate change governance faces various challenges in terms of national and local implementations. For example, the voluntary character of the Nationally Determined Contributions (NDCs), which are the core mechanism of the Paris Agreement, illustrates this decentralized nature. Climate governance can be observed at a number of different scales, such as the international, national, subnational, regional, and local levels [@bruhl2001international]. Microlevel implementation requires regulations and attribution of responsibilities for each level, for which we propose the first dimension of our conjoint experiment is *Who makes the rules?*. To mirror the choices available for climate agreement signatory countries, we included the following levels of ruling: international organizations; federal governments; local governments; local community members; and non-governmental organizations. These levels reveal differences in rule making and top-down versus bottom-up possible schemes.

Moving to the *International Organizations* dimension, currently there is no international institution that centralizes agenda-setting, rule making, conflict resolution, and law enforcement for climate-related issues. On the other hand, the Conference of Parties is the highest decision-making body of the UNFCCC (United Nations Framework Convention on Climate Change), and the UN Climate Convention could possibly evolve into a more institutionalized model of international governance. The very signing of the many agreements that comprise the broad climate mitigation international framework---the Rio Convention, the Kyoto Protocol, and the Paris Agreement---constitute a step in this direction of further institutionalization.

We can take the WTO (World Trade Organization) as an example. After the establishment of the GATT (General Agreement on Tariffs and Trade), trade organizations progressively increased their role as gatekeepers, issue entrepreneurs and proponents of international norms. In turn, this movement solidified the international trade regime and established dedicated international organizations. The process of continuous, and eventually formal, institutionalization of trade regimes has been discussed by the rational functionalist school of international relations [e.g., @simons2002handbook]. This mechanism could plausibly be valued by specific country elites as a reliable and legitimate potential source of climate mitigation rules.

We also include the *Federal government* as an alternative rule provider. Once a treaty is signed and ratified, it must be internalized by a minimum number of signatories in order to entry into force. In other words, federal governments have to comply with the engagement, which highlights the essential role states play in the implementation of international treaties. Despite the growing importance of transnational networks and non-state actors in global regimes, states remain the fundamental actors in the international realm.

Concerning climate-related agreements, the bare existence of international institutions is not sufficient for global climate governance. In fact, scholars have posited that these engagements can be understood as a discursive strategy, one which allows states to satisfy internal public opinion demands without taking costly global warming mitigation initiatives [e.g., @dimitrov2005hostage]. In that sense, elites may see the federal government as a more credible source of effective environmental legislation. Latin American countries provide examples of national-level climate legislation, such as Mexico's General Law for Climate Change (2012), Bolivia's Mother Earth Legal Framework (Law 300/2012), and Costa Rica's Biodiversity Law (1998).

Local actors, such as municipal governments, have also been active players in the creation of climate mitigation norms [@barber2013if; @fraundorfer2017role]. Significantly, cities have built a number of networks dedicated to climate issues, such as the ICLEI (Local Governments for Sustainability) and the C40. They include 94 cities and promote coordinated actions to fight climate change. While some countries opt to internalize and implement climate rules at the national level, others prefer to withdraw from the climate treaties and allow local governments to voluntarily comply with international conventions. The state of California is a recent example. Although the United States had chosen to withdraw from the Paris Agreement, the state government offered a plan to follow the mitigation schemes proposed by the treaty.

Likewise, climate mitigation practices and norms are strongly influenced by local community members and non-governmental organizations. The civil society has long played a substantive role in setting the agenda and monitoring compliance of climate legislation [@bruhl2001international]. This is also true for Latin America. Indigenous groups have been particularly active in the local sphere and the international arena, as climate change directly endangers traditional communities [@bellier2012peuples]. At the international level, we highlight the creation of the United Nations Permanent Forum on Indigenous Issues. Locally, Latin American communities have founded many non-governmental organizations to fight climate change, such as the _Asamblea Nacional de Afectados Ambientales_ (ANAA) in Mexico, the _Red de los Pueblos Fumigados_ in Argentina and the _Rede Brasileira de Justiça Ambiental_ (RBJA) in Brazil.

## Attribute Two: *Conflict Resolution Mechanism*

International regimes require efficient institutions to deal with non-compliance and cost reparation to aggrieved parties. We consider five institutions that can plausibly serve as instances for conflict mediation in climate issues: the United Nations; government bureaucracies; local courts; private arbitration; and informal norms.

International trade disputes, such as those concerning unfair practices, are already solved by international courts. The WTO Dispute Settlement Body and the International Criminal Court are two well-known examples. In this sense, the United Nations can also provide a similar resolution mechanism to solve climate disputes.

In other situations, government bureaucracies have the discretion to settle conflicts and enforce punishment to violators [@biesbroek2018public]. In Latin America, one example is the Brazilian Environmental Protection Agency (IBAMA), which is run by the federal government. The IBAMA is free to set rules nationally and punish any violation of environmental laws in Brazil. Another example is provided by the Costa Rican National Environmental Office (SETENA). It is responsible for receiving denunciations and conducting investigations on environmental crimes.

Local courts can also function as conflict resolution mechanisms in case of non-compliance. A widely-publicized case in Latin America involves the lawsuit filed by almost 30 thousand Ecuadorian citizens against oil companies Texaco/Chevron. The case was brought to a provincial court in Ecuador [@pigrau2014texaco]. Alternatively, private arbitration is a low-cost and prompt approach for the problem in the case of violations. Chile and Argentina resorted to private arbitration in 1972 to address a border dispute, and the Permanent Court of Arbitration frequently addresses environmental issues.

Informal norms could also be a plausible and low-cost form of solving conflicts. For example, non-institutionalized forms of reputational sanction within the climate international regime, such as *naming and shaming*, can act as a peer-pressure mechanism pushing for compliance. Note that informal norms within communities usually boost compliance in many situations and allow for more flexible and context adapted solutions (@ostrom1990governing).

## Attributes Three and Four: *Punishment* and *Punishment for Repeated Violations*

The third and fourth attributes relate to punishments against non-compliance. We design the first feature, *Punishment*, to capture what types of punishment elites would like to apply to climate agreement violators. The options we present closely match existing Latin American regulations on the subject. As such, they constitute plausible answers for the interviewees. For instance, in Brazil, violations are punishable with a range of options, from community services and fines to imprisonment (Law 9605/98, available in: <http://www.planalto.gov.br/ccivil_03/leis/l9605.htm>. That is also the case for Peru (Law 29263, available in: <http://www.minam.gob.pe/delitosambientales>). Similarly, Ecuador has recently approved legislation that adds severe penalties on perpetrators of climate violations to the national constitution. Incarceration, for instance, is one of the possible punishments (Article 245, Código Orgánico Integral Penal, available in: <http://www.oas.org/juridico/PDFs/mesicic5_ecu_ane_con_judi_c%C3%B3d_org_int_pen.pdf>). In this respect, our selected treaty characteristics are plausible to Latin American elites.

Following the same logic, the fourth experimental feature, *Punishment for repeated violations*, captures the actions that should be taken upon subsequent violations of climate change agreements.

## Attribute Five: *Agreement Costs*

The fifth design principle refers to the distribution of the agreement's costs. Here we consider the normative debates on climate equality, burden-sharing, and distributive justice in global public goods provision. More specifically, developing nations have stressed how wealthier countries have historically contributed more to global warming [@santos2017global]. For instance, under Copenhagen Accord and the Kyoto Protocol, developed countries were deemed as the sole responsible for CO2 emissions [@kartha2011comparison]. Over time, the wealthier nations successfully contested this position, arguing that emerging economies such as China and India accounted for a large share of greenhouse gas emissions as well. Hence, a new compensation principle emerged, that of *common but differentiated responsibilities and respective capabilities*. While accounting for past environmental pollution, the model also request the contribution of developing countries.

Our design feature allows us to capture the position of Latin American elites in this debate. We add four variables derived from the discussion presented above. The first posits that only rich countries should pay the coses of climate agreements, reflecting the original stance of developing nations. The second states that rich countries should pay more than poor countries, which also test the elites' willing to contribute to public goods. The third alternative states that countries should pay costs according to the history of emissions, while the last one affirms that only current emissions should be taken into account. Understanding how Latin American elites stand on this normative issue is essential for assuring compliance in the local level and for crafting national level climate governance schemes.

## Attribute Six: *Renegotiation*

Finally, our last conjoint experiment dimension investigates the rigidity of the agreement. From a theoretical standpoint, we know from the contract theory literature that more rigid agreements may fall short when their underlying conditions change. However, agreements that change constantly might be too unstable to form a path dependency and provide focal points for long-run coordination. Thus, parties have to strike a balance between legal stability and the need for adaptation. For instance, if a climate agreement has deleterious effects in the economy while only marginally improving slightly the country's environmental output, decision-makers will push for treaty renegotiations. Conversely, longer agreements reduce coordination costs and give predictability to the system. We analyse temporal preferences of Latin American elites by including five different climate agreement renegotiation schedules: one year, five years, twenty years, fifty years, or never.

Latin America has a history of political and economic volatility, so it is unclear whether elites will prefer long-term or short-term climate change initiatives. On the one hand, elites may perceive longer contracts as a means to force collective commitment to fighting global warming. On the other hand, elites can also anticipate future volatility discount the future more heavily in favor of immediate circumstances.

# Descriptive Statistics \label{sec:descstats}

The main demographic statistics we collected are: country indicators, elite types, and the geolocation of the respondents. We show their distributions in the graphs below.

```{r demogGraph, fig.align='center', fig.height=3.5, message=FALSE,warning=F}
library(plm); library(tidyverse)
library(haven); library(clusterSEs)
library(stargazer); library(cjoint)
library(sp); library(cregg)
library(kableExtra)
load('data.RData')

aux <- cj %>%
  select(Response.ID, countryOrigin, groupOrigin, LocationLongitude, LocationLatitude) %>%
  unique()

# Country
tab <- data.frame(table(aux$countryOrigin))
names(tab) <- c('Country', 'Frequency')
p<- ggplot(data = tab, aes(x = Country, y = Frequency)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 50, linetype = "dashed", color = "black") + 
  theme_bw() + annotate(geom = "text", label = tab$Frequency,
                        x = tab$Country, y = tab$Frequency + 2)
p

# Elite type
tab <- data.frame(table(aux$groupOrigin))
names(tab) <- c('Group Origin', 'Frequency')
p<- ggplot(data = tab, aes(x = `Group Origin`, y = Frequency)) +
  geom_bar(stat = "identity") +
  theme_bw() + annotate(geom = "text", label = tab$Frequency,
                        x = tab$`Group Origin`, y = tab$Frequency + 5)
p

# Elite type and group type
tab <- data.frame(table(aux$groupOrigin, aux$countryOrigin))
names(tab) <- c('Group Origin','Country', 'Frequency')
p<- ggplot(data = tab, aes(x = `Group Origin`, y = Frequency)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Country) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
p
```

As we can see, we managed to contact more than fifty elite members for all countries but Bolivia and Peru. To compensate for the low response rates in these two countries, we exceeded the sample size in all the remaining states.

Regarding elite type, we have more data on academics and members of civil society. They are followed by members the Executive and of the Legislative, respectively. This result is in line with our expectations given the difficulty of accessing high profile government members.

In the map, we see that respondents are concentrated in the largest municipalities of their respective countries. As elites tend to be more urban and well-educated than the average citizen, we expected the clustering of elites in big cities.

# Frequency of Features Selected by Each Attribute \label{sec:features}

The figures below display the freature frequency for each attribute, by each task.

```{r attributes, fig.align='center', fig.height=4}
aux <- cj %>%
  select(Response.ID, `Who makes the rules?`, `How are conflicts resolved?`, `What punishments do they use?`,`How are repeated violations punished?`, `How are costs distributed?`,`How often will the agreement be renegotiated?`) %>%
  unique()

# Who makes the rules
tab <- data.frame(table(aux$`Who makes the rules?`))
names(tab) <- c('Levels', 'Frequency')
p <- ggplot(data=tab, aes(x = Levels, y = Frequency)) +
  geom_bar(stat="identity") +
  theme_bw() + xlab('Who makes the rules?') +
  annotate(geom = "text", label = tab$Frequency,
           x = tab$Levels, y = tab$Frequency + 35) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p

# How are conflicts resolved?
tab <- data.frame(table(aux$`How are conflicts resolved?`))
names(tab) <- c('Levels', 'Frequency')
p <- ggplot(data=tab, aes(x = Levels, y = Frequency)) +
  geom_bar(stat="identity") +
  theme_bw() + xlab('How are conflicts resolved?') +
  annotate(geom = "text", label = tab$Frequency,
           x = tab$Levels, y = tab$Frequency + 35) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p

# What punishments do they use?
tab <- data.frame(table(aux$`What punishments do they use?`))
names(tab) <- c('Levels', 'Frequency')
p <- ggplot(data=tab, aes(x = Levels, y = Frequency)) +
  geom_bar(stat="identity") +
  theme_bw() + xlab('What punishments do they use?') +
  annotate(geom = "text", label = tab$Frequency,
           x = tab$Levels, y = tab$Frequency + 40) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p

# How are repeated violations punished?
tab <- data.frame(table(aux$`How are repeated violations punished?`))
names(tab) <- c('Levels', 'Frequency')
p <- ggplot(data=tab, aes(x = Levels, y = Frequency)) +
  geom_bar(stat="identity") +
  theme_bw() + xlab('How are repeated violations punished?') +
  annotate(geom = "text", label = tab$Frequency,
           x = tab$Levels, y = tab$Frequency + 50) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p

# How are costs distributed?
tab <- data.frame(table(aux$`How are costs distributed?`))
names(tab) <- c('Levels', 'Frequency')
p <- ggplot(data=tab, aes(x = Levels, y = Frequency)) +
  geom_bar(stat="identity") +
  theme_bw() + xlab('How are costs distributed?') +
  annotate(geom = "text", label = tab$Frequency,
           x = tab$Levels, y = tab$Frequency + 50) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p

# How often will the agreement be renegotiated?
tab <- data.frame(table(aux$`How often will the agreement be renegotiated?`))
names(tab) <- c('Levels', 'Frequency')
p <- ggplot(data=tab, aes(x = Levels, y = Frequency)) +
  geom_bar(stat="identity") +
  theme_bw() + xlab('How often will the agreement be renegotiated?') +
  annotate(geom = "text", label = tab$Frequency,
           x = tab$Levels, y = tab$Frequency + 50) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
p
```

Two remarkable issues in our data are the low frequency of *Less penalty* in the attribute *How are repeated violations punished?* and the low frequency of *None* for the attribute *What punishments do they use?*. This because when we draw the level none for the attribute *What punishments do they use?* or the level less penalty for *How are repeated violations punished?*, then it makes no sense to have less penalty than the minimum possible.

# Code for the Main Paper

The code for Figure 2 of the paper follows below. We also include the point estimates in companion tables.

```{r codefig2}
## Main plot
# Main estimation equation
fm <- selected ~ `Who makes the rules?` +
  `How are conflicts resolved?` +
  `What punishments do they use?` + 
  `How are repeated violations punished?` +
  `How are costs distributed?` +
  `How often will the agreement be renegotiated?`

# Plot
mms <- mm(cj, fm, id = ~Response.ID, alpha = .1, h0 = 0.5)

myFaces <- c(rep('plain', 5), "bold",
             rep('plain', 4), "bold",
             rep('plain', 3), "bold",
             rep('plain', 4), "bold",
             rep('plain', 5), "bold",
             rep('plain', 5), "bold")

p <- plot(mms, vline = 0.5, header_fmt = "%s", size = 2) +
  ggplot2::theme(
    legend.position = "none",
    panel.grid.major = ggplot2::element_blank(),
    panel.grid.minor = ggplot2::element_blank(),
    axis.text.y = element_text(face=myFaces, size = 11)) + 
    ggplot2::geom_errorbarh(ggplot2::aes_string(xmin = "lower",
                                                xmax = "upper"),
                            size = 0.5, height = 0, na.rm = TRUE)
p

ggsave(filename = 'MM_all.pdf', plot = p, width = 6, height = 5)

# Table
table_mm <- function(mms, capt) {
  dfr <- data.frame(feature = mms[,c(4)], 
             round(mms[,c(5,6,8,9,10)], digits=3))
  names(dfr) <- c('Feature', 'Estimate', 'Std.Error',
                'P-Value', 'Lower', 'Upper')
  return(kable(dfr, "latex", caption = capt, booktabs = T) %>%
    kable_styling(font_size = 10) %>%
    group_rows('Who makes the rules?', 1, 5) %>%
    group_rows('How are conflicts resolved?', 6, 10) %>%
    group_rows('What punishments do they use?', 11, 14) %>%
    group_rows('How are repeated violations punished?', 15, 17) %>%
    group_rows('How are costs distributed?', 18, 21) %>%
    group_rows('How often will the agreement be renegotiated?', 22, 26))
}

table_mm(mms, capt = 'Marginal Means -- Full Dataset')
```

The code for Figure 3 of the paper follows below, with also the point estimates in a companion table.

\newpage

```{r codefig3}
## Marginal means by Country

# Argentina
arg <- mm(subset(cj, countryOrigin=='Argentina'),  
        fm, id = ~Response.ID, alpha = 0.1, h0 = 0.5)
table_mm(arg, capt = 'Marginal Means -- Argentina Only')
```

\newpage

```{r}
# Bolivia
bol <- mm(subset(cj, countryOrigin=='Bolivia'),  
        fm, id = ~Response.ID, alpha = 0.1, h0 = 0.5)
table_mm(bol, capt = 'Marginal Means -- Bolivia Only')
```

\newpage

```{r}
# Brazil
bra <- mm(subset(cj, countryOrigin=='Brazil'),  
        fm, id = ~Response.ID, alpha = 0.1, h0 = 0.5)
table_mm(bra, capt = 'Marginal Means -- Brazil Only')
```

\newpage

```{r}
# Chile
chi <- mm(subset(cj, countryOrigin=='Chile'),  
        fm, id = ~Response.ID, alpha = 0.1, h0 = 0.5)
table_mm(chi, capt = 'Marginal Means -- Chile Only')
```

\newpage

```{r}
# Colombia
col <- mm(subset(cj, countryOrigin=='Colombia'),  
        fm, id = ~Response.ID, alpha = 0.1, h0 = 0.5)
table_mm(col, capt = 'Marginal Means -- Colombia Only')
```

\newpage

```{r}
# Costa Rica
cri <- mm(subset(cj, countryOrigin=='Costa Rica'),  
        fm, id = ~Response.ID, alpha = 0.1, h0 = 0.5)
table_mm(cri, capt = 'Marginal Means -- Costa Rica Only')
```

\newpage

```{r}
# Ecuador
ecu <- mm(subset(cj, countryOrigin=='Ecuador'),  
        fm, id = ~Response.ID, alpha = 0.1, h0 = 0.5)
table_mm(ecu, capt = 'Marginal Means -- Ecuador Only')
```

\newpage

```{r}
# Mexico
mex <- mm(subset(cj, countryOrigin=='Mexico'),  
        fm, id = ~Response.ID, alpha = 0.1, h0 = 0.5)
table_mm(mex, capt = 'Marginal Means -- Mexico Only')
```

\newpage

```{r}
# Panama
pan <- mm(subset(cj, countryOrigin=='Panama'),  
        fm, id = ~Response.ID, alpha = 0.1, h0 = 0.5)
table_mm(pan, capt = 'Marginal Means -- Panama Only')
```

\newpage

```{r}
# Peru
per <- mm(subset(cj, countryOrigin=='Peru'),  
        fm, id = ~Response.ID, alpha = 0.1, h0 = 0.5)
table_mm(per, capt = 'Marginal Means -- Peru Only')
```

\newpage

```{r, fig.width = 13, fig.height = 13}
# Changing country labels
arg$country <- 'Argentina'
bol$country <- 'Bolivia'
bra$country <- 'Brazil'
chi$country <- 'Chile'
col$country <- 'Colombia'
cri$country <- 'Costa Rica'
ecu$country <- 'Ecuador'
mex$country <- 'Mexico'
pan$country <- 'Panama'
per$country <- 'Peru'
# Plot by country
p <- plot(rbind(arg, bol, bra, chi, col, 
              cri, ecu, mex, pan, per), 
        group = 'country', vline = 0.5, nr = 10,
        header_fmt = "%s", size = 2) +
  facet_wrap( ~ country, ncol = 5) +
  ggplot2::theme(
    axis.text.y = element_text(face=myFaces, size = 13)) + 
    ggplot2::geom_errorbarh(ggplot2::aes_string(xmin = "lower",
                                                xmax = "upper"),
                            size = 0.75, height = 0, na.rm = TRUE)

p

ggsave(filename = 'MM_country.pdf', 
       plot = p, width = 10, height = 12)
```

The code for Figure 4 of the paper is available below. Tables displaying the point estimates are also included.

```{r}
## Elite type graphs and estimates

# Table for the Executive
res1 <- mm(subset(cj, groupOrigin=='Executive'), 
           fm, id = ~Response.ID, alpha = 0.1, h0 = 0.5)
table_mm(res1, capt = 'Marginal Means -- Executive Only')
```

\newpage

```{r}
# Legislative
res2 <- mm(subset(cj, groupOrigin=='Legislative'), 
           fm, id = ~Response.ID, alpha = 0.1, h0 = 0.5)
table_mm(res2, capt = 'Marginal Means -- Legislative Only')
```

\newpage

```{r}
# Civil Society
res3 <- mm(subset(cj, groupOrigin=='Civil Society'), 
           fm, id = ~Response.ID, alpha = 0.1, h0 = 0.5)
table_mm(res3, capt = 'Marginal Means -- Civil Society Only')
```

\newpage

```{r}
# Academia
res4 <- mm(subset(cj, groupOrigin=='Academia'), 
           fm, id = ~Response.ID, alpha = 0.1, h0 = 0.5)
table_mm(res4, capt = 'Marginal Means -- Academia Only')

```

\newpage

```{r, fig.width = 10, fig.height = 8}
# Changing labels
res1$MemberType <- 'Executive'
res2$MemberType <- 'Legislative'
res3$MemberType <- 'Civil Society'
res4$MemberType <- 'Academia'

# Plot by Elite Type
p <- plot(rbind(res1, res2, res3, res4), 
     group = 'MemberType', vline = 0.5, nr = 4,
        header_fmt = "%s", size = 2) + 
  facet_wrap( ~ MemberType, ncol = 4) +
  ggplot2::theme(
    axis.text.y = element_text(face=myFaces, size = 11)) + 
    ggplot2::geom_errorbarh(ggplot2::aes_string(xmin = "lower",
                                                xmax = "upper"),
                            size = 0.5, height = 0, na.rm = TRUE,
                            position = ggstance::position_dodgev(height = 1))
p

ggsave(filename = 'MM_membertype.pdf', plot = p,
       width = 8, height = 6)
```

# Average Marginal Component Effect (AMCE) Estimator \label{sec:amce}

This estimator fixes one category and look at changes from this baseline category. Below follows the plots for the main model and the country and elite type subsamples.

```{r, fig.width = 7, fig.height = 8}
## Main model
rm(cj)
load('data.RData')

# Set conjoint baselines
baselines <- list()
baselines$`How often will the agreement be renegotiated?` <- "One year"
baselines$`What punishments do they use?` <- 'None'
baselines$`How are conflicts resolved?` <- 'Government bureaucracy'
baselines$`How are costs distributed?` <- 'Rich countries pay more than poor countries'
baselines$`Who makes the rules?` <- 'Federal government'

attrs <- c("Who makes the rules?",
           "How are conflicts resolved?",
           "What punishments do they use?",
           "How are repeated violations punished?",
           "How are costs distributed?",
           "How often will the agreement be renegotiated?")

results <- cjoint::amce(fm, data = cj, cluster = TRUE, 
                        respondent.id = "Response.ID", 
                        design = conjDesign, 
                        baselines = baselines, na.ignore = T)

# Table AMCE -- Full Model 
tableAMCE <- function(results, capt = 'Main Model') {
  aux <- plotAMCE(results, ci = 0.9, point.size = .8, dodge.size = 1,
     text.size = 10, main = 'AMCE -- Full Model',
     attribute.names = attrs, tblfy = T)
  aux <- aux %>%
    filter(!is.na(group)) %>%
    select(printvar, pe, se, lower, upper)
  indx <- grep('Baseline =', aux$printvar)
  aux$pe <- round(aux$pe, digits = 3)
  aux$se <- round(aux$se, digits = 3)
  aux$lower <- round(aux$lower, digits = 3)
  aux$upper <- round(aux$upper, digits = 3)
  aux$pe[indx] = ''
  aux$se[indx] = ''
  aux$lower[indx] = ''
  aux$upper[indx] = ''
  names(aux) <- c('Feature', 'Estimate', 'Std.Error', 
                  'Lower', 'Upper')
  return(kable(aux, "latex", caption = capt, booktabs = T, align = c('l', rep('c', 4))) %>%
    kable_styling(font_size = 10) %>%
    group_rows('How are conflicts resolved?', 1, 5) %>%
    group_rows('How are costs distributed?', 6, 9) %>%
    group_rows('How are repeated violations punished?', 10, 12) %>%
    group_rows('How often will the agreement be renegotiated?', 13, 17) %>%
    group_rows('What punishments do they use?', 18, 21) %>%
    group_rows('Who makes the rules?', 22, 26))
}

tableAMCE(results, capt = 'AMCE -- Full Model')
```

\newpage

```{r, fig.width = 7, fig.height = 8}
# Plot AMCE -- Full Model
plot(results, ci = 0.9, point.size = .8, dodge.size = 1,
     text.size = 10, main = 'AMCE -- Full Model',
     attribute.names = attrs)
```

\newpage

```{r}
## Country subsample

## Argentina
results <- cjoint::amce(fm, data = cj, cluster = TRUE,
                        respondent.id = "Response.ID", 
                        design = conjDesign, baselines = baselines,
                        subset = cj$countryOrigin == 'Argentina')
# AMCE -- Argentina Only
tableAMCE(results, capt = 'AMCE -- Argentina Only')
```

\newpage

```{r, fig.width = 7, fig.height = 8}
plot(results, ci=0.9, point.size = .8, dodge.size = 1,
     text.size = 10, main = 'AMCE -- Argentina Only',
     attribute.names = attrs)
```

\newpage

```{r}
## Bolivia
results <- cjoint::amce(fm, data = cj, cluster = TRUE,
                        respondent.id = "Response.ID", 
                        design = conjDesign, baselines = baselines,
                        subset = cj$countryOrigin == 'Bolivia')
# Table AMCE -- Full Model 
tableAMCE(results, capt = 'AMCE -- Bolivia Only')
```

\newpage

```{r, fig.width = 7, fig.height = 8}
plot(results, ci=0.9, point.size = .8, dodge.size = 1,
     text.size = 10, main = 'AMCE -- Bolivia Only',
     attribute.names = attrs)
```

\newpage

```{r}
## Brazil
results <- cjoint::amce(fm, data = cj, cluster = TRUE,
                        respondent.id = "Response.ID", 
                        design = conjDesign, baselines = baselines,
                        subset = cj$countryOrigin == 'Brazil')
# Table AMCE -- Brazil Only
tableAMCE(results, capt = 'AMCE -- Brazil Only')
```

\newpage

```{r, fig.width = 7, fig.height = 8}
plot(results, ci=0.9, point.size = .8, dodge.size = 1,
     text.size = 10, main = 'AMCE -- Brazil Only',
     attribute.names = attrs)
```

\newpage

```{r}
## Chile
results <- cjoint::amce(fm, data = cj, cluster = TRUE,
                        respondent.id = "Response.ID", 
                        design = conjDesign, baselines = baselines,
                        subset = cj$countryOrigin == 'Chile')
# Table AMCE -- Chile Only
tableAMCE(results, capt = 'AMCE -- Chile Only')
```

\newpage

```{r, fig.width = 7, fig.height = 8}
plot(results, ci=0.9, point.size = .8, dodge.size = 1,
     text.size = 10, main = 'AMCE -- Chile Only',
     attribute.names = attrs)
```

\newpage

```{r}
## Colombia
results <- cjoint::amce(fm, data = cj, cluster = TRUE,
                        respondent.id = "Response.ID", 
                        design = conjDesign, baselines = baselines,
                        subset = cj$countryOrigin == 'Colombia')
# Table AMCE -- Colombia Only
tableAMCE(results, capt = 'AMCE -- Colombia Only')
```

\newpage

```{r, fig.width = 7, fig.height = 8}
plot(results, ci=0.9, point.size = .8, dodge.size = 1,
     text.size = 10, main = 'AMCE -- Colombia Only',
     attribute.names = attrs)
```

\newpage

```{r}
## Costa Rica
results <- cjoint::amce(fm, data = cj, cluster = TRUE,
                        respondent.id = "Response.ID", 
                        design = conjDesign, baselines = baselines,
                        subset = cj$countryOrigin == 'Costa Rica')
# Table AMCE -- Costa Rica Only
tableAMCE(results, capt = 'AMCE -- Costa Rica Only')
```

\newpage

```{r, fig.width = 7, fig.height = 8}
plot(results, ci=0.9, point.size = .8, dodge.size = 1,
     text.size = 10, main = 'AMCE -- Costa Rica Only',
     attribute.names = attrs)
```

\newpage

```{r}
## Ecuador
results <- cjoint::amce(fm, data = cj, cluster = TRUE,
                        respondent.id = "Response.ID", 
                        design = conjDesign, baselines = baselines,
                        subset = cj$countryOrigin == 'Ecuador')
# Table AMCE -- Ecuador Only
tableAMCE(results, capt = 'AMCE -- Ecuador Only')
```

\newpage

```{r, fig.width = 7, fig.height = 8}
plot(results, ci=0.9, point.size = .8, dodge.size = 1,
     text.size = 10, main = 'AMCE -- Ecuador Only',
     attribute.names = attrs)
```

\newpage

```{r}
## Mexico
results <- cjoint::amce(fm, data = cj, cluster = TRUE,
                        respondent.id = "Response.ID", 
                        design = conjDesign, baselines = baselines,
                        subset = cj$countryOrigin == 'Mexico')
# Table AMCE -- Full Model 
tableAMCE(results, capt = 'AMCE -- Mexico Only')
```

\newpage

```{r, fig.width = 7, fig.height = 8}
plot(results, ci=0.9, point.size = .8, dodge.size = 1,
     text.size = 10, main = 'AMCE -- Mexico Only',
     attribute.names = attrs)
```

\newpage

```{r}
## Panama
results <- cjoint::amce(fm, data = cj, cluster = TRUE,
                        respondent.id = "Response.ID", 
                        design = conjDesign, baselines = baselines,
                        subset = cj$countryOrigin == 'Panama')
# Table AMCE -- Panama Only
tableAMCE(results, capt = 'AMCE -- Panama Only')
```

\newpage

```{r, fig.width = 7, fig.height = 8}
plot(results, ci=0.9, point.size = .8, dodge.size = 1,
     text.size = 10, main = 'AMCE -- Panama Only',
     attribute.names = attrs)
```

\newpage

```{r}
## Peru
results <- cjoint::amce(fm, data = cj, cluster = TRUE,
                        respondent.id = "Response.ID", 
                        design = conjDesign, baselines = baselines,
                        subset = cj$countryOrigin == 'Peru')
# Table AMCE -- Peru Only
tableAMCE(results, capt = 'AMCE -- Peru Only')
```

\newpage

```{r, fig.width = 7, fig.height = 8}
plot(results, ci=0.9, point.size = .8, dodge.size = 1,
     text.size = 10, main = 'AMCE -- Peru Only',
     attribute.names = attrs)
```

\newpage

```{r}
## Elite Subsample

# Executive
results <- cjoint::amce(fm, data = cj, cluster = TRUE,
                        respondent.id = "Response.ID", 
                        design = conjDesign, baselines = baselines,
                        subset = cj$groupOrigin == 'Executive')
# Table AMCE -- Executive Only
tableAMCE(results, capt = 'AMCE -- Executive Only')
```

\newpage

```{r, fig.width = 7, fig.height = 8}
plot(results, ci=0.9, point.size = .8, dodge.size = 1,
     text.size = 10, main = 'AMCE -- Executive Only',
     attribute.names = attrs)
```

\newpage

```{r}
## Legislative
results <- cjoint::amce(fm, data = cj, cluster = TRUE,
                        respondent.id = "Response.ID", 
                        design = conjDesign, baselines = baselines,
                        subset = cj$groupOrigin == 'Legislative')
# Table AMCE -- Legislative Only
tableAMCE(results, capt = 'AMCE -- Legislative Only')
```

\newpage

```{r, fig.width = 7, fig.height = 8}
plot(results, ci=0.9, point.size = .8, dodge.size = 1,
     text.size = 10, main = 'AMCE -- Legislative Only',
     attribute.names = attrs)
```

\newpage

```{r}
## Civil society
results <- cjoint::amce(fm, data = cj, cluster = TRUE,
                        respondent.id = "Response.ID", 
                        design = conjDesign, baselines = baselines,
                        subset = cj$groupOrigin == 'Civil Society')
# Table AMCE -- Civil Society Only 
tableAMCE(results, capt = 'AMCE -- Civil Society Only')
```

\newpage

```{r, fig.width = 7, fig.height = 8}
plot(results, ci=0.9, point.size = .8, dodge.size = 1,
     text.size = 10, main = 'AMCE -- Civil Society Only',
     attribute.names = attrs)
```

\newpage

```{r}
## Academia Only
results <- cjoint::amce(fm, data = cj, cluster = TRUE,
                        respondent.id = "Response.ID", 
                        design = conjDesign, baselines = baselines,
                        subset = cj$groupOrigin == 'Academia')
# Table AMCE -- Full Model 
tableAMCE(results, capt = 'AMCE -- Academia Only')
```

\newpage

```{r, fig.width = 7, fig.height = 8}
plot(results, ci=0.9, point.size = .8, dodge.size = 1,
     text.size = 10, main = 'AMCE -- Academia Only',
     attribute.names = attrs)
```

\newpage

# APSA Experimental Section Standard Report for Experiments

## Hypothesis

The experiment was designed to study the characteristics of the climate change mitigation treaties favored by Latin American elites.

## Subjects and Context

The eligibility criteria for the research was to belong to one of the four elite profiles listed below.

1. **Executive members**: members of regulatory agencies, politicians (mayors, governors, presidents), members of ministries and secretaries at the federal and state levels.
2. **Legislative members**: legislators and staff officers from the federal and state level in the researched countries.
3. **Civil society**: environmental and energy-related NGO members; oil, renewables, and environmental solutions firm owners in the researched countries.
4. **Academics**: professors from energy and engineering departments of the most prestigious universities in the countries researched.

We decided to include these profiles based on the stakes and influence that such elites have on climate change policies. According to constructivist international relations theories, academics have notable influence on our topic of interest as they form an epistemic community around climate change policies. Civil society members that have stakes on climate policies are also influential, such as oil-extracting companies, lobby groups, and environmental NGOs. Executive and Legislative elites handle the decisions themselves, what justifies their inclusion in our sample.

We selected ten countries to our sample: Argentina, Bolivia, Brazil, Chile, Colombia, Costa Rica, Ecuador, Mexico, Panama, and Peru. Our selection criteria were: 1) the impact of climate change on each nation, 2) their size and regional importance, and 3) natural resources availability. For each country, we built a dataset of potential respondents with at least:

* 150 potential legislative respondents
* 150 potential executive respondents
* 200 potential civil society respondents
* 200 potential academic respondents

These numbers represent ten times the number of respondents we were aiming to interview by the end of the survey. The dataset was build from October 5th to November 12th, 2018. We hired a team of enumerators that searched the internet for people in the countries that fulfilled our pre-determined elite profiles.

We ranked the possible respondents in two tiers. In the first tier we placed people that we had complete information about their profiles: office telephone numbers, emails, social media contacts, and so on. In the second tier we included participants that had incomplete profiles, such as those with only online information available. The dataset including the first-tier respondents was randomly divided between enumerators in Rio de Janeiro and São Paulo to avoid any eventual biases at the collecting stage. Both groups conducted the interviews by phone from November 12th to December 5th, 2018. The second tier dataset was used for the online only version of the survey.

The response rate in each of the enumeration sites and in the online only survey follows in the table below.

\vspace{.5cm}

| Country    | Enumeration Place | Type             | Total | Phone  | Online |
|:-----------|:------------------|:-----------------|------:|-------:|-------:|
| Argentina  | Rio de Janeiro    | Online and Phone | 300   | 33     | 17     |
| Bolivia    | Rio de Janeiro    | Online and Phone | 372   | 31     | 11     |
| Brazil     | Rio de Janeiro    | Online and Phone | 346   | 34     | 12     |
| Chile      | Rio de Janeiro    | Online and Phone | 417   | 35     | 12     |
| Colombia   | Rio de Janeiro    | Online and Phone | 445   | 35     | 21     |
| Costa Rica | Rio de Janeiro    | Online and Phone | 409   | 32     | 11     |
| Ecuador    | Rio de Janeiro    | Online and Phone | 512   | 30     | 12     |
| Mexico     | Rio de Janeiro    | Online and Phone | 468   | 41     | 17     |
| Panama     | Rio de Janeiro    | Online and Phone | 291   | 30     | 14     |
| Peru       | Rio de Janeiro    | Online and Phone | 243   | 32     | 11     |
| *TOTAL*    |                   | Online and Phone | 3803  | 333    | 138    |
|            |                   |                  |       |        |        |
| Argentina  | Sao Paulo         | Online and Phone | 292   | 48     | 16     |
| Bolivia    | Sao Paulo         | Online and Phone | 373   | 50     | 13     |
| Brazil     | Sao Paulo         | Online and Phone | 344   | 43     | 15     |
| Chile      | Sao Paulo         | Online and Phone | 416   | 41     | 15     |
| Colombia   | Sao Paulo         | Online and Phone | 448   | 42     | 19     |
| Costa Rica | Sao Paulo         | Online and Phone | 412   | 44     | 20     |
| Ecuador    | Sao Paulo         | Online and Phone | 507   | 43     | 18     |
| Mexico     | Sao Paulo         | Online and Phone | 464   | 41     | 17     |
| Panama     | Sao Paulo         | Online and Phone | 291   | 50     | 19     |
| Peru       | Sao Paulo         | Online and Phone | 248   | 44     | 9      |
| *TOTAL*    |                   | Online and Phone | 3803  | 446    | 161    |
|            |                   |                  |       |        |        |
| Argentina  |                   | Online Only      | 517   |        | 19     |
| Bolivia    |                   | Online Only      | 132   |        | 12     |
| Brazil     |                   | Online Only      | 1183  |        | 52     |
| Chile      |                   | Online Only      | 470   |        | 31     |
| Colombia   |                   | Online Only      | 522   |        | 29     |
| Costa Rica |                   | Online Only      | 325   |        | 20     |
| Ecuador    |                   | Online Only      | 460   |        | 37     |
| Mexico     |                   | Online Only      | 955   |        | 29     |
| Panama     |                   | Online Only      | 319   |        | 26     |
| Peru       |                   | Online Only      | 540   |        | 25     |
| *TOTAL*    |                   | Online Only      | 5569  |        | 293    |
Table: Survey response rates.

The response rate was 10.24 percent of the original population in the telephone survey (779 of 7606 possible respondents). We had an attrition rate of 61.62 percent from the telephone (779) to the online survey (299). The response rate for the online only survey was 5.26 percent (293 of 5569 invited by email).

The conjoint experiment analyzed in this paper was in the online and online only dataset. Therefore, the relevant column for our data is the last one.

## Allocation Methods

**Random Assignment:** We wrote the computer code for our experiment using a Python application provided by @strezhnev2013conjoint. We translated the original survey questions (as shown in the text) into Portuguese and Spanish, and then embedded the PHP file with the randomization parameters in a Qualtrics survey. The PHP code we used to randomize the values of the attributes is available at the project's GitHub repository: <https://github.com/danilofreire/climate-governance>.

For each attribute, the probability of selecting a given component follows a uniform distribution function. The table below describes the results.

\vspace{.5cm}

| **Attribute**                          | **Values**                                   | **Probabilities** |
|:---------------------------------------|:---------------------------------------------|------------------:|
| **Who makes the rules?**               | International organizations                  | 1/5               | 
|                                        | Federal government                           | 1/5               |
|                                        | Local government                             | 1/5               |
|                                        | Local community members                      | 1/5               |
|                                        | Non-governmental organizations               | 1/5               |
|                                        |                                              |                   |
| **Conflict resolution mechanism**      | United Nations                               | 1/5               |
|                                        | Government bureaucracy                       | 1/5               |
|                                        | Local courts                                 | 1/5               |
|                                        | Private arbitration                          | 1/5               |
|                                        | Informal norms                               | 1/5               |
|                                        |                                              |                   |
| **Punishment**                         | Imprisonment                                 | 1/4               |
|                                        | Fines                                        | 1/4               |
|                                        | Blacklist                                    | 1/4               |
|                                        | None                                         | 1/4               |
|                                        |                                              |                   |
| **Punishment for repeated violations** | More penalty                                 | 1/3               |
|                                        | Same                                         | 1/3               |
|                                        | Less penalty                                 | 1/3               |
|                                        |                                              |                   |
| **Agreement costs**                    | Rich countries pay more than poor countries  | 1/4               |
|                                        | Proportional to history of emissions         | 1/4               |
|                                        | Proportional to current emissions            | 1/4               |
|                                        | Only rich countries pay                      | 1/4               |
|                                        |                                              |                   |
| **Renegotiation**                      | Never                                        | 1/5               |
|                                        | Fifty years                                  | 1/5               |
|                                        | Twenty years                                 | 1/5               |
|                                        | Five years                                   | 1/5               |
|                                        | One year                                     | 1/5               |
Table: Conjoint experiment attributes and their respective probabilities.

We added one logical restriction to the set climate change treaties: **Punishment = None** can never appear together with **Punishment for repeated violations = Less penalty**. The number of possible treaties are the product of the attributes ($5\times 5 \times 4 \times 3 \times 4 \times 5 = 6,000$) minus the number of removed possibilities ($5\times 5  \times 4 \times 5 = 500$). The total number of possible treaties is $5,500$.

## Treatments

**Descriptions of the intervention:** Our treatment was the randomization of components in the conjoined climate mitigation treaties. We provided the following description before the conjoint experiment starts, in Spanish and Portuguese:

\vspace{.5cm}

```
For the following questions, read carefully the instructions below.

Imagine that your country would sign an international treaty to mitigate climate change. 
A climate mitigation treaty has, in general, the following attributes:

1. Who defines the rules? Which group will define the parameters for the treaty?
2. How are the conflicts that might happen when the treaty begin be resolved?
3. What punishments should be applied to ensure compliance?
4. How will repeated violations be punished?
5. How are costs for implementing the treaty be distributed?
6. How often the treaty has to be renegotiated?

In the following questions, you will see variations in treaties on these six attributes.
Please select the treaty that has the characteristics that you believe are best for your country.
```

\vspace{.5cm}

The conjoint experiment consisted in comparing two hypothetical climate change treaties that vary across six attributes. We presented each respondent with seven pairs of possible climate agreements.

\noindent \textbf{Software:} We used the Python application provided by @strezhnev2013conjoint to program the PHP randomization device. We hosted the survey questionnaire on Qualtrics.

\noindent \textbf{Delivery:} We repeated the conjoint experiment seven times, to improve test power.

\noindent \textbf{Dataset description:} Each line in the dataset corresponds to a given treaty, which indicates the selection status of a given treaty (selected versus non-selected) for a specific respondent. When a respondent finished an entire response set, this added 14 rows to the dataset. Seven rows with the selected treaties and other seven rows with the non-selected treaties.

\noindent \textbf{Deception:} We used no deception in this survey experiment.

## Results

\noindent \textbf{Outcome measures and covariates:} The main outcome is a binary indicator for the selection of a given conjoined package. We added two covariates to measure heterogeneous effects: *Country* and *Elite type*.

\noindent \textbf{Statistical analysis:} We fit a marginal means estimator and an AMCE estimator. The estimation conditions the results within subjects and the presented conjoined packages. The results are available in sections \ref{sec:features} and \ref{sec:amce}.

\noindent \textbf{Missing data and attrition:} Conditional on having started the conjoint experiment, for each task we had the following completion rates:

```{r, fig.width = 7, fig.height = 5, tidy=FALSE}
barplot(100*table(cj$task)/max(table(cj$task)),
        xlab = 'Conjoint task',
        ylab = 'Percentage completed')
```

## Other information

\noindent \textbf{IRB:} This research received IRB approval from Brown University (Protocol 2195/2018) and Fundação Getulio Vargas (Protocol 83/2018).

\noindent \textbf{Pre-registration:} The result was not pre-registered.

\noindent \textbf{Funding:} The research was funded by Konrad Adenauer Stiftung (KAS). KAS provided EUR$ 55,000.00 for this research. They have not interfered in the research design and in the question choices.

\noindent \textbf{Replication materials:} Available at <http://github.com/danilofreire/climate-governance>.

# Bibliography