### 2020-05-27 supplementary-material/supplementary-material-climate-policy.tex
Accepted version

% !TEX program = XeLaTeX
\documentclass[12pt,a4paper,]{article}

% Fonts
\usepackage[]{libertine}
% This is a nice mathfont, it fits well with Libertine
% Comment the line if you want to use a different one
\usepackage[libertine]{newtxmath}
% The default monofont 
% Again, comment the line below if you want to change it
\usepackage[scaled=.95]{inconsolata}

% Add the following packages to support kableExtra
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{etoolbox}
\usepackage{tocloft}

% Dots in toc
\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}

% Colours
\usepackage[usenames,dvipsnames]{xcolor}
\definecolor{darkblue}{rgb}{0.0,0.0,0.55}

% Spacing
\usepackage{setspace}

% Margin
\usepackage[top=2cm,bottom=2cm,left=2cm,right=2cm]{geometry}

% Packages I've been using for different reasons...
\usepackage[backref,pagebackref]{hyperref}
\usepackage{dcolumn}
\usepackage{graphicx}
\usepackage{float}
\floatplacement{figure}{H}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{positioning}
\usepackage{mathtools}
\usepackage{caption}

% UK English
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\cleanlookdateon

% Penalties
\exhyphenpenalty=1000
\hyphenpenalty=1000
\widowpenalty=1000
\clubpenalty=1000

% Back references
\renewcommand*{\backref}[1]{}
\renewcommand*{\backrefalt}[4]{%
    \ifcase #1 (Not cited.)%
    \or        Cited on page~#2.%
    \else      Cited on pages~#2.%
    \fi}
\renewcommand{\backreftwosep}{ and~} 
\renewcommand{\backreflastsep}{ and~}

% Hypersetup
\hypersetup{
  linkcolor=Mahogany,
  citecolor=Mahogany,
  urlcolor=darkblue, 
  breaklinks=true, 
  colorlinks=true,
      pdfauthor={Danilo Freire; Umberto Mignozzetti; David Skarbek},
    }

% If XeTex, LuaLaTeX, etc
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{amssymb,amsmath}
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}

% Language

% Bibliography
\usepackage{natbib}
\bibliographystyle{apalike}
\makeatletter
% Remove comma after author
\setcitestyle{aysep={}}
\patchcmd{\NAT@citex}
	  {\@citea\NAT@hyper@{%
		 \NAT@nmfmt{\NAT@nm}%
		 \hyper@natlinkbreak{\NAT@aysep\NAT@spacechar}{\@citeb\@extra@b@citeb}%
		 \NAT@date}}
	  {\@citea\NAT@nmfmt{\NAT@nm}%
	   \NAT@aysep\NAT@spacechar\NAT@hyper@{\NAT@date}}{}{}
	\patchcmd{\NAT@citex}
	  {\@citea\NAT@hyper@{%
		 \NAT@nmfmt{\NAT@nm}%
		 \hyper@natlinkbreak{\NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi}%
		   {\@citeb\@extra@b@citeb}%
		 \NAT@date}}
	  {\@citea\NAT@nmfmt{\NAT@nm}%
	   \NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi\NAT@hyper@{\NAT@date}}
	  {}{}
% Patch case where name and year are separated by aysep
\patchcmd{\NAT@citex}
  {\@citea\NAT@hyper@{%
     \NAT@nmfmt{\NAT@nm}%
     \hyper@natlinkbreak{\NAT@aysep\NAT@spacechar}{\@citeb\@extra@b@citeb}%
     \NAT@date}}
  {\@citea\NAT@nmfmt{\NAT@nm}%
   \NAT@aysep\NAT@spacechar\NAT@hyper@{\NAT@date}}{}{}
% Patch case where name and year are separated by opening bracket
\patchcmd{\NAT@citex}
  {\@citea\NAT@hyper@{%
     \NAT@nmfmt{\NAT@nm}%
     \hyper@natlinkbreak{\NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi}%
       {\@citeb\@extra@b@citeb}%
     \NAT@date}}
  {\@citea\NAT@nmfmt{\NAT@nm}%
   \NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi\NAT@hyper@{\NAT@date}}
  {}{}
\makeatother

% Listings
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

% Verbatim

% Tables
\usepackage{longtable,booktabs}

% Graphics
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}

% Make links footnotes instead of hotlinks:
%  \setlength{\emergencystretch}{3em}  % prevent overfull lines
 \providecommand{\tightlist}{%
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Numbered sections
\setcounter{secnumdepth}{5}
% % % Redefines (sub)paragraphs to behave more like sections
% \ifx\paragraph\undefined\else
% \let\oldparagraph\paragraph
% \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
% \fi
% \ifx\subparagraph\undefined\else
% \let\oldsubparagraph\subparagraph
% \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
% \fi
% 
% Spacing
\doublespacing

% Title
\title{Supplementary Materials for ``Institutional Design and Elite Support for
Climate Policies: Evidence from Latin American Countries''}

% Author
\author{Danilo Freire\footnote{Postdoctoral Research Associate, The Political
  Theory Project, Brown University, Providence, RI 02912, USA,
  \href{mailto:danilofreire@brown.edu}{\nolinkurl{danilofreire@brown.edu}},
  \url{http://danilofreire.github.io}. Corresponding author.} \and Umberto Mignozzetti\footnote{School of International Relations, Fundação
  Getulio Vargas, São Paulo, SP, Brazil and Wilf Family Department of
  Politics, NYU, NY, USA,
  \href{mailto:umberto.mig@nyu.edu}{\nolinkurl{umberto.mig@nyu.edu}},
  \url{http://umbertomig.com}.} \and David Skarbek\footnote{The Department of Political Science and the
  Political Theory Project, Brown University, Providence, RI, USA,
  \href{mailto:davidskarbek@brown.edu}{\nolinkurl{davidskarbek@brown.edu}},
  \url{http://davidskarbek.com}.}}

% Date
\date{\today}

% Begin document
\begin{document}
\maketitle

% Abstract

% Table of Contents
{
\hypersetup{linkcolor=black}
\setcounter{tocdepth}{2}
\tableofcontents
}
\hypertarget{latin-american-elite-survey}{%
\section{\texorpdfstring{Latin American Elite Survey
\label{sec:latam}}{Latin American Elite Survey }}\label{latin-american-elite-survey}}

\setlength{\parindent}{1cm}
\setlength{\parskip}{0pt}

From October 1 to December 5, 2018, we ran an elite survey with
respondents from ten Latin American countries: Argentina, Bolivia,
Brazil, Chile, Colombia, Costa Rica, Ecuador, Mexico, Panama, and Peru.
The survey proceeded in two steps.

We collected information on Latin American elites. The target was:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Members of the Executive branch
\item
  Members of the Legislative branch
\item
  Civil society
\item
  Academics
\end{enumerate}

For each profile, in each country, we aimed at surveying:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  10 members of the Executive branch
\item
  10 members of the Legislative branch
\item
  15 Civil society
\item
  15 Academics
\end{enumerate}

To do so, we built a prospective respondents dataset with ten times
these members for each country. Then, we ran an online and telephone
survey where we collected information regarding the climate change
agreements and other climate change-related problems.

After finishing this dataset, we started the surveying stage. From
November 12 to December 5, we ran the survey by contacting elite members
by phone or by email. We had two teams of enumerators, one based in São
Paulo and another based in Rio de Janeiro, Brazil, comprised of
Portuguese and Spanish native speakers. Although the overall aim of the
survey was 500 responses, the teams managed to collect 654 interviews.

\hypertarget{conjoint-experiment-analysis-identification-and-estimation}{%
\section{\texorpdfstring{Conjoint Experiment Analysis: Identification
and Estimation
\label{sec:conjoint}}{Conjoint Experiment Analysis: Identification and Estimation }}\label{conjoint-experiment-analysis-identification-and-estimation}}

A conjoint experiment is a statistical technique that allows individuals
to express their preferences on multiple attributes of a single topic
\citep{hainmueller2014causal, bansak2016economic}. Individuals are
presented with two hypothetical scenarios, A and B, each containing a
series of characteristics a researcher wants to evaluate. The individual
chooses one of them. As the attributes are randomized and individuals
choose between different pairs of hypothetical scenarios, we can
estimate how individuals value each of the conjoined elements.

In our research, we ask what features elite members would like to see
implemented in a climate change agreement. We ask their opinion on 6
questions: 1) who makes the rules; 2) who enforces the laws; 3) what
punishment should be used against lawbreakers; 4) how are repeated
violations punished; 5) how are costs distributed, and 6) how often the
agreement will be renegotiated. Each of the questions has four to five
different attributes. For the first two questions, we ask individuals if
they would rather have the community, local governments, federal
governments, or international organizations to create or enforce rules.
As for punishments, we have options ranging from do nothing to fines and
incarceration. We ask whether the costs of climate change agreements
should be paid mostly or exclusively by developed countries, by polluter
countries regardless of their wealth, or if costs should be allocated
according to the history of emissions of a given nation. Individuals
also state their preferences for short-term agreements, a proxy we use
for flexibility, or for long-term ones, what we see as a proxy for the
stability of rules. Lastly, we ask which of the two conjoined choices
individuals perceive as more legitimate and more efficient.

Conjoint experiments have many advantages \citep{hainmueller2014causal}.
First, as each individual chooses between many pairs of possible climate
agreements -- seven in our case -- we can drastically increase our
sample size without incurring in further financial costs. The design
thus maximizes our research budget. Second, individuals rarely decide
considering one attribute at a time, as presented in other types of
survey experiments. In that regard, the conjoint analysis mirrors how
people naturally make their choices, that is, by simultaneously taking
several characteristics into consideration. Finally, as the experiments
consist of a simple choice between A and B, conjoint analyses are easy
to understand and to implement. This research design reduces the
cognitive load interviewees have while answering surveys and as a
result, we can expect more accurate responses from our sample.

\hypertarget{descriptive-statistics}{%
\section{\texorpdfstring{Descriptive Statistics
\label{sec:descstats}}{Descriptive Statistics }}\label{descriptive-statistics}}

The main demographic statistics collected were the country and elite
types, and the geolocation of the respondents. They follow in the graphs
below.

\begin{center}\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/demogGraph-1} \end{center}

\begin{center}\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/demogGraph-2} \end{center}

As we can see, we managed to contact more than fifty elite members for
all countries but Bolivia and Peru. To compensate for the low collection
rates in these two countries, note that we exceeded the sample size in
all the remaining states.

Considering the elite type, note that we have, as expected, more people
from academia and the civil society, followed by those from the
Executive and Legislative branches. The data collection for members of
the Legislative branch was below the original target, but above the
minimum mark of 5 for each country. This is expected, given the
difficulty of accessing these elite members.

In the map, we can see that the answers concentrate on the country's
capitals, with some variation within the state, but with sample
concentrated in the largest municipalities. This is also in line with
our expectations, as elite members are assumed to gravitate around
capital cities and large urban centers.

\hypertarget{frequency-of-features-selected-by-each-attribute}{%
\section{\texorpdfstring{Frequency of Features Selected by Each
Attribute
\label{sec:features}}{Frequency of Features Selected by Each Attribute }}\label{frequency-of-features-selected-by-each-attribute}}

The figures below display the feature frequency for each attribute, by
each task.

\begin{center}\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/attributes-1} \end{center}

\begin{center}\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/attributes-2} \end{center}

\begin{center}\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/attributes-3} \end{center}

\begin{center}\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/attributes-4} \end{center}

\begin{center}\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/attributes-5} \end{center}

\begin{center}\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/attributes-6} \end{center}

The two remarkable issues are the low frequency of \emph{Less penalty}
in the attribute \emph{How are repeated violations punished?} and the
low frequency of \emph{None} for the attribute \emph{What punishments do
they use?}. This because when we draw the level none for the attribute
\emph{What punishments do they use?} or the level less penalty for
\emph{How are repeated violations punished?}, then it makes no sense to
have less penalty than the minimum possible.

\hypertarget{code-for-the-main-paper}{%
\section{Code for the Main Paper}\label{code-for-the-main-paper}}

The code for Figure 2 of the paper follows below. We also include the
point estimates in companion tables.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## Main plot}
\CommentTok{# Main estimation equation}
\NormalTok{fm <-}\StringTok{ }\NormalTok{selected }\OperatorTok{~}\StringTok{ `}\DataTypeTok{Who makes the rules?}\StringTok{`} \OperatorTok{+}
\StringTok{  `}\DataTypeTok{How are conflicts resolved?}\StringTok{`} \OperatorTok{+}
\StringTok{  `}\DataTypeTok{What punishments do they use?}\StringTok{`} \OperatorTok{+}\StringTok{ }
\StringTok{  `}\DataTypeTok{How are repeated violations punished?}\StringTok{`} \OperatorTok{+}
\StringTok{  `}\DataTypeTok{How are costs distributed?}\StringTok{`} \OperatorTok{+}
\StringTok{  `}\DataTypeTok{How often will the agreement be renegotiated?}\StringTok{`}

\CommentTok{# Plot}
\NormalTok{mms <-}\StringTok{ }\KeywordTok{mm}\NormalTok{(cj, fm, }\DataTypeTok{id =} \OperatorTok{~}\NormalTok{Response.ID, }\DataTypeTok{alpha =} \FloatTok{.1}\NormalTok{, }\DataTypeTok{h0 =} \FloatTok{0.5}\NormalTok{)}

\NormalTok{myFaces <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\StringTok{'plain'}\NormalTok{, }\DecValTok{5}\NormalTok{), }\StringTok{"bold"}\NormalTok{,}
             \KeywordTok{rep}\NormalTok{(}\StringTok{'plain'}\NormalTok{, }\DecValTok{4}\NormalTok{), }\StringTok{"bold"}\NormalTok{,}
             \KeywordTok{rep}\NormalTok{(}\StringTok{'plain'}\NormalTok{, }\DecValTok{3}\NormalTok{), }\StringTok{"bold"}\NormalTok{,}
             \KeywordTok{rep}\NormalTok{(}\StringTok{'plain'}\NormalTok{, }\DecValTok{4}\NormalTok{), }\StringTok{"bold"}\NormalTok{,}
             \KeywordTok{rep}\NormalTok{(}\StringTok{'plain'}\NormalTok{, }\DecValTok{5}\NormalTok{), }\StringTok{"bold"}\NormalTok{,}
             \KeywordTok{rep}\NormalTok{(}\StringTok{'plain'}\NormalTok{, }\DecValTok{5}\NormalTok{), }\StringTok{"bold"}\NormalTok{)}

\NormalTok{p <-}\StringTok{ }\KeywordTok{plot}\NormalTok{(mms, }\DataTypeTok{vline =} \FloatTok{0.5}\NormalTok{, }\DataTypeTok{header_fmt =} \StringTok{"%s"}\NormalTok{, }\DataTypeTok{size =} \DecValTok{2}\NormalTok{) }\OperatorTok{+}\StringTok{ }\NormalTok{ggplot2}\OperatorTok{::}\KeywordTok{theme}\NormalTok{(}
    \DataTypeTok{legend.position =} \StringTok{"none"}\NormalTok{,}
    \DataTypeTok{panel.grid.major =}\NormalTok{ ggplot2}\OperatorTok{::}\KeywordTok{element_blank}\NormalTok{(),}
    \DataTypeTok{panel.grid.minor =}\NormalTok{ ggplot2}\OperatorTok{::}\KeywordTok{element_blank}\NormalTok{(),}
    \DataTypeTok{axis.text.y =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{face=}\NormalTok{myFaces, }\DataTypeTok{size =} \DecValTok{11}\NormalTok{)) }\OperatorTok{+}\StringTok{ }
\StringTok{    }\NormalTok{ggplot2}\OperatorTok{::}\KeywordTok{geom_errorbarh}\NormalTok{(ggplot2}\OperatorTok{::}\KeywordTok{aes_string}\NormalTok{(}\DataTypeTok{xmin =} \StringTok{"lower"}\NormalTok{,}
                                                \DataTypeTok{xmax =} \StringTok{"upper"}\NormalTok{),}
                            \DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{height =} \DecValTok{0}\NormalTok{, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{,}
                            \DataTypeTok{position =}\NormalTok{ ggstance}\OperatorTok{::}\KeywordTok{position_dodgev}\NormalTok{(}\DataTypeTok{height =} \DecValTok{1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: Vectorized input to `element_text()` is not officially supported.
## Results may be unexpected or may change in future versions of ggplot2.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{p}
\end{Highlighting}
\end{Shaded}

\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/codefig2-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggsave}\NormalTok{(}\DataTypeTok{filename =} \StringTok{'MM_all.pdf'}\NormalTok{, }\DataTypeTok{plot =}\NormalTok{ p, }\DataTypeTok{width =} \DecValTok{8}\NormalTok{, }\DataTypeTok{height =} \DecValTok{6}\NormalTok{)}

\CommentTok{# Table}
\NormalTok{table_mm <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(mms, capt) \{}
\NormalTok{  dfr <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{feature =}\NormalTok{ mms[,}\KeywordTok{c}\NormalTok{(}\DecValTok{4}\NormalTok{)], }
             \KeywordTok{round}\NormalTok{(mms[,}\KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{,}\DecValTok{6}\NormalTok{,}\DecValTok{8}\NormalTok{,}\DecValTok{9}\NormalTok{,}\DecValTok{10}\NormalTok{)], }\DataTypeTok{digits=}\DecValTok{3}\NormalTok{))}
  \KeywordTok{names}\NormalTok{(dfr) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{'Feature'}\NormalTok{, }\StringTok{'Estimate'}\NormalTok{, }\StringTok{'Std.Error'}\NormalTok{,}
                \StringTok{'P-Value'}\NormalTok{, }\StringTok{'Lower'}\NormalTok{, }\StringTok{'Upper'}\NormalTok{)}
  \KeywordTok{return}\NormalTok{(}\KeywordTok{kable}\NormalTok{(dfr, }\StringTok{"latex"}\NormalTok{, }\DataTypeTok{caption =}\NormalTok{ capt, }\DataTypeTok{booktabs =}\NormalTok{ T) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{font_size =} \DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{group_rows}\NormalTok{(}\StringTok{'Who makes the rules?'}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{5}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{group_rows}\NormalTok{(}\StringTok{'How are conflicts resolved?'}\NormalTok{, }\DecValTok{6}\NormalTok{, }\DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{group_rows}\NormalTok{(}\StringTok{'What punishments do they use?'}\NormalTok{, }\DecValTok{11}\NormalTok{, }\DecValTok{14}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{group_rows}\NormalTok{(}\StringTok{'How are repeated violations punished?'}\NormalTok{, }\DecValTok{15}\NormalTok{, }\DecValTok{17}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{group_rows}\NormalTok{(}\StringTok{'How are costs distributed?'}\NormalTok{, }\DecValTok{18}\NormalTok{, }\DecValTok{21}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{group_rows}\NormalTok{(}\StringTok{'How often will the agreement be renegotiated?'}\NormalTok{, }\DecValTok{22}\NormalTok{, }\DecValTok{26}\NormalTok{))}
\NormalTok{\}}

\KeywordTok{table_mm}\NormalTok{(mms, }\DataTypeTok{capt =} \StringTok{'Marginal Means -- Full Dataset'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:codefig2}Marginal Means -- Full Dataset}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrrrrr}
\toprule
Feature & Estimate & Std.Error & P-Value & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}NGOs & 0.455 & 0.013 & 0.000 & 0.434 & 0.476\\
\hspace{1em}Local community members & 0.516 & 0.013 & 0.202 & 0.495 & 0.537\\
\hspace{1em}Local government & 0.498 & 0.012 & 0.901 & 0.478 & 0.519\\
\hspace{1em}Federal government & 0.486 & 0.013 & 0.271 & 0.466 & 0.507\\
\hspace{1em}International organizations & 0.543 & 0.012 & 0.001 & 0.522 & 0.563\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}Informal norms & 0.410 & 0.013 & 0.000 & 0.390 & 0.431\\
\hspace{1em}Private arbitration & 0.500 & 0.013 & 0.980 & 0.480 & 0.521\\
\hspace{1em}Local courts & 0.546 & 0.012 & 0.000 & 0.526 & 0.567\\
\hspace{1em}Government bureaucracy & 0.465 & 0.013 & 0.005 & 0.444 & 0.485\\
\hspace{1em}United Nations & 0.573 & 0.012 & 0.000 & 0.553 & 0.593\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}None & 0.448 & 0.013 & 0.000 & 0.427 & 0.470\\
\hspace{1em}Blacklist & 0.523 & 0.011 & 0.035 & 0.505 & 0.540\\
\hspace{1em}Fines & 0.553 & 0.011 & 0.000 & 0.536 & 0.571\\
\hspace{1em}Imprisonment & 0.458 & 0.011 & 0.000 & 0.440 & 0.476\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}Less penalty & 0.464 & 0.011 & 0.001 & 0.447 & 0.481\\
\hspace{1em}Same & 0.497 & 0.009 & 0.739 & 0.482 & 0.512\\
\hspace{1em}More penalty & 0.532 & 0.009 & 0.001 & 0.517 & 0.548\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}Only rich countries pay & 0.438 & 0.011 & 0.000 & 0.420 & 0.456\\
\hspace{1em}Proportional to current emissions & 0.524 & 0.011 & 0.030 & 0.506 & 0.542\\
\hspace{1em}Proportional to history of emissions & 0.536 & 0.011 & 0.001 & 0.518 & 0.555\\
\hspace{1em}Rich countries pay more than poor countries & 0.505 & 0.011 & 0.684 & 0.486 & 0.523\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}One year & 0.535 & 0.012 & 0.004 & 0.515 & 0.556\\
\hspace{1em}Five years & 0.590 & 0.012 & 0.000 & 0.570 & 0.610\\
\hspace{1em}Twenty years & 0.519 & 0.012 & 0.121 & 0.499 & 0.540\\
\hspace{1em}Fifty years & 0.442 & 0.013 & 0.000 & 0.421 & 0.463\\
\hspace{1em}Never & 0.404 & 0.013 & 0.000 & 0.383 & 0.424\\
\bottomrule
\end{tabular}
\end{table}

The code for Figure 3 of the paper follows below, with also the point
estimates in a companion table.

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## Marginal means by Country}

\CommentTok{# Argentina}
\NormalTok{arg <-}\StringTok{ }\KeywordTok{mm}\NormalTok{(}\KeywordTok{subset}\NormalTok{(cj, countryOrigin}\OperatorTok{==}\StringTok{'Argentina'}\NormalTok{),  }
\NormalTok{        fm, }\DataTypeTok{id =} \OperatorTok{~}\NormalTok{Response.ID, }\DataTypeTok{alpha =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{h0 =} \FloatTok{0.5}\NormalTok{)}
\KeywordTok{table_mm}\NormalTok{(arg, }\DataTypeTok{capt =} \StringTok{'Marginal Means -- Argentina Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:codefig3}Marginal Means -- Argentina Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrrrrr}
\toprule
Feature & Estimate & Std.Error & P-Value & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}NGOs & 0.417 & 0.043 & 0.052 & 0.346 & 0.487\\
\hspace{1em}Local community members & 0.449 & 0.044 & 0.246 & 0.376 & 0.521\\
\hspace{1em}Local government & 0.503 & 0.042 & 0.933 & 0.435 & 0.572\\
\hspace{1em}Federal government & 0.504 & 0.045 & 0.928 & 0.429 & 0.579\\
\hspace{1em}International organizations & 0.619 & 0.041 & 0.004 & 0.551 & 0.686\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}Informal norms & 0.377 & 0.043 & 0.004 & 0.307 & 0.447\\
\hspace{1em}Private arbitration & 0.522 & 0.047 & 0.638 & 0.445 & 0.599\\
\hspace{1em}Local courts & 0.574 & 0.042 & 0.074 & 0.506 & 0.643\\
\hspace{1em}Government bureaucracy & 0.446 & 0.044 & 0.217 & 0.374 & 0.518\\
\hspace{1em}United Nations & 0.568 & 0.041 & 0.097 & 0.501 & 0.635\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}None & 0.361 & 0.043 & 0.001 & 0.289 & 0.432\\
\hspace{1em}Blacklist & 0.514 & 0.037 & 0.710 & 0.453 & 0.575\\
\hspace{1em}Fines & 0.587 & 0.037 & 0.019 & 0.526 & 0.647\\
\hspace{1em}Imprisonment & 0.494 & 0.037 & 0.882 & 0.433 & 0.556\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}Less penalty & 0.465 & 0.035 & 0.319 & 0.406 & 0.523\\
\hspace{1em}Same & 0.516 & 0.032 & 0.608 & 0.464 & 0.569\\
\hspace{1em}More penalty & 0.514 & 0.034 & 0.686 & 0.458 & 0.569\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}Only rich countries pay & 0.408 & 0.035 & 0.009 & 0.350 & 0.466\\
\hspace{1em}Proportional to current emissions & 0.534 & 0.041 & 0.410 & 0.466 & 0.601\\
\hspace{1em}Proportional to history of emissions & 0.576 & 0.038 & 0.045 & 0.514 & 0.638\\
\hspace{1em}Rich countries pay more than poor countries & 0.500 & 0.041 & 1.000 & 0.432 & 0.568\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}One year & 0.604 & 0.039 & 0.008 & 0.539 & 0.669\\
\hspace{1em}Five years & 0.545 & 0.043 & 0.298 & 0.474 & 0.616\\
\hspace{1em}Twenty years & 0.438 & 0.044 & 0.157 & 0.367 & 0.510\\
\hspace{1em}Fifty years & 0.438 & 0.044 & 0.154 & 0.365 & 0.510\\
\hspace{1em}Never & 0.448 & 0.046 & 0.263 & 0.372 & 0.524\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Bolivia}
\NormalTok{bol <-}\StringTok{ }\KeywordTok{mm}\NormalTok{(}\KeywordTok{subset}\NormalTok{(cj, countryOrigin}\OperatorTok{==}\StringTok{'Bolivia'}\NormalTok{),  }
\NormalTok{        fm, }\DataTypeTok{id =} \OperatorTok{~}\NormalTok{Response.ID, }\DataTypeTok{alpha =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{h0 =} \FloatTok{0.5}\NormalTok{)}
\KeywordTok{table_mm}\NormalTok{(bol, }\DataTypeTok{capt =} \StringTok{'Marginal Means -- Bolivia Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-1}Marginal Means -- Bolivia Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrrrrr}
\toprule
Feature & Estimate & Std.Error & P-Value & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}NGOs & 0.421 & 0.051 & 0.119 & 0.338 & 0.504\\
\hspace{1em}Local community members & 0.575 & 0.053 & 0.159 & 0.488 & 0.662\\
\hspace{1em}Local government & 0.560 & 0.054 & 0.272 & 0.470 & 0.649\\
\hspace{1em}Federal government & 0.467 & 0.048 & 0.498 & 0.388 & 0.547\\
\hspace{1em}International organizations & 0.494 & 0.056 & 0.910 & 0.401 & 0.586\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}Informal norms & 0.369 & 0.053 & 0.013 & 0.282 & 0.456\\
\hspace{1em}Private arbitration & 0.543 & 0.051 & 0.408 & 0.458 & 0.627\\
\hspace{1em}Local courts & 0.538 & 0.052 & 0.467 & 0.453 & 0.623\\
\hspace{1em}Government bureaucracy & 0.397 & 0.055 & 0.064 & 0.306 & 0.489\\
\hspace{1em}United Nations & 0.612 & 0.048 & 0.020 & 0.533 & 0.691\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}None & 0.442 & 0.054 & 0.278 & 0.354 & 0.530\\
\hspace{1em}Blacklist & 0.466 & 0.044 & 0.431 & 0.394 & 0.537\\
\hspace{1em}Fines & 0.570 & 0.044 & 0.108 & 0.498 & 0.642\\
\hspace{1em}Imprisonment & 0.505 & 0.048 & 0.923 & 0.425 & 0.584\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}Less penalty & 0.451 & 0.042 & 0.238 & 0.382 & 0.519\\
\hspace{1em}Same & 0.534 & 0.038 & 0.362 & 0.472 & 0.597\\
\hspace{1em}More penalty & 0.507 & 0.043 & 0.864 & 0.437 & 0.578\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}Only rich countries pay & 0.424 & 0.045 & 0.094 & 0.349 & 0.499\\
\hspace{1em}Proportional to current emissions & 0.593 & 0.047 & 0.050 & 0.515 & 0.670\\
\hspace{1em}Proportional to history of emissions & 0.427 & 0.046 & 0.112 & 0.352 & 0.503\\
\hspace{1em}Rich countries pay more than poor countries & 0.569 & 0.047 & 0.147 & 0.491 & 0.647\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}One year & 0.485 & 0.050 & 0.763 & 0.402 & 0.567\\
\hspace{1em}Five years & 0.624 & 0.048 & 0.010 & 0.544 & 0.703\\
\hspace{1em}Twenty years & 0.575 & 0.053 & 0.159 & 0.488 & 0.662\\
\hspace{1em}Fifty years & 0.385 & 0.055 & 0.036 & 0.294 & 0.475\\
\hspace{1em}Never & 0.402 & 0.053 & 0.063 & 0.316 & 0.489\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Brazil}
\NormalTok{bra <-}\StringTok{ }\KeywordTok{mm}\NormalTok{(}\KeywordTok{subset}\NormalTok{(cj, countryOrigin}\OperatorTok{==}\StringTok{'Brazil'}\NormalTok{),  }
\NormalTok{        fm, }\DataTypeTok{id =} \OperatorTok{~}\NormalTok{Response.ID, }\DataTypeTok{alpha =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{h0 =} \FloatTok{0.5}\NormalTok{)}
\KeywordTok{table_mm}\NormalTok{(bra, }\DataTypeTok{capt =} \StringTok{'Marginal Means -- Brazil Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-2}Marginal Means -- Brazil Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrrrrr}
\toprule
Feature & Estimate & Std.Error & P-Value & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}NGOs & 0.492 & 0.036 & 0.829 & 0.433 & 0.551\\
\hspace{1em}Local community members & 0.495 & 0.036 & 0.886 & 0.436 & 0.554\\
\hspace{1em}Local government & 0.456 & 0.036 & 0.219 & 0.397 & 0.515\\
\hspace{1em}Federal government & 0.481 & 0.036 & 0.610 & 0.422 & 0.541\\
\hspace{1em}International organizations & 0.563 & 0.033 & 0.053 & 0.509 & 0.617\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}Informal norms & 0.434 & 0.037 & 0.079 & 0.373 & 0.496\\
\hspace{1em}Private arbitration & 0.475 & 0.035 & 0.477 & 0.416 & 0.533\\
\hspace{1em}Local courts & 0.556 & 0.035 & 0.114 & 0.498 & 0.615\\
\hspace{1em}Government bureaucracy & 0.476 & 0.035 & 0.485 & 0.418 & 0.533\\
\hspace{1em}United Nations & 0.547 & 0.033 & 0.160 & 0.492 & 0.601\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}None & 0.465 & 0.038 & 0.356 & 0.402 & 0.528\\
\hspace{1em}Blacklist & 0.523 & 0.031 & 0.461 & 0.472 & 0.573\\
\hspace{1em}Fines & 0.569 & 0.030 & 0.021 & 0.520 & 0.618\\
\hspace{1em}Imprisonment & 0.434 & 0.029 & 0.024 & 0.386 & 0.482\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}Less penalty & 0.416 & 0.029 & 0.004 & 0.368 & 0.464\\
\hspace{1em}Same & 0.518 & 0.027 & 0.490 & 0.475 & 0.562\\
\hspace{1em}More penalty & 0.549 & 0.026 & 0.063 & 0.506 & 0.592\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}Only rich countries pay & 0.429 & 0.033 & 0.028 & 0.375 & 0.482\\
\hspace{1em}Proportional to current emissions & 0.496 & 0.031 & 0.901 & 0.445 & 0.547\\
\hspace{1em}Proportional to history of emissions & 0.546 & 0.032 & 0.152 & 0.493 & 0.599\\
\hspace{1em}Rich countries pay more than poor countries & 0.524 & 0.030 & 0.431 & 0.474 & 0.574\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}One year & 0.535 & 0.034 & 0.305 & 0.479 & 0.591\\
\hspace{1em}Five years & 0.568 & 0.036 & 0.064 & 0.508 & 0.627\\
\hspace{1em}Twenty years & 0.537 & 0.034 & 0.275 & 0.481 & 0.593\\
\hspace{1em}Fifty years & 0.469 & 0.035 & 0.368 & 0.412 & 0.526\\
\hspace{1em}Never & 0.377 & 0.037 & 0.001 & 0.317 & 0.437\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Chile}
\NormalTok{chi <-}\StringTok{ }\KeywordTok{mm}\NormalTok{(}\KeywordTok{subset}\NormalTok{(cj, countryOrigin}\OperatorTok{==}\StringTok{'Chile'}\NormalTok{),  }
\NormalTok{        fm, }\DataTypeTok{id =} \OperatorTok{~}\NormalTok{Response.ID, }\DataTypeTok{alpha =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{h0 =} \FloatTok{0.5}\NormalTok{)}
\KeywordTok{table_mm}\NormalTok{(chi, }\DataTypeTok{capt =} \StringTok{'Marginal Means -- Chile Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-3}Marginal Means -- Chile Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrrrrr}
\toprule
Feature & Estimate & Std.Error & P-Value & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}NGOs & 0.400 & 0.043 & 0.020 & 0.329 & 0.471\\
\hspace{1em}Local community members & 0.630 & 0.039 & 0.001 & 0.566 & 0.694\\
\hspace{1em}Local government & 0.538 & 0.040 & 0.335 & 0.473 & 0.604\\
\hspace{1em}Federal government & 0.403 & 0.043 & 0.025 & 0.332 & 0.474\\
\hspace{1em}International organizations & 0.496 & 0.043 & 0.932 & 0.426 & 0.567\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}Informal norms & 0.403 & 0.042 & 0.022 & 0.333 & 0.473\\
\hspace{1em}Private arbitration & 0.480 & 0.045 & 0.652 & 0.406 & 0.554\\
\hspace{1em}Local courts & 0.606 & 0.039 & 0.006 & 0.543 & 0.670\\
\hspace{1em}Government bureaucracy & 0.445 & 0.041 & 0.183 & 0.378 & 0.513\\
\hspace{1em}United Nations & 0.545 & 0.042 & 0.275 & 0.477 & 0.614\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}None & 0.425 & 0.043 & 0.081 & 0.355 & 0.496\\
\hspace{1em}Blacklist & 0.607 & 0.036 & 0.003 & 0.547 & 0.666\\
\hspace{1em}Fines & 0.512 & 0.035 & 0.726 & 0.455 & 0.570\\
\hspace{1em}Imprisonment & 0.435 & 0.036 & 0.076 & 0.376 & 0.495\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}Less penalty & 0.476 & 0.036 & 0.512 & 0.416 & 0.536\\
\hspace{1em}Same & 0.438 & 0.030 & 0.041 & 0.388 & 0.488\\
\hspace{1em}More penalty & 0.583 & 0.031 & 0.007 & 0.532 & 0.634\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}Only rich countries pay & 0.462 & 0.040 & 0.338 & 0.397 & 0.527\\
\hspace{1em}Proportional to current emissions & 0.524 & 0.037 & 0.508 & 0.464 & 0.585\\
\hspace{1em}Proportional to history of emissions & 0.497 & 0.037 & 0.940 & 0.436 & 0.559\\
\hspace{1em}Rich countries pay more than poor countries & 0.511 & 0.037 & 0.768 & 0.450 & 0.571\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}One year & 0.418 & 0.042 & 0.050 & 0.350 & 0.487\\
\hspace{1em}Five years & 0.645 & 0.038 & 0.000 & 0.582 & 0.708\\
\hspace{1em}Twenty years & 0.503 & 0.042 & 0.934 & 0.435 & 0.572\\
\hspace{1em}Fifty years & 0.489 & 0.043 & 0.795 & 0.417 & 0.560\\
\hspace{1em}Never & 0.424 & 0.043 & 0.078 & 0.353 & 0.495\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Colombia}
\NormalTok{col <-}\StringTok{ }\KeywordTok{mm}\NormalTok{(}\KeywordTok{subset}\NormalTok{(cj, countryOrigin}\OperatorTok{==}\StringTok{'Colombia'}\NormalTok{),  }
\NormalTok{        fm, }\DataTypeTok{id =} \OperatorTok{~}\NormalTok{Response.ID, }\DataTypeTok{alpha =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{h0 =} \FloatTok{0.5}\NormalTok{)}
\KeywordTok{table_mm}\NormalTok{(col, }\DataTypeTok{capt =} \StringTok{'Marginal Means -- Colombia Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-4}Marginal Means -- Colombia Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrrrrr}
\toprule
Feature & Estimate & Std.Error & P-Value & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}NGOs & 0.430 & 0.042 & 0.090 & 0.361 & 0.498\\
\hspace{1em}Local community members & 0.595 & 0.038 & 0.012 & 0.533 & 0.658\\
\hspace{1em}Local government & 0.465 & 0.038 & 0.359 & 0.403 & 0.528\\
\hspace{1em}Federal government & 0.463 & 0.037 & 0.327 & 0.402 & 0.525\\
\hspace{1em}International organizations & 0.544 & 0.041 & 0.285 & 0.476 & 0.611\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}Informal norms & 0.410 & 0.038 & 0.018 & 0.347 & 0.472\\
\hspace{1em}Private arbitration & 0.510 & 0.040 & 0.808 & 0.443 & 0.576\\
\hspace{1em}Local courts & 0.533 & 0.039 & 0.394 & 0.469 & 0.596\\
\hspace{1em}Government bureaucracy & 0.466 & 0.039 & 0.385 & 0.401 & 0.531\\
\hspace{1em}United Nations & 0.584 & 0.039 & 0.031 & 0.520 & 0.648\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}None & 0.497 & 0.038 & 0.939 & 0.434 & 0.560\\
\hspace{1em}Blacklist & 0.531 & 0.034 & 0.370 & 0.474 & 0.587\\
\hspace{1em}Fines & 0.551 & 0.033 & 0.123 & 0.497 & 0.606\\
\hspace{1em}Imprisonment & 0.414 & 0.035 & 0.013 & 0.357 & 0.471\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}Less penalty & 0.479 & 0.034 & 0.537 & 0.423 & 0.535\\
\hspace{1em}Same & 0.492 & 0.029 & 0.772 & 0.444 & 0.539\\
\hspace{1em}More penalty & 0.523 & 0.029 & 0.417 & 0.476 & 0.571\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}Only rich countries pay & 0.475 & 0.034 & 0.459 & 0.420 & 0.530\\
\hspace{1em}Proportional to current emissions & 0.516 & 0.034 & 0.633 & 0.460 & 0.572\\
\hspace{1em}Proportional to history of emissions & 0.538 & 0.035 & 0.286 & 0.480 & 0.596\\
\hspace{1em}Rich countries pay more than poor countries & 0.468 & 0.038 & 0.402 & 0.406 & 0.531\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}One year & 0.547 & 0.038 & 0.221 & 0.484 & 0.609\\
\hspace{1em}Five years & 0.558 & 0.039 & 0.137 & 0.494 & 0.621\\
\hspace{1em}Twenty years & 0.547 & 0.039 & 0.232 & 0.482 & 0.612\\
\hspace{1em}Fifty years & 0.423 & 0.042 & 0.069 & 0.354 & 0.493\\
\hspace{1em}Never & 0.417 & 0.037 & 0.026 & 0.356 & 0.478\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Costa Rica}
\NormalTok{cri <-}\StringTok{ }\KeywordTok{mm}\NormalTok{(}\KeywordTok{subset}\NormalTok{(cj, countryOrigin}\OperatorTok{==}\StringTok{'Costa Rica'}\NormalTok{),  }
\NormalTok{        fm, }\DataTypeTok{id =} \OperatorTok{~}\NormalTok{Response.ID, }\DataTypeTok{alpha =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{h0 =} \FloatTok{0.5}\NormalTok{)}
\KeywordTok{table_mm}\NormalTok{(cri, }\DataTypeTok{capt =} \StringTok{'Marginal Means -- Costa Rica Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-5}Marginal Means -- Costa Rica Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrrrrr}
\toprule
Feature & Estimate & Std.Error & P-Value & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}NGOs & 0.524 & 0.044 & 0.593 & 0.451 & 0.597\\
\hspace{1em}Local community members & 0.439 & 0.042 & 0.146 & 0.370 & 0.508\\
\hspace{1em}Local government & 0.565 & 0.043 & 0.134 & 0.494 & 0.636\\
\hspace{1em}Federal government & 0.511 & 0.043 & 0.798 & 0.441 & 0.581\\
\hspace{1em}International organizations & 0.459 & 0.048 & 0.387 & 0.380 & 0.537\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}Informal norms & 0.402 & 0.043 & 0.021 & 0.331 & 0.472\\
\hspace{1em}Private arbitration & 0.496 & 0.044 & 0.930 & 0.424 & 0.569\\
\hspace{1em}Local courts & 0.497 & 0.042 & 0.934 & 0.428 & 0.565\\
\hspace{1em}Government bureaucracy & 0.500 & 0.048 & 1.000 & 0.421 & 0.579\\
\hspace{1em}United Nations & 0.609 & 0.043 & 0.011 & 0.538 & 0.680\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}None & 0.383 & 0.045 & 0.010 & 0.308 & 0.457\\
\hspace{1em}Blacklist & 0.477 & 0.038 & 0.541 & 0.414 & 0.539\\
\hspace{1em}Fines & 0.588 & 0.036 & 0.014 & 0.529 & 0.647\\
\hspace{1em}Imprisonment & 0.506 & 0.039 & 0.877 & 0.442 & 0.569\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}Less penalty & 0.461 & 0.035 & 0.261 & 0.403 & 0.518\\
\hspace{1em}Same & 0.481 & 0.034 & 0.584 & 0.425 & 0.537\\
\hspace{1em}More penalty & 0.554 & 0.033 & 0.107 & 0.499 & 0.608\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}Only rich countries pay & 0.396 & 0.039 & 0.007 & 0.332 & 0.460\\
\hspace{1em}Proportional to current emissions & 0.582 & 0.040 & 0.041 & 0.516 & 0.647\\
\hspace{1em}Proportional to history of emissions & 0.515 & 0.039 & 0.695 & 0.451 & 0.580\\
\hspace{1em}Rich countries pay more than poor countries & 0.509 & 0.039 & 0.816 & 0.445 & 0.573\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}One year & 0.598 & 0.046 & 0.034 & 0.522 & 0.674\\
\hspace{1em}Five years & 0.674 & 0.040 & 0.000 & 0.608 & 0.740\\
\hspace{1em}Twenty years & 0.471 & 0.042 & 0.498 & 0.402 & 0.541\\
\hspace{1em}Fifty years & 0.381 & 0.042 & 0.004 & 0.312 & 0.450\\
\hspace{1em}Never & 0.380 & 0.044 & 0.007 & 0.308 & 0.453\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Ecuador}
\NormalTok{ecu <-}\StringTok{ }\KeywordTok{mm}\NormalTok{(}\KeywordTok{subset}\NormalTok{(cj, countryOrigin}\OperatorTok{==}\StringTok{'Ecuador'}\NormalTok{),  }
\NormalTok{        fm, }\DataTypeTok{id =} \OperatorTok{~}\NormalTok{Response.ID, }\DataTypeTok{alpha =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{h0 =} \FloatTok{0.5}\NormalTok{)}
\KeywordTok{table_mm}\NormalTok{(ecu, }\DataTypeTok{capt =} \StringTok{'Marginal Means -- Ecuador Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-6}Marginal Means -- Ecuador Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrrrrr}
\toprule
Feature & Estimate & Std.Error & P-Value & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}NGOs & 0.427 & 0.038 & 0.053 & 0.365 & 0.489\\
\hspace{1em}Local community members & 0.548 & 0.038 & 0.215 & 0.484 & 0.611\\
\hspace{1em}Local government & 0.518 & 0.039 & 0.643 & 0.454 & 0.581\\
\hspace{1em}Federal government & 0.475 & 0.039 & 0.529 & 0.411 & 0.540\\
\hspace{1em}International organizations & 0.531 & 0.038 & 0.405 & 0.469 & 0.593\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}Informal norms & 0.420 & 0.039 & 0.039 & 0.356 & 0.484\\
\hspace{1em}Private arbitration & 0.500 & 0.036 & 1.000 & 0.440 & 0.560\\
\hspace{1em}Local courts & 0.556 & 0.037 & 0.131 & 0.495 & 0.617\\
\hspace{1em}Government bureaucracy & 0.424 & 0.039 & 0.053 & 0.359 & 0.489\\
\hspace{1em}United Nations & 0.596 & 0.039 & 0.014 & 0.532 & 0.661\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}None & 0.490 & 0.040 & 0.810 & 0.424 & 0.556\\
\hspace{1em}Blacklist & 0.572 & 0.033 & 0.030 & 0.517 & 0.627\\
\hspace{1em}Fines & 0.544 & 0.033 & 0.184 & 0.490 & 0.598\\
\hspace{1em}Imprisonment & 0.397 & 0.032 & 0.001 & 0.345 & 0.450\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}Less penalty & 0.468 & 0.032 & 0.329 & 0.415 & 0.522\\
\hspace{1em}Same & 0.480 & 0.029 & 0.490 & 0.433 & 0.527\\
\hspace{1em}More penalty & 0.544 & 0.029 & 0.121 & 0.497 & 0.591\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}Only rich countries pay & 0.412 & 0.033 & 0.007 & 0.359 & 0.466\\
\hspace{1em}Proportional to current emissions & 0.581 & 0.034 & 0.016 & 0.526 & 0.637\\
\hspace{1em}Proportional to history of emissions & 0.546 & 0.036 & 0.197 & 0.487 & 0.604\\
\hspace{1em}Rich countries pay more than poor countries & 0.468 & 0.035 & 0.363 & 0.411 & 0.526\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}One year & 0.465 & 0.040 & 0.379 & 0.399 & 0.530\\
\hspace{1em}Five years & 0.554 & 0.038 & 0.149 & 0.492 & 0.616\\
\hspace{1em}Twenty years & 0.531 & 0.038 & 0.405 & 0.469 & 0.593\\
\hspace{1em}Fifty years & 0.447 & 0.038 & 0.165 & 0.384 & 0.510\\
\hspace{1em}Never & 0.497 & 0.039 & 0.938 & 0.433 & 0.561\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Mexico}
\NormalTok{mex <-}\StringTok{ }\KeywordTok{mm}\NormalTok{(}\KeywordTok{subset}\NormalTok{(cj, countryOrigin}\OperatorTok{==}\StringTok{'Mexico'}\NormalTok{),  }
\NormalTok{        fm, }\DataTypeTok{id =} \OperatorTok{~}\NormalTok{Response.ID, }\DataTypeTok{alpha =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{h0 =} \FloatTok{0.5}\NormalTok{)}
\KeywordTok{table_mm}\NormalTok{(mex, }\DataTypeTok{capt =} \StringTok{'Marginal Means -- Mexico Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-7}Marginal Means -- Mexico Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrrrrr}
\toprule
Feature & Estimate & Std.Error & P-Value & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}NGOs & 0.457 & 0.042 & 0.309 & 0.388 & 0.526\\
\hspace{1em}Local community members & 0.470 & 0.043 & 0.489 & 0.399 & 0.541\\
\hspace{1em}Local government & 0.438 & 0.040 & 0.122 & 0.372 & 0.504\\
\hspace{1em}Federal government & 0.497 & 0.040 & 0.937 & 0.432 & 0.562\\
\hspace{1em}International organizations & 0.620 & 0.038 & 0.001 & 0.559 & 0.682\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}Informal norms & 0.405 & 0.040 & 0.019 & 0.339 & 0.472\\
\hspace{1em}Private arbitration & 0.520 & 0.041 & 0.626 & 0.453 & 0.586\\
\hspace{1em}Local courts & 0.507 & 0.042 & 0.867 & 0.438 & 0.576\\
\hspace{1em}Government bureaucracy & 0.461 & 0.040 & 0.329 & 0.394 & 0.527\\
\hspace{1em}United Nations & 0.601 & 0.039 & 0.009 & 0.537 & 0.665\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}None & 0.472 & 0.044 & 0.534 & 0.400 & 0.545\\
\hspace{1em}Blacklist & 0.478 & 0.035 & 0.531 & 0.421 & 0.535\\
\hspace{1em}Fines & 0.571 & 0.035 & 0.040 & 0.514 & 0.629\\
\hspace{1em}Imprisonment & 0.470 & 0.034 & 0.374 & 0.414 & 0.526\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}Less penalty & 0.505 & 0.035 & 0.888 & 0.447 & 0.563\\
\hspace{1em}Same & 0.503 & 0.029 & 0.908 & 0.456 & 0.551\\
\hspace{1em}More penalty & 0.492 & 0.031 & 0.802 & 0.441 & 0.544\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}Only rich countries pay & 0.429 & 0.036 & 0.047 & 0.369 & 0.488\\
\hspace{1em}Proportional to current emissions & 0.500 & 0.035 & 1.000 & 0.442 & 0.558\\
\hspace{1em}Proportional to history of emissions & 0.620 & 0.038 & 0.001 & 0.559 & 0.682\\
\hspace{1em}Rich countries pay more than poor countries & 0.466 & 0.036 & 0.348 & 0.407 & 0.525\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}One year & 0.507 & 0.041 & 0.871 & 0.440 & 0.573\\
\hspace{1em}Five years & 0.616 & 0.038 & 0.002 & 0.553 & 0.678\\
\hspace{1em}Twenty years & 0.487 & 0.041 & 0.746 & 0.420 & 0.554\\
\hspace{1em}Fifty years & 0.493 & 0.043 & 0.864 & 0.422 & 0.563\\
\hspace{1em}Never & 0.385 & 0.040 & 0.004 & 0.319 & 0.451\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Panama}
\NormalTok{pan <-}\StringTok{ }\KeywordTok{mm}\NormalTok{(}\KeywordTok{subset}\NormalTok{(cj, countryOrigin}\OperatorTok{==}\StringTok{'Panama'}\NormalTok{),  }
\NormalTok{        fm, }\DataTypeTok{id =} \OperatorTok{~}\NormalTok{Response.ID, }\DataTypeTok{alpha =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{h0 =} \FloatTok{0.5}\NormalTok{)}
\KeywordTok{table_mm}\NormalTok{(pan, }\DataTypeTok{capt =} \StringTok{'Marginal Means -- Panama Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-8}Marginal Means -- Panama Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrrrrr}
\toprule
Feature & Estimate & Std.Error & P-Value & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}NGOs & 0.478 & 0.040 & 0.578 & 0.413 & 0.543\\
\hspace{1em}Local community members & 0.526 & 0.040 & 0.521 & 0.460 & 0.591\\
\hspace{1em}Local government & 0.477 & 0.041 & 0.566 & 0.409 & 0.544\\
\hspace{1em}Federal government & 0.497 & 0.040 & 0.936 & 0.431 & 0.563\\
\hspace{1em}International organizations & 0.522 & 0.040 & 0.576 & 0.457 & 0.588\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}Informal norms & 0.439 & 0.040 & 0.127 & 0.374 & 0.505\\
\hspace{1em}Private arbitration & 0.488 & 0.039 & 0.756 & 0.424 & 0.552\\
\hspace{1em}Local courts & 0.549 & 0.041 & 0.241 & 0.480 & 0.617\\
\hspace{1em}Government bureaucracy & 0.477 & 0.040 & 0.571 & 0.411 & 0.544\\
\hspace{1em}United Nations & 0.551 & 0.040 & 0.198 & 0.486 & 0.617\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}None & 0.455 & 0.042 & 0.275 & 0.386 & 0.523\\
\hspace{1em}Blacklist & 0.522 & 0.035 & 0.531 & 0.465 & 0.579\\
\hspace{1em}Fines & 0.531 & 0.033 & 0.349 & 0.476 & 0.586\\
\hspace{1em}Imprisonment & 0.475 & 0.035 & 0.481 & 0.417 & 0.533\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}Less penalty & 0.432 & 0.033 & 0.042 & 0.378 & 0.487\\
\hspace{1em}Same & 0.466 & 0.030 & 0.253 & 0.416 & 0.515\\
\hspace{1em}More penalty & 0.588 & 0.030 & 0.003 & 0.540 & 0.637\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}Only rich countries pay & 0.473 & 0.035 & 0.442 & 0.416 & 0.531\\
\hspace{1em}Proportional to current emissions & 0.497 & 0.036 & 0.942 & 0.438 & 0.557\\
\hspace{1em}Proportional to history of emissions & 0.513 & 0.037 & 0.715 & 0.453 & 0.573\\
\hspace{1em}Rich countries pay more than poor countries & 0.518 & 0.036 & 0.614 & 0.459 & 0.577\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}One year & 0.605 & 0.040 & 0.008 & 0.540 & 0.670\\
\hspace{1em}Five years & 0.552 & 0.038 & 0.170 & 0.490 & 0.614\\
\hspace{1em}Twenty years & 0.573 & 0.040 & 0.069 & 0.507 & 0.640\\
\hspace{1em}Fifty years & 0.418 & 0.041 & 0.044 & 0.351 & 0.485\\
\hspace{1em}Never & 0.344 & 0.038 & 0.000 & 0.281 & 0.407\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Peru}
\NormalTok{per <-}\StringTok{ }\KeywordTok{mm}\NormalTok{(}\KeywordTok{subset}\NormalTok{(cj, countryOrigin}\OperatorTok{==}\StringTok{'Peru'}\NormalTok{),  }
\NormalTok{        fm, }\DataTypeTok{id =} \OperatorTok{~}\NormalTok{Response.ID, }\DataTypeTok{alpha =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{h0 =} \FloatTok{0.5}\NormalTok{)}
\KeywordTok{table_mm}\NormalTok{(per, }\DataTypeTok{capt =} \StringTok{'Marginal Means -- Peru Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-9}Marginal Means -- Peru Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrrrrr}
\toprule
Feature & Estimate & Std.Error & P-Value & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}NGOs & 0.442 & 0.049 & 0.236 & 0.362 & 0.522\\
\hspace{1em}Local community members & 0.410 & 0.048 & 0.059 & 0.331 & 0.488\\
\hspace{1em}Local government & 0.568 & 0.047 & 0.151 & 0.490 & 0.645\\
\hspace{1em}Federal government & 0.537 & 0.048 & 0.440 & 0.458 & 0.616\\
\hspace{1em}International organizations & 0.533 & 0.046 & 0.464 & 0.458 & 0.608\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}Informal norms & 0.351 & 0.048 & 0.002 & 0.271 & 0.430\\
\hspace{1em}Private arbitration & 0.607 & 0.045 & 0.018 & 0.533 & 0.681\\
\hspace{1em}Local courts & 0.505 & 0.051 & 0.919 & 0.422 & 0.589\\
\hspace{1em}Government bureaucracy & 0.488 & 0.045 & 0.788 & 0.414 & 0.562\\
\hspace{1em}United Nations & 0.527 & 0.047 & 0.570 & 0.449 & 0.604\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}None & 0.441 & 0.047 & 0.214 & 0.364 & 0.519\\
\hspace{1em}Blacklist & 0.500 & 0.043 & 1.000 & 0.430 & 0.570\\
\hspace{1em}Fines & 0.564 & 0.042 & 0.125 & 0.495 & 0.633\\
\hspace{1em}Imprisonment & 0.484 & 0.040 & 0.692 & 0.419 & 0.549\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}Less penalty & 0.472 & 0.042 & 0.501 & 0.403 & 0.541\\
\hspace{1em}Same & 0.536 & 0.035 & 0.298 & 0.479 & 0.593\\
\hspace{1em}More penalty & 0.482 & 0.036 & 0.618 & 0.424 & 0.541\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}Only rich countries pay & 0.468 & 0.040 & 0.419 & 0.401 & 0.534\\
\hspace{1em}Proportional to current emissions & 0.497 & 0.040 & 0.936 & 0.430 & 0.563\\
\hspace{1em}Proportional to history of emissions & 0.571 & 0.047 & 0.127 & 0.495 & 0.648\\
\hspace{1em}Rich countries pay more than poor countries & 0.481 & 0.044 & 0.660 & 0.408 & 0.553\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}One year & 0.548 & 0.045 & 0.279 & 0.475 & 0.622\\
\hspace{1em}Five years & 0.583 & 0.047 & 0.079 & 0.505 & 0.661\\
\hspace{1em}Twenty years & 0.517 & 0.054 & 0.748 & 0.429 & 0.605\\
\hspace{1em}Fifty years & 0.420 & 0.047 & 0.085 & 0.343 & 0.496\\
\hspace{1em}Never & 0.436 & 0.046 & 0.162 & 0.360 & 0.511\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Changing country labels}
\NormalTok{arg}\OperatorTok{$}\NormalTok{country <-}\StringTok{ 'Argentina'}
\NormalTok{bol}\OperatorTok{$}\NormalTok{country <-}\StringTok{ 'Bolivia'}
\NormalTok{bra}\OperatorTok{$}\NormalTok{country <-}\StringTok{ 'Brazil'}
\NormalTok{chi}\OperatorTok{$}\NormalTok{country <-}\StringTok{ 'Chile'}
\NormalTok{col}\OperatorTok{$}\NormalTok{country <-}\StringTok{ 'Colombia'}
\NormalTok{cri}\OperatorTok{$}\NormalTok{country <-}\StringTok{ 'Costa Rica'}
\NormalTok{ecu}\OperatorTok{$}\NormalTok{country <-}\StringTok{ 'Ecuador'}
\NormalTok{mex}\OperatorTok{$}\NormalTok{country <-}\StringTok{ 'Mexico'}
\NormalTok{pan}\OperatorTok{$}\NormalTok{country <-}\StringTok{ 'Panama'}
\NormalTok{per}\OperatorTok{$}\NormalTok{country <-}\StringTok{ 'Peru'}
\CommentTok{# Plot by country}
\NormalTok{p <-}\StringTok{ }\KeywordTok{plot}\NormalTok{(}\KeywordTok{rbind}\NormalTok{(arg, bol, bra, chi, col, }
\NormalTok{              cri, ecu, mex, pan, per), }
        \DataTypeTok{group =} \StringTok{'country'}\NormalTok{, }\DataTypeTok{vline =} \FloatTok{0.5}\NormalTok{, }\DataTypeTok{nr =} \DecValTok{10}\NormalTok{,}
        \DataTypeTok{header_fmt =} \StringTok{"%s"}\NormalTok{, }\DataTypeTok{size =} \DecValTok{2}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{( }\OperatorTok{~}\StringTok{ }\NormalTok{country, }\DataTypeTok{ncol =} \DecValTok{5}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\NormalTok{ggplot2}\OperatorTok{::}\KeywordTok{theme}\NormalTok{(}
    \DataTypeTok{axis.text.y =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{face=}\NormalTok{myFaces, }\DataTypeTok{size =} \DecValTok{11}\NormalTok{)) }\OperatorTok{+}\StringTok{ }
\StringTok{    }\NormalTok{ggplot2}\OperatorTok{::}\KeywordTok{geom_errorbarh}\NormalTok{(ggplot2}\OperatorTok{::}\KeywordTok{aes_string}\NormalTok{(}\DataTypeTok{xmin =} \StringTok{"lower"}\NormalTok{,}
                                                \DataTypeTok{xmax =} \StringTok{"upper"}\NormalTok{),}
                            \DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{height =} \DecValTok{0}\NormalTok{, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{,}
                            \DataTypeTok{position =}\NormalTok{ ggstance}\OperatorTok{::}\KeywordTok{position_dodgev}\NormalTok{(}\DataTypeTok{height =} \DecValTok{1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: Vectorized input to `element_text()` is not officially supported.
## Results may be unexpected or may change in future versions of ggplot2.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{p}
\end{Highlighting}
\end{Shaded}

\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/unnamed-chunk-10-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggsave}\NormalTok{(}\DataTypeTok{filename =} \StringTok{'MM_country.pdf'}\NormalTok{, }
       \DataTypeTok{plot =}\NormalTok{ p, }\DataTypeTok{width =} \DecValTok{12}\NormalTok{, }\DataTypeTok{height =} \DecValTok{12}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The code for Figure 4 of the paper is available below. Tables displaying
the point estimates are also included.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## Elite type graphs and estimates}

\CommentTok{# Table for the Executive branch}
\NormalTok{res1 <-}\StringTok{ }\KeywordTok{mm}\NormalTok{(}\KeywordTok{subset}\NormalTok{(cj, groupOrigin}\OperatorTok{==}\StringTok{'Executive'}\NormalTok{), }
\NormalTok{           fm, }\DataTypeTok{id =} \OperatorTok{~}\NormalTok{Response.ID, }\DataTypeTok{alpha =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{h0 =} \FloatTok{0.5}\NormalTok{)}
\KeywordTok{table_mm}\NormalTok{(res1, }\DataTypeTok{capt =} \StringTok{'Marginal Means -- Executive Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-11}Marginal Means -- Executive Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrrrrr}
\toprule
Feature & Estimate & Std.Error & P-Value & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}NGOs & 0.392 & 0.034 & 0.001 & 0.336 & 0.447\\
\hspace{1em}Local community members & 0.516 & 0.031 & 0.618 & 0.464 & 0.567\\
\hspace{1em}Local government & 0.502 & 0.030 & 0.952 & 0.452 & 0.552\\
\hspace{1em}Federal government & 0.529 & 0.032 & 0.365 & 0.476 & 0.582\\
\hspace{1em}International organizations & 0.545 & 0.031 & 0.150 & 0.494 & 0.596\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}Informal norms & 0.429 & 0.034 & 0.037 & 0.373 & 0.485\\
\hspace{1em}Private arbitration & 0.508 & 0.032 & 0.797 & 0.455 & 0.561\\
\hspace{1em}Local courts & 0.523 & 0.031 & 0.456 & 0.472 & 0.574\\
\hspace{1em}Government bureaucracy & 0.462 & 0.031 & 0.214 & 0.411 & 0.512\\
\hspace{1em}United Nations & 0.564 & 0.030 & 0.036 & 0.514 & 0.614\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}None & 0.480 & 0.034 & 0.545 & 0.424 & 0.535\\
\hspace{1em}Blacklist & 0.541 & 0.028 & 0.144 & 0.495 & 0.587\\
\hspace{1em}Fines & 0.568 & 0.026 & 0.010 & 0.524 & 0.611\\
\hspace{1em}Imprisonment & 0.406 & 0.026 & 0.000 & 0.363 & 0.450\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}Less penalty & 0.447 & 0.026 & 0.039 & 0.405 & 0.489\\
\hspace{1em}Same & 0.501 & 0.023 & 0.963 & 0.463 & 0.539\\
\hspace{1em}More penalty & 0.549 & 0.025 & 0.050 & 0.508 & 0.590\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}Only rich countries pay & 0.442 & 0.029 & 0.043 & 0.395 & 0.489\\
\hspace{1em}Proportional to current emissions & 0.550 & 0.028 & 0.072 & 0.504 & 0.596\\
\hspace{1em}Proportional to history of emissions & 0.524 & 0.028 & 0.396 & 0.478 & 0.570\\
\hspace{1em}Rich countries pay more than poor countries & 0.480 & 0.029 & 0.491 & 0.433 & 0.527\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}One year & 0.543 & 0.030 & 0.159 & 0.493 & 0.593\\
\hspace{1em}Five years & 0.586 & 0.030 & 0.005 & 0.536 & 0.636\\
\hspace{1em}Twenty years & 0.500 & 0.032 & 1.000 & 0.448 & 0.552\\
\hspace{1em}Fifty years & 0.438 & 0.032 & 0.052 & 0.386 & 0.490\\
\hspace{1em}Never & 0.414 & 0.033 & 0.009 & 0.359 & 0.468\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Legislative branch}
\NormalTok{res2 <-}\StringTok{ }\KeywordTok{mm}\NormalTok{(}\KeywordTok{subset}\NormalTok{(cj, groupOrigin}\OperatorTok{==}\StringTok{'Legislative'}\NormalTok{), }
\NormalTok{           fm, }\DataTypeTok{id =} \OperatorTok{~}\NormalTok{Response.ID, }\DataTypeTok{alpha =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{h0 =} \FloatTok{0.5}\NormalTok{)}
\KeywordTok{table_mm}\NormalTok{(res2, }\DataTypeTok{capt =} \StringTok{'Marginal Means -- Legislative Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-12}Marginal Means -- Legislative Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrrrrr}
\toprule
Feature & Estimate & Std.Error & P-Value & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}NGOs & 0.400 & 0.038 & 0.008 & 0.338 & 0.462\\
\hspace{1em}Local community members & 0.519 & 0.039 & 0.637 & 0.454 & 0.583\\
\hspace{1em}Local government & 0.549 & 0.039 & 0.209 & 0.485 & 0.613\\
\hspace{1em}Federal government & 0.517 & 0.041 & 0.680 & 0.449 & 0.585\\
\hspace{1em}International organizations & 0.524 & 0.042 & 0.558 & 0.456 & 0.593\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}Informal norms & 0.420 & 0.037 & 0.032 & 0.358 & 0.481\\
\hspace{1em}Private arbitration & 0.506 & 0.039 & 0.876 & 0.442 & 0.570\\
\hspace{1em}Local courts & 0.547 & 0.039 & 0.235 & 0.482 & 0.611\\
\hspace{1em}Government bureaucracy & 0.500 & 0.041 & 1.000 & 0.432 & 0.568\\
\hspace{1em}United Nations & 0.539 & 0.042 & 0.353 & 0.470 & 0.608\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}None & 0.438 & 0.042 & 0.143 & 0.368 & 0.508\\
\hspace{1em}Blacklist & 0.546 & 0.034 & 0.172 & 0.491 & 0.602\\
\hspace{1em}Fines & 0.581 & 0.033 & 0.013 & 0.528 & 0.635\\
\hspace{1em}Imprisonment & 0.403 & 0.034 & 0.005 & 0.347 & 0.459\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}Less penalty & 0.455 & 0.033 & 0.178 & 0.400 & 0.510\\
\hspace{1em}Same & 0.521 & 0.029 & 0.481 & 0.472 & 0.569\\
\hspace{1em}More penalty & 0.515 & 0.030 & 0.629 & 0.465 & 0.564\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}Only rich countries pay & 0.411 & 0.033 & 0.007 & 0.357 & 0.465\\
\hspace{1em}Proportional to current emissions & 0.475 & 0.035 & 0.481 & 0.417 & 0.533\\
\hspace{1em}Proportional to history of emissions & 0.552 & 0.039 & 0.181 & 0.488 & 0.616\\
\hspace{1em}Rich countries pay more than poor countries & 0.584 & 0.035 & 0.017 & 0.526 & 0.642\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}One year & 0.533 & 0.040 & 0.416 & 0.466 & 0.599\\
\hspace{1em}Five years & 0.557 & 0.038 & 0.139 & 0.494 & 0.620\\
\hspace{1em}Twenty years & 0.526 & 0.040 & 0.519 & 0.460 & 0.592\\
\hspace{1em}Fifty years & 0.485 & 0.039 & 0.695 & 0.420 & 0.549\\
\hspace{1em}Never & 0.393 & 0.040 & 0.007 & 0.328 & 0.459\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Civil society}
\NormalTok{res3 <-}\StringTok{ }\KeywordTok{mm}\NormalTok{(}\KeywordTok{subset}\NormalTok{(cj, groupOrigin}\OperatorTok{==}\StringTok{'Civil Society'}\NormalTok{), }
\NormalTok{           fm, }\DataTypeTok{id =} \OperatorTok{~}\NormalTok{Response.ID, }\DataTypeTok{alpha =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{h0 =} \FloatTok{0.5}\NormalTok{)}
\KeywordTok{table_mm}\NormalTok{(res3, }\DataTypeTok{capt =} \StringTok{'Marginal Means -- Civil Society Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-13}Marginal Means -- Civil Society Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrrrrr}
\toprule
Feature & Estimate & Std.Error & P-Value & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}NGOs & 0.500 & 0.023 & 1.000 & 0.463 & 0.537\\
\hspace{1em}Local community members & 0.520 & 0.022 & 0.368 & 0.483 & 0.557\\
\hspace{1em}Local government & 0.462 & 0.023 & 0.093 & 0.425 & 0.499\\
\hspace{1em}Federal government & 0.478 & 0.022 & 0.309 & 0.441 & 0.514\\
\hspace{1em}International organizations & 0.538 & 0.022 & 0.080 & 0.502 & 0.574\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}Informal norms & 0.410 & 0.022 & 0.000 & 0.373 & 0.446\\
\hspace{1em}Private arbitration & 0.503 & 0.023 & 0.892 & 0.466 & 0.540\\
\hspace{1em}Local courts & 0.551 & 0.022 & 0.021 & 0.515 & 0.587\\
\hspace{1em}Government bureaucracy & 0.430 & 0.022 & 0.002 & 0.393 & 0.467\\
\hspace{1em}United Nations & 0.594 & 0.021 & 0.000 & 0.559 & 0.629\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}None & 0.453 & 0.023 & 0.036 & 0.416 & 0.490\\
\hspace{1em}Blacklist & 0.518 & 0.020 & 0.364 & 0.485 & 0.550\\
\hspace{1em}Fines & 0.548 & 0.019 & 0.011 & 0.517 & 0.579\\
\hspace{1em}Imprisonment & 0.467 & 0.019 & 0.089 & 0.436 & 0.499\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}Less penalty & 0.469 & 0.019 & 0.112 & 0.437 & 0.501\\
\hspace{1em}Same & 0.479 & 0.016 & 0.203 & 0.452 & 0.506\\
\hspace{1em}More penalty & 0.545 & 0.017 & 0.007 & 0.517 & 0.572\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}Only rich countries pay & 0.449 & 0.019 & 0.008 & 0.417 & 0.481\\
\hspace{1em}Proportional to current emissions & 0.520 & 0.020 & 0.315 & 0.487 & 0.553\\
\hspace{1em}Proportional to history of emissions & 0.560 & 0.020 & 0.003 & 0.527 & 0.593\\
\hspace{1em}Rich countries pay more than poor countries & 0.476 & 0.020 & 0.230 & 0.443 & 0.509\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}One year & 0.548 & 0.022 & 0.026 & 0.513 & 0.584\\
\hspace{1em}Five years & 0.595 & 0.022 & 0.000 & 0.559 & 0.630\\
\hspace{1em}Twenty years & 0.493 & 0.022 & 0.757 & 0.457 & 0.530\\
\hspace{1em}Fifty years & 0.448 & 0.023 & 0.026 & 0.409 & 0.486\\
\hspace{1em}Never & 0.404 & 0.022 & 0.000 & 0.368 & 0.441\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Academia}
\NormalTok{res4 <-}\StringTok{ }\KeywordTok{mm}\NormalTok{(}\KeywordTok{subset}\NormalTok{(cj, groupOrigin}\OperatorTok{==}\StringTok{'Academia'}\NormalTok{), }
\NormalTok{           fm, }\DataTypeTok{id =} \OperatorTok{~}\NormalTok{Response.ID, }\DataTypeTok{alpha =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{h0 =} \FloatTok{0.5}\NormalTok{)}
\KeywordTok{table_mm}\NormalTok{(res4, }\DataTypeTok{capt =} \StringTok{'Marginal Means -- Academia Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-14}Marginal Means -- Academia Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lrrrrr}
\toprule
Feature & Estimate & Std.Error & P-Value & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}NGOs & 0.447 & 0.022 & 0.014 & 0.411 & 0.483\\
\hspace{1em}Local community members & 0.515 & 0.022 & 0.483 & 0.479 & 0.551\\
\hspace{1em}Local government & 0.524 & 0.022 & 0.260 & 0.489 & 0.560\\
\hspace{1em}Federal government & 0.458 & 0.021 & 0.048 & 0.423 & 0.493\\
\hspace{1em}International organizations & 0.556 & 0.021 & 0.009 & 0.521 & 0.591\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}Informal norms & 0.388 & 0.021 & 0.000 & 0.353 & 0.423\\
\hspace{1em}Private arbitration & 0.517 & 0.022 & 0.437 & 0.481 & 0.552\\
\hspace{1em}Local courts & 0.549 & 0.022 & 0.023 & 0.513 & 0.584\\
\hspace{1em}Government bureaucracy & 0.476 & 0.022 & 0.275 & 0.440 & 0.512\\
\hspace{1em}United Nations & 0.565 & 0.021 & 0.002 & 0.530 & 0.599\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}None & 0.430 & 0.022 & 0.002 & 0.393 & 0.467\\
\hspace{1em}Blacklist & 0.510 & 0.018 & 0.582 & 0.480 & 0.540\\
\hspace{1em}Fines & 0.554 & 0.019 & 0.004 & 0.523 & 0.585\\
\hspace{1em}Imprisonment & 0.483 & 0.019 & 0.371 & 0.453 & 0.514\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}Less penalty & 0.461 & 0.018 & 0.030 & 0.431 & 0.490\\
\hspace{1em}Same & 0.499 & 0.016 & 0.974 & 0.473 & 0.526\\
\hspace{1em}More penalty & 0.532 & 0.016 & 0.048 & 0.505 & 0.559\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}Only rich countries pay & 0.434 & 0.019 & 0.001 & 0.403 & 0.466\\
\hspace{1em}Proportional to current emissions & 0.540 & 0.019 & 0.036 & 0.509 & 0.571\\
\hspace{1em}Proportional to history of emissions & 0.519 & 0.020 & 0.327 & 0.487 & 0.551\\
\hspace{1em}Rich countries pay more than poor countries & 0.507 & 0.020 & 0.723 & 0.475 & 0.539\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}One year & 0.509 & 0.022 & 0.664 & 0.474 & 0.545\\
\hspace{1em}Five years & 0.595 & 0.021 & 0.000 & 0.560 & 0.629\\
\hspace{1em}Twenty years & 0.549 & 0.022 & 0.023 & 0.514 & 0.585\\
\hspace{1em}Fifty years & 0.422 & 0.021 & 0.000 & 0.386 & 0.457\\
\hspace{1em}Never & 0.421 & 0.022 & 0.000 & 0.386 & 0.456\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Changing labels}
\NormalTok{res1}\OperatorTok{$}\NormalTok{MemberType <-}\StringTok{ 'Executive'}
\NormalTok{res2}\OperatorTok{$}\NormalTok{MemberType <-}\StringTok{ 'Legislative'}
\NormalTok{res3}\OperatorTok{$}\NormalTok{MemberType <-}\StringTok{ 'Civil Society'}
\NormalTok{res4}\OperatorTok{$}\NormalTok{MemberType <-}\StringTok{ 'Academia'}

\CommentTok{# Plot by Elite Type}
\NormalTok{p <-}\StringTok{ }\KeywordTok{plot}\NormalTok{(}\KeywordTok{rbind}\NormalTok{(res1, res2, res3, res4), }
     \DataTypeTok{group =} \StringTok{'MemberType'}\NormalTok{, }\DataTypeTok{vline =} \FloatTok{0.5}\NormalTok{, }\DataTypeTok{nr =} \DecValTok{4}\NormalTok{,}
        \DataTypeTok{header_fmt =} \StringTok{"%s"}\NormalTok{, }\DataTypeTok{size =} \DecValTok{2}\NormalTok{) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{( }\OperatorTok{~}\StringTok{ }\NormalTok{MemberType, }\DataTypeTok{ncol =} \DecValTok{4}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\NormalTok{ggplot2}\OperatorTok{::}\KeywordTok{theme}\NormalTok{(}
    \DataTypeTok{axis.text.y =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{face=}\NormalTok{myFaces, }\DataTypeTok{size =} \DecValTok{11}\NormalTok{)) }\OperatorTok{+}\StringTok{ }
\StringTok{    }\NormalTok{ggplot2}\OperatorTok{::}\KeywordTok{geom_errorbarh}\NormalTok{(ggplot2}\OperatorTok{::}\KeywordTok{aes_string}\NormalTok{(}\DataTypeTok{xmin =} \StringTok{"lower"}\NormalTok{,}
                                                \DataTypeTok{xmax =} \StringTok{"upper"}\NormalTok{),}
                            \DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{height =} \DecValTok{0}\NormalTok{, }\DataTypeTok{na.rm =} \OtherTok{TRUE}\NormalTok{,}
                            \DataTypeTok{position =}\NormalTok{ ggstance}\OperatorTok{::}\KeywordTok{position_dodgev}\NormalTok{(}\DataTypeTok{height =} \DecValTok{1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: Vectorized input to `element_text()` is not officially supported.
## Results may be unexpected or may change in future versions of ggplot2.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{p}
\end{Highlighting}
\end{Shaded}

\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/unnamed-chunk-15-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggsave}\NormalTok{(}\DataTypeTok{filename =} \StringTok{'MM_membertype.pdf'}\NormalTok{, }\DataTypeTok{plot =}\NormalTok{ p, }
       \DataTypeTok{width =} \DecValTok{8}\NormalTok{, }\DataTypeTok{height =} \DecValTok{6}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{average-marginal-component-effect-amce-estimator}{%
\section{\texorpdfstring{Average Marginal Component Effect (AMCE)
Estimator
\label{sec:amce}}{Average Marginal Component Effect (AMCE) Estimator }}\label{average-marginal-component-effect-amce-estimator}}

This estimator fixes one category and look at changes from this baseline
category. Below follows the plots for the main model and the country and
elite type subsamples.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## Main model}
\KeywordTok{rm}\NormalTok{(cj)}
\KeywordTok{load}\NormalTok{(}\StringTok{'freire-mignozzetti-skarbek.RData'}\NormalTok{)}

\CommentTok{# Set conjoint baselines}
\NormalTok{baselines <-}\StringTok{ }\KeywordTok{list}\NormalTok{()}
\NormalTok{baselines}\OperatorTok{$}\StringTok{`}\DataTypeTok{How often will the agreement be renegotiated?}\StringTok{`}\NormalTok{ <-}\StringTok{ "One year"}
\NormalTok{baselines}\OperatorTok{$}\StringTok{`}\DataTypeTok{What punishments do they use?}\StringTok{`}\NormalTok{ <-}\StringTok{ 'None'}
\NormalTok{baselines}\OperatorTok{$}\StringTok{`}\DataTypeTok{How are conflicts resolved?}\StringTok{`}\NormalTok{ <-}\StringTok{ 'Government bureaucracy'}
\NormalTok{baselines}\OperatorTok{$}\StringTok{`}\DataTypeTok{How are costs distributed?}\StringTok{`}\NormalTok{ <-}\StringTok{ 'Rich countries pay more than poor countries'}
\NormalTok{baselines}\OperatorTok{$}\StringTok{`}\DataTypeTok{Who makes the rules?}\StringTok{`}\NormalTok{ <-}\StringTok{ 'Federal government'}

\NormalTok{attrs <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Who makes the rules?"}\NormalTok{,}
           \StringTok{"How are conflicts resolved?"}\NormalTok{,}
           \StringTok{"What punishments do they use?"}\NormalTok{,}
           \StringTok{"How are repeated violations punished?"}\NormalTok{,}
           \StringTok{"How are costs distributed?"}\NormalTok{,}
           \StringTok{"How often will the agreement be renegotiated?"}\NormalTok{)}

\NormalTok{results <-}\StringTok{ }\NormalTok{cjoint}\OperatorTok{::}\KeywordTok{amce}\NormalTok{(fm, }\DataTypeTok{data =}\NormalTok{ cj, }\DataTypeTok{cluster =} \OtherTok{TRUE}\NormalTok{, }
                        \DataTypeTok{respondent.id =} \StringTok{"Response.ID"}\NormalTok{, }
                        \DataTypeTok{design =}\NormalTok{ conjDesign, }
                        \DataTypeTok{baselines =}\NormalTok{ baselines, }\DataTypeTok{na.ignore =}\NormalTok{ T)}

\CommentTok{# Table AMCE -- Full Model }
\NormalTok{tableAMCE <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(results, }\DataTypeTok{capt =} \StringTok{'Main Model'}\NormalTok{) \{}
\NormalTok{  aux <-}\StringTok{ }\KeywordTok{plotAMCE}\NormalTok{(results, }\DataTypeTok{ci =} \FloatTok{0.9}\NormalTok{, }\DataTypeTok{point.size =} \FloatTok{.8}\NormalTok{, }\DataTypeTok{dodge.size =} \DecValTok{1}\NormalTok{,}
     \DataTypeTok{text.size =} \DecValTok{10}\NormalTok{, }\DataTypeTok{main =} \StringTok{'AMCE -- Full Model'}\NormalTok{,}
     \DataTypeTok{attribute.names =}\NormalTok{ attrs, }\DataTypeTok{tblfy =}\NormalTok{ T)}
\NormalTok{  aux <-}\StringTok{ }\NormalTok{aux }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(group)) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{select}\NormalTok{(printvar, pe, se, lower, upper)}
\NormalTok{  indx <-}\StringTok{ }\KeywordTok{grep}\NormalTok{(}\StringTok{'Baseline ='}\NormalTok{, aux}\OperatorTok{$}\NormalTok{printvar)}
\NormalTok{  aux}\OperatorTok{$}\NormalTok{pe <-}\StringTok{ }\KeywordTok{round}\NormalTok{(aux}\OperatorTok{$}\NormalTok{pe, }\DataTypeTok{digits =} \DecValTok{3}\NormalTok{)}
\NormalTok{  aux}\OperatorTok{$}\NormalTok{se <-}\StringTok{ }\KeywordTok{round}\NormalTok{(aux}\OperatorTok{$}\NormalTok{se, }\DataTypeTok{digits =} \DecValTok{3}\NormalTok{)}
\NormalTok{  aux}\OperatorTok{$}\NormalTok{lower <-}\StringTok{ }\KeywordTok{round}\NormalTok{(aux}\OperatorTok{$}\NormalTok{lower, }\DataTypeTok{digits =} \DecValTok{3}\NormalTok{)}
\NormalTok{  aux}\OperatorTok{$}\NormalTok{upper <-}\StringTok{ }\KeywordTok{round}\NormalTok{(aux}\OperatorTok{$}\NormalTok{upper, }\DataTypeTok{digits =} \DecValTok{3}\NormalTok{)}
\NormalTok{  aux}\OperatorTok{$}\NormalTok{pe[indx] =}\StringTok{ ''}
\NormalTok{  aux}\OperatorTok{$}\NormalTok{se[indx] =}\StringTok{ ''}
\NormalTok{  aux}\OperatorTok{$}\NormalTok{lower[indx] =}\StringTok{ ''}
\NormalTok{  aux}\OperatorTok{$}\NormalTok{upper[indx] =}\StringTok{ ''}
  \KeywordTok{names}\NormalTok{(aux) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{'Feature'}\NormalTok{, }\StringTok{'Estimate'}\NormalTok{, }\StringTok{'Std.Error'}\NormalTok{, }
                  \StringTok{'Lower'}\NormalTok{, }\StringTok{'Upper'}\NormalTok{)}
  \KeywordTok{return}\NormalTok{(}\KeywordTok{kable}\NormalTok{(aux, }\StringTok{"latex"}\NormalTok{, }\DataTypeTok{caption =}\NormalTok{ capt, }\DataTypeTok{booktabs =}\NormalTok{ T, }\DataTypeTok{align =} \KeywordTok{c}\NormalTok{(}\StringTok{'l'}\NormalTok{, }\KeywordTok{rep}\NormalTok{(}\StringTok{'c'}\NormalTok{, }\DecValTok{4}\NormalTok{))) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{kable_styling}\NormalTok{(}\DataTypeTok{font_size =} \DecValTok{10}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{group_rows}\NormalTok{(}\StringTok{'How are conflicts resolved?'}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{5}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{group_rows}\NormalTok{(}\StringTok{'How are costs distributed?'}\NormalTok{, }\DecValTok{6}\NormalTok{, }\DecValTok{9}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{group_rows}\NormalTok{(}\StringTok{'How are repeated violations punished?'}\NormalTok{, }\DecValTok{10}\NormalTok{, }\DecValTok{12}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{group_rows}\NormalTok{(}\StringTok{'How often will the agreement be renegotiated?'}\NormalTok{, }\DecValTok{13}\NormalTok{, }\DecValTok{17}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{group_rows}\NormalTok{(}\StringTok{'What punishments do they use?'}\NormalTok{, }\DecValTok{18}\NormalTok{, }\DecValTok{21}\NormalTok{) }\OperatorTok{%>%}
\StringTok{    }\KeywordTok{group_rows}\NormalTok{(}\StringTok{'Who makes the rules?'}\NormalTok{, }\DecValTok{22}\NormalTok{, }\DecValTok{26}\NormalTok{))}
\NormalTok{\}}

\KeywordTok{tableAMCE}\NormalTok{(results, }\DataTypeTok{capt =} \StringTok{'AMCE -- Full Model'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-16}AMCE -- Full Model}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lcccc}
\toprule
Feature & Estimate & Std.Error & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}(Baseline = Government bureaucracy) &  &  &  & \\
\hspace{1em}Informal norms & -0.054 & 0.018 & -0.083 & -0.025\\
\hspace{1em}Private arbitration & 0.043 & 0.018 & 0.014 & 0.072\\
\hspace{1em}Local courts & 0.088 & 0.017 & 0.059 & 0.116\\
\hspace{1em}United Nations & 0.111 & 0.019 & 0.081 & 0.142\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}(Baseline = Rich countries pay more than poor countries) &  &  &  & \\
\hspace{1em}Only rich countries pay & -0.071 & 0.016 & -0.098 & -0.044\\
\hspace{1em}Proportional to current emissions & 0.018 & 0.016 & -0.008 & 0.044\\
\hspace{1em}Proportional to history of emissions & 0.029 & 0.017 & 0.002 & 0.056\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}(Baseline = Less penalty) &  &  &  & \\
\hspace{1em}Same & 0.058 & 0.015 & 0.034 & 0.082\\
\hspace{1em}More penalty & 0.096 & 0.015 & 0.072 & 0.121\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}(Baseline = One year) &  &  &  & \\
\hspace{1em}Five years & 0.052 & 0.018 & 0.023 & 0.082\\
\hspace{1em}Twenty years & -0.012 & 0.018 & -0.041 & 0.017\\
\hspace{1em}Fifty years & -0.093 & 0.018 & -0.124 & -0.063\\
\hspace{1em}Never & -0.132 & 0.018 & -0.162 & -0.102\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}(Baseline = None) &  &  &  & \\
\hspace{1em}Blacklist & 0.102 & 0.017 & 0.074 & 0.13\\
\hspace{1em}Fines & 0.132 & 0.016 & 0.105 & 0.158\\
\hspace{1em}Imprisonment & 0.035 & 0.018 & 0.006 & 0.065\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}(Baseline = Federal government) &  &  &  & \\
\hspace{1em}NGOs & -0.031 & 0.019 & -0.062 & 0\\
\hspace{1em}Local community members & 0.024 & 0.019 & -0.008 & 0.055\\
\hspace{1em}Local government & 0.011 & 0.018 & -0.019 & 0.041\\
\hspace{1em}International organisations & 0.058 & 0.018 & 0.029 & 0.088\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Plot AMCE -- Full Model}
\KeywordTok{plot}\NormalTok{(results, }\DataTypeTok{ci =} \FloatTok{0.9}\NormalTok{, }\DataTypeTok{point.size =} \FloatTok{.8}\NormalTok{, }\DataTypeTok{dodge.size =} \DecValTok{1}\NormalTok{,}
     \DataTypeTok{text.size =} \DecValTok{10}\NormalTok{, }\DataTypeTok{main =} \StringTok{'AMCE -- Full Model'}\NormalTok{,}
     \DataTypeTok{attribute.names =}\NormalTok{ attrs)}
\end{Highlighting}
\end{Shaded}

\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/unnamed-chunk-17-1.pdf}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## Country subsample}

\CommentTok{## Argentina}
\NormalTok{results <-}\StringTok{ }\NormalTok{cjoint}\OperatorTok{::}\KeywordTok{amce}\NormalTok{(fm, }\DataTypeTok{data =}\NormalTok{ cj, }\DataTypeTok{cluster =} \OtherTok{TRUE}\NormalTok{,}
                        \DataTypeTok{respondent.id =} \StringTok{"Response.ID"}\NormalTok{, }
                        \DataTypeTok{design =}\NormalTok{ conjDesign, }\DataTypeTok{baselines =}\NormalTok{ baselines,}
                        \DataTypeTok{subset =}\NormalTok{ cj}\OperatorTok{$}\NormalTok{countryOrigin }\OperatorTok{==}\StringTok{ 'Argentina'}\NormalTok{)}
\CommentTok{# AMCE -- Argentina Only}
\KeywordTok{tableAMCE}\NormalTok{(results, }\DataTypeTok{capt =} \StringTok{'AMCE -- Argentina Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-18}AMCE -- Argentina Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lcccc}
\toprule
Feature & Estimate & Std.Error & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}(Baseline = Government bureaucracy) &  &  &  & \\
\hspace{1em}Informal norms & -0.09 & 0.061 & -0.191 & 0.01\\
\hspace{1em}Private arbitration & 0.055 & 0.068 & -0.056 & 0.167\\
\hspace{1em}Local courts & 0.103 & 0.062 & 0.001 & 0.204\\
\hspace{1em}United Nations & 0.125 & 0.067 & 0.016 & 0.235\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}(Baseline = Rich countries pay more than poor countries) &  &  &  & \\
\hspace{1em}Only rich countries pay & -0.078 & 0.054 & -0.166 & 0.01\\
\hspace{1em}Proportional to current emissions & 0.031 & 0.056 & -0.062 & 0.124\\
\hspace{1em}Proportional to history of emissions & 0.09 & 0.048 & 0.011 & 0.169\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}(Baseline = Less penalty) &  &  &  & \\
\hspace{1em}Same & 0.095 & 0.046 & 0.018 & 0.171\\
\hspace{1em}More penalty & 0.122 & 0.055 & 0.031 & 0.213\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}(Baseline = One year) &  &  &  & \\
\hspace{1em}Five years & -0.061 & 0.06 & -0.159 & 0.037\\
\hspace{1em}Twenty years & -0.132 & 0.068 & -0.243 & -0.021\\
\hspace{1em}Fifty years & -0.166 & 0.068 & -0.278 & -0.054\\
\hspace{1em}Never & -0.171 & 0.061 & -0.271 & -0.071\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}(Baseline = None) &  &  &  & \\
\hspace{1em}Blacklist & 0.185 & 0.056 & 0.093 & 0.278\\
\hspace{1em}Fines & 0.262 & 0.053 & 0.174 & 0.35\\
\hspace{1em}Imprisonment & 0.154 & 0.062 & 0.052 & 0.256\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}(Baseline = Federal government) &  &  &  & \\
\hspace{1em}NGOs & -0.118 & 0.069 & -0.23 & -0.005\\
\hspace{1em}Local community members & -0.074 & 0.068 & -0.185 & 0.038\\
\hspace{1em}Local government & 0.003 & 0.067 & -0.108 & 0.114\\
\hspace{1em}International organisations & 0.08 & 0.06 & -0.018 & 0.178\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(results, }\DataTypeTok{ci=}\FloatTok{0.9}\NormalTok{, }\DataTypeTok{point.size =} \FloatTok{.8}\NormalTok{, }\DataTypeTok{dodge.size =} \DecValTok{1}\NormalTok{,}
     \DataTypeTok{text.size =} \DecValTok{10}\NormalTok{, }\DataTypeTok{main =} \StringTok{'AMCE -- Argentina Only'}\NormalTok{,}
     \DataTypeTok{attribute.names =}\NormalTok{ attrs)}
\end{Highlighting}
\end{Shaded}

\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/unnamed-chunk-19-1.pdf}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## Bolivia}
\NormalTok{results <-}\StringTok{ }\NormalTok{cjoint}\OperatorTok{::}\KeywordTok{amce}\NormalTok{(fm, }\DataTypeTok{data =}\NormalTok{ cj, }\DataTypeTok{cluster =} \OtherTok{TRUE}\NormalTok{,}
                        \DataTypeTok{respondent.id =} \StringTok{"Response.ID"}\NormalTok{, }
                        \DataTypeTok{design =}\NormalTok{ conjDesign, }\DataTypeTok{baselines =}\NormalTok{ baselines,}
                        \DataTypeTok{subset =}\NormalTok{ cj}\OperatorTok{$}\NormalTok{countryOrigin }\OperatorTok{==}\StringTok{ 'Bolivia'}\NormalTok{)}
\CommentTok{# Table AMCE -- Full Model }
\KeywordTok{tableAMCE}\NormalTok{(results, }\DataTypeTok{capt =} \StringTok{'AMCE -- Bolivia Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-20}AMCE -- Bolivia Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lcccc}
\toprule
Feature & Estimate & Std.Error & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}(Baseline = Government bureaucracy) &  &  &  & \\
\hspace{1em}Informal norms & -0.056 & 0.067 & -0.167 & 0.054\\
\hspace{1em}Private arbitration & 0.127 & 0.074 & 0.005 & 0.249\\
\hspace{1em}Local courts & 0.093 & 0.057 & -0.001 & 0.186\\
\hspace{1em}United Nations & 0.17 & 0.08 & 0.038 & 0.302\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}(Baseline = Rich countries pay more than poor countries) &  &  &  & \\
\hspace{1em}Only rich countries pay & -0.142 & 0.068 & -0.254 & -0.029\\
\hspace{1em}Proportional to current emissions & 0.012 & 0.066 & -0.097 & 0.12\\
\hspace{1em}Proportional to history of emissions & -0.149 & 0.058 & -0.245 & -0.054\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}(Baseline = Less penalty) &  &  &  & \\
\hspace{1em}Same & 0.126 & 0.052 & 0.039 & 0.212\\
\hspace{1em}More penalty & 0.117 & 0.061 & 0.016 & 0.218\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}(Baseline = One year) &  &  &  & \\
\hspace{1em}Five years & 0.125 & 0.072 & 0.006 & 0.244\\
\hspace{1em}Twenty years & 0.075 & 0.077 & -0.052 & 0.202\\
\hspace{1em}Fifty years & -0.116 & 0.065 & -0.223 & -0.008\\
\hspace{1em}Never & -0.114 & 0.078 & -0.243 & 0.015\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}(Baseline = None) &  &  &  & \\
\hspace{1em}Blacklist & 0.037 & 0.075 & -0.086 & 0.16\\
\hspace{1em}Fines & 0.155 & 0.056 & 0.063 & 0.247\\
\hspace{1em}Imprisonment & 0.069 & 0.074 & -0.052 & 0.19\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}(Baseline = Federal government) &  &  &  & \\
\hspace{1em}NGOs & -0.033 & 0.071 & -0.15 & 0.084\\
\hspace{1em}Local community members & 0.122 & 0.08 & -0.009 & 0.253\\
\hspace{1em}Local government & 0.061 & 0.07 & -0.055 & 0.176\\
\hspace{1em}International organisations & 0.061 & 0.076 & -0.065 & 0.186\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(results, }\DataTypeTok{ci=}\FloatTok{0.9}\NormalTok{, }\DataTypeTok{point.size =} \FloatTok{.8}\NormalTok{, }\DataTypeTok{dodge.size =} \DecValTok{1}\NormalTok{,}
     \DataTypeTok{text.size =} \DecValTok{10}\NormalTok{, }\DataTypeTok{main =} \StringTok{'AMCE -- Bolivia Only'}\NormalTok{,}
     \DataTypeTok{attribute.names =}\NormalTok{ attrs)}
\end{Highlighting}
\end{Shaded}

\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/unnamed-chunk-21-1.pdf}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## Brazil}
\NormalTok{results <-}\StringTok{ }\NormalTok{cjoint}\OperatorTok{::}\KeywordTok{amce}\NormalTok{(fm, }\DataTypeTok{data =}\NormalTok{ cj, }\DataTypeTok{cluster =} \OtherTok{TRUE}\NormalTok{,}
                        \DataTypeTok{respondent.id =} \StringTok{"Response.ID"}\NormalTok{, }
                        \DataTypeTok{design =}\NormalTok{ conjDesign, }\DataTypeTok{baselines =}\NormalTok{ baselines,}
                        \DataTypeTok{subset =}\NormalTok{ cj}\OperatorTok{$}\NormalTok{countryOrigin }\OperatorTok{==}\StringTok{ 'Brazil'}\NormalTok{)}
\CommentTok{# Table AMCE -- Brazil Only}
\KeywordTok{tableAMCE}\NormalTok{(results, }\DataTypeTok{capt =} \StringTok{'AMCE -- Brazil Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-22}AMCE -- Brazil Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lcccc}
\toprule
Feature & Estimate & Std.Error & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}(Baseline = Government bureaucracy) &  &  &  & \\
\hspace{1em}Informal norms & -0.03 & 0.054 & -0.118 & 0.058\\
\hspace{1em}Private arbitration & 0.021 & 0.044 & -0.052 & 0.093\\
\hspace{1em}Local courts & 0.091 & 0.05 & 0.009 & 0.174\\
\hspace{1em}United Nations & 0.082 & 0.047 & 0.004 & 0.16\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}(Baseline = Rich countries pay more than poor countries) &  &  &  & \\
\hspace{1em}Only rich countries pay & -0.107 & 0.051 & -0.191 & -0.023\\
\hspace{1em}Proportional to current emissions & -0.022 & 0.05 & -0.104 & 0.06\\
\hspace{1em}Proportional to history of emissions & 0.021 & 0.049 & -0.059 & 0.101\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}(Baseline = Less penalty) &  &  &  & \\
\hspace{1em}Same & 0.134 & 0.039 & 0.069 & 0.199\\
\hspace{1em}More penalty & 0.169 & 0.044 & 0.097 & 0.241\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}(Baseline = One year) &  &  &  & \\
\hspace{1em}Five years & 0.025 & 0.044 & -0.048 & 0.098\\
\hspace{1em}Twenty years & -0.012 & 0.047 & -0.09 & 0.065\\
\hspace{1em}Fifty years & -0.087 & 0.057 & -0.181 & 0.007\\
\hspace{1em}Never & -0.166 & 0.054 & -0.256 & -0.077\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}(Baseline = None) &  &  &  & \\
\hspace{1em}Blacklist & 0.13 & 0.045 & 0.055 & 0.205\\
\hspace{1em}Fines & 0.169 & 0.042 & 0.1 & 0.238\\
\hspace{1em}Imprisonment & 0.032 & 0.055 & -0.058 & 0.122\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}(Baseline = Federal government) &  &  &  & \\
\hspace{1em}NGOs & 0.001 & 0.051 & -0.083 & 0.086\\
\hspace{1em}Local community members & 0.003 & 0.054 & -0.086 & 0.092\\
\hspace{1em}Local government & -0.017 & 0.053 & -0.105 & 0.071\\
\hspace{1em}International organisations & 0.087 & 0.049 & 0.008 & 0.167\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(results, }\DataTypeTok{ci=}\FloatTok{0.9}\NormalTok{, }\DataTypeTok{point.size =} \FloatTok{.8}\NormalTok{, }\DataTypeTok{dodge.size =} \DecValTok{1}\NormalTok{,}
     \DataTypeTok{text.size =} \DecValTok{10}\NormalTok{, }\DataTypeTok{main =} \StringTok{'AMCE -- Brazil Only'}\NormalTok{,}
     \DataTypeTok{attribute.names =}\NormalTok{ attrs)}
\end{Highlighting}
\end{Shaded}

\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/unnamed-chunk-23-1.pdf}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## Chile}
\NormalTok{results <-}\StringTok{ }\NormalTok{cjoint}\OperatorTok{::}\KeywordTok{amce}\NormalTok{(fm, }\DataTypeTok{data =}\NormalTok{ cj, }\DataTypeTok{cluster =} \OtherTok{TRUE}\NormalTok{,}
                        \DataTypeTok{respondent.id =} \StringTok{"Response.ID"}\NormalTok{, }
                        \DataTypeTok{design =}\NormalTok{ conjDesign, }\DataTypeTok{baselines =}\NormalTok{ baselines,}
                        \DataTypeTok{subset =}\NormalTok{ cj}\OperatorTok{$}\NormalTok{countryOrigin }\OperatorTok{==}\StringTok{ 'Chile'}\NormalTok{)}
\CommentTok{# Table AMCE -- Chile Only}
\KeywordTok{tableAMCE}\NormalTok{(results, }\DataTypeTok{capt =} \StringTok{'AMCE -- Chile Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-24}AMCE -- Chile Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lcccc}
\toprule
Feature & Estimate & Std.Error & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}(Baseline = Government bureaucracy) &  &  &  & \\
\hspace{1em}Informal norms & -0.021 & 0.063 & -0.125 & 0.083\\
\hspace{1em}Private arbitration & 0.055 & 0.049 & -0.026 & 0.136\\
\hspace{1em}Local courts & 0.166 & 0.062 & 0.064 & 0.268\\
\hspace{1em}United Nations & 0.089 & 0.07 & -0.026 & 0.204\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}(Baseline = Rich countries pay more than poor countries) &  &  &  & \\
\hspace{1em}Only rich countries pay & 0.001 & 0.052 & -0.084 & 0.086\\
\hspace{1em}Proportional to current emissions & 0.054 & 0.045 & -0.021 & 0.128\\
\hspace{1em}Proportional to history of emissions & 0.003 & 0.054 & -0.086 & 0.092\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}(Baseline = Less penalty) &  &  &  & \\
\hspace{1em}Same & -0.022 & 0.045 & -0.096 & 0.052\\
\hspace{1em}More penalty & 0.106 & 0.05 & 0.024 & 0.188\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}(Baseline = One year) &  &  &  & \\
\hspace{1em}Five years & 0.207 & 0.053 & 0.12 & 0.294\\
\hspace{1em}Twenty years & 0.083 & 0.049 & 0.003 & 0.163\\
\hspace{1em}Fifty years & 0.057 & 0.066 & -0.052 & 0.165\\
\hspace{1em}Never & -0.029 & 0.062 & -0.13 & 0.072\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}(Baseline = None) &  &  &  & \\
\hspace{1em}Blacklist & 0.193 & 0.049 & 0.112 & 0.274\\
\hspace{1em}Fines & 0.103 & 0.044 & 0.03 & 0.176\\
\hspace{1em}Imprisonment & 0.013 & 0.052 & -0.073 & 0.099\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}(Baseline = Federal government) &  &  &  & \\
\hspace{1em}NGOs & 0.001 & 0.066 & -0.107 & 0.11\\
\hspace{1em}Local community members & 0.196 & 0.063 & 0.092 & 0.301\\
\hspace{1em}Local government & 0.119 & 0.058 & 0.024 & 0.215\\
\hspace{1em}International organisations & 0.069 & 0.061 & -0.03 & 0.169\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(results, }\DataTypeTok{ci=}\FloatTok{0.9}\NormalTok{, }\DataTypeTok{point.size =} \FloatTok{.8}\NormalTok{, }\DataTypeTok{dodge.size =} \DecValTok{1}\NormalTok{,}
     \DataTypeTok{text.size =} \DecValTok{10}\NormalTok{, }\DataTypeTok{main =} \StringTok{'AMCE -- Chile Only'}\NormalTok{,}
     \DataTypeTok{attribute.names =}\NormalTok{ attrs)}
\end{Highlighting}
\end{Shaded}

\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/unnamed-chunk-25-1.pdf}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## Colombia}
\NormalTok{results <-}\StringTok{ }\NormalTok{cjoint}\OperatorTok{::}\KeywordTok{amce}\NormalTok{(fm, }\DataTypeTok{data =}\NormalTok{ cj, }\DataTypeTok{cluster =} \OtherTok{TRUE}\NormalTok{,}
                        \DataTypeTok{respondent.id =} \StringTok{"Response.ID"}\NormalTok{, }
                        \DataTypeTok{design =}\NormalTok{ conjDesign, }\DataTypeTok{baselines =}\NormalTok{ baselines,}
                        \DataTypeTok{subset =}\NormalTok{ cj}\OperatorTok{$}\NormalTok{countryOrigin }\OperatorTok{==}\StringTok{ 'Colombia'}\NormalTok{)}
\CommentTok{# Table AMCE -- Colombia Only}
\KeywordTok{tableAMCE}\NormalTok{(results, }\DataTypeTok{capt =} \StringTok{'AMCE -- Colombia Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-26}AMCE -- Colombia Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lcccc}
\toprule
Feature & Estimate & Std.Error & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}(Baseline = Government bureaucracy) &  &  &  & \\
\hspace{1em}Informal norms & -0.054 & 0.05 & -0.137 & 0.029\\
\hspace{1em}Private arbitration & 0.057 & 0.053 & -0.03 & 0.144\\
\hspace{1em}Local courts & 0.078 & 0.051 & -0.005 & 0.162\\
\hspace{1em}United Nations & 0.132 & 0.054 & 0.043 & 0.221\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}(Baseline = Rich countries pay more than poor countries) &  &  &  & \\
\hspace{1em}Only rich countries pay & -0.017 & 0.053 & -0.105 & 0.07\\
\hspace{1em}Proportional to current emissions & 0.046 & 0.052 & -0.04 & 0.131\\
\hspace{1em}Proportional to history of emissions & 0.059 & 0.06 & -0.039 & 0.158\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}(Baseline = Less penalty) &  &  &  & \\
\hspace{1em}Same & 0.014 & 0.046 & -0.06 & 0.089\\
\hspace{1em}More penalty & 0.051 & 0.044 & -0.021 & 0.124\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}(Baseline = One year) &  &  &  & \\
\hspace{1em}Five years & 0.001 & 0.046 & -0.075 & 0.077\\
\hspace{1em}Twenty years & 0.005 & 0.052 & -0.081 & 0.091\\
\hspace{1em}Fifty years & -0.126 & 0.051 & -0.21 & -0.041\\
\hspace{1em}Never & -0.122 & 0.048 & -0.201 & -0.043\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}(Baseline = None) &  &  &  & \\
\hspace{1em}Blacklist & 0.037 & 0.059 & -0.059 & 0.133\\
\hspace{1em}Fines & 0.068 & 0.052 & -0.019 & 0.154\\
\hspace{1em}Imprisonment & -0.072 & 0.053 & -0.159 & 0.015\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}(Baseline = Federal government) &  &  &  & \\
\hspace{1em}NGOs & -0.023 & 0.059 & -0.12 & 0.074\\
\hspace{1em}Local community members & 0.131 & 0.06 & 0.032 & 0.229\\
\hspace{1em}Local government & -0.003 & 0.049 & -0.083 & 0.078\\
\hspace{1em}International organisations & 0.094 & 0.051 & 0.009 & 0.178\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(results, }\DataTypeTok{ci=}\FloatTok{0.9}\NormalTok{, }\DataTypeTok{point.size =} \FloatTok{.8}\NormalTok{, }\DataTypeTok{dodge.size =} \DecValTok{1}\NormalTok{,}
     \DataTypeTok{text.size =} \DecValTok{10}\NormalTok{, }\DataTypeTok{main =} \StringTok{'AMCE -- Colombia Only'}\NormalTok{,}
     \DataTypeTok{attribute.names =}\NormalTok{ attrs)}
\end{Highlighting}
\end{Shaded}

\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/unnamed-chunk-27-1.pdf}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## Costa Rica}
\NormalTok{results <-}\StringTok{ }\NormalTok{cjoint}\OperatorTok{::}\KeywordTok{amce}\NormalTok{(fm, }\DataTypeTok{data =}\NormalTok{ cj, }\DataTypeTok{cluster =} \OtherTok{TRUE}\NormalTok{,}
                        \DataTypeTok{respondent.id =} \StringTok{"Response.ID"}\NormalTok{, }
                        \DataTypeTok{design =}\NormalTok{ conjDesign, }\DataTypeTok{baselines =}\NormalTok{ baselines,}
                        \DataTypeTok{subset =}\NormalTok{ cj}\OperatorTok{$}\NormalTok{countryOrigin }\OperatorTok{==}\StringTok{ 'Costa Rica'}\NormalTok{)}
\CommentTok{# Table AMCE -- Costa Rica Only}
\KeywordTok{tableAMCE}\NormalTok{(results, }\DataTypeTok{capt =} \StringTok{'AMCE -- Costa Rica Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-28}AMCE -- Costa Rica Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lcccc}
\toprule
Feature & Estimate & Std.Error & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}(Baseline = Government bureaucracy) &  &  &  & \\
\hspace{1em}Informal norms & -0.085 & 0.066 & -0.193 & 0.023\\
\hspace{1em}Private arbitration & -0.018 & 0.064 & -0.123 & 0.087\\
\hspace{1em}Local courts & 0.013 & 0.066 & -0.095 & 0.121\\
\hspace{1em}United Nations & 0.094 & 0.071 & -0.022 & 0.21\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}(Baseline = Rich countries pay more than poor countries) &  &  &  & \\
\hspace{1em}Only rich countries pay & -0.106 & 0.06 & -0.205 & -0.007\\
\hspace{1em}Proportional to current emissions & 0.058 & 0.054 & -0.031 & 0.146\\
\hspace{1em}Proportional to history of emissions & 0.015 & 0.066 & -0.094 & 0.123\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}(Baseline = Less penalty) &  &  &  & \\
\hspace{1em}Same & 0.083 & 0.044 & 0.01 & 0.155\\
\hspace{1em}More penalty & 0.135 & 0.042 & 0.066 & 0.204\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}(Baseline = One year) &  &  &  & \\
\hspace{1em}Five years & 0.108 & 0.063 & 0.004 & 0.211\\
\hspace{1em}Twenty years & -0.115 & 0.063 & -0.219 & -0.011\\
\hspace{1em}Fifty years & -0.183 & 0.06 & -0.282 & -0.084\\
\hspace{1em}Never & -0.201 & 0.061 & -0.303 & -0.1\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}(Baseline = None) &  &  &  & \\
\hspace{1em}Blacklist & 0.139 & 0.051 & 0.055 & 0.223\\
\hspace{1em}Fines & 0.245 & 0.051 & 0.161 & 0.328\\
\hspace{1em}Imprisonment & 0.167 & 0.057 & 0.074 & 0.261\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}(Baseline = Federal government) &  &  &  & \\
\hspace{1em}NGOs & 0.032 & 0.063 & -0.071 & 0.136\\
\hspace{1em}Local community members & -0.081 & 0.074 & -0.203 & 0.042\\
\hspace{1em}Local government & 0.087 & 0.059 & -0.01 & 0.184\\
\hspace{1em}International organisations & -0.042 & 0.068 & -0.154 & 0.069\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(results, }\DataTypeTok{ci=}\FloatTok{0.9}\NormalTok{, }\DataTypeTok{point.size =} \FloatTok{.8}\NormalTok{, }\DataTypeTok{dodge.size =} \DecValTok{1}\NormalTok{,}
     \DataTypeTok{text.size =} \DecValTok{10}\NormalTok{, }\DataTypeTok{main =} \StringTok{'AMCE -- Costa Rica Only'}\NormalTok{,}
     \DataTypeTok{attribute.names =}\NormalTok{ attrs)}
\end{Highlighting}
\end{Shaded}

\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/unnamed-chunk-29-1.pdf}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## Ecuador}
\NormalTok{results <-}\StringTok{ }\NormalTok{cjoint}\OperatorTok{::}\KeywordTok{amce}\NormalTok{(fm, }\DataTypeTok{data =}\NormalTok{ cj, }\DataTypeTok{cluster =} \OtherTok{TRUE}\NormalTok{,}
                        \DataTypeTok{respondent.id =} \StringTok{"Response.ID"}\NormalTok{, }
                        \DataTypeTok{design =}\NormalTok{ conjDesign, }\DataTypeTok{baselines =}\NormalTok{ baselines,}
                        \DataTypeTok{subset =}\NormalTok{ cj}\OperatorTok{$}\NormalTok{countryOrigin }\OperatorTok{==}\StringTok{ 'Ecuador'}\NormalTok{)}
\CommentTok{# Table AMCE -- Ecuador Only}
\KeywordTok{tableAMCE}\NormalTok{(results, }\DataTypeTok{capt =} \StringTok{'AMCE -- Ecuador Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-30}AMCE -- Ecuador Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lcccc}
\toprule
Feature & Estimate & Std.Error & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}(Baseline = Government bureaucracy) &  &  &  & \\
\hspace{1em}Informal norms & -0.007 & 0.055 & -0.098 & 0.084\\
\hspace{1em}Private arbitration & 0.068 & 0.052 & -0.018 & 0.154\\
\hspace{1em}Local courts & 0.122 & 0.049 & 0.042 & 0.203\\
\hspace{1em}United Nations & 0.151 & 0.058 & 0.057 & 0.246\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}(Baseline = Rich countries pay more than poor countries) &  &  &  & \\
\hspace{1em}Only rich countries pay & -0.044 & 0.054 & -0.134 & 0.045\\
\hspace{1em}Proportional to current emissions & 0.114 & 0.05 & 0.032 & 0.196\\
\hspace{1em}Proportional to history of emissions & 0.085 & 0.055 & -0.005 & 0.175\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}(Baseline = Less penalty) &  &  &  & \\
\hspace{1em}Same & 0.01 & 0.052 & -0.075 & 0.095\\
\hspace{1em}More penalty & 0.085 & 0.05 & 0.003 & 0.167\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}(Baseline = One year) &  &  &  & \\
\hspace{1em}Five years & 0.085 & 0.063 & -0.018 & 0.188\\
\hspace{1em}Twenty years & 0.076 & 0.05 & -0.006 & 0.159\\
\hspace{1em}Fifty years & -0.017 & 0.05 & -0.1 & 0.065\\
\hspace{1em}Never & 0.036 & 0.066 & -0.073 & 0.144\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}(Baseline = None) &  &  &  & \\
\hspace{1em}Blacklist & 0.079 & 0.049 & -0.002 & 0.16\\
\hspace{1em}Fines & 0.062 & 0.055 & -0.029 & 0.153\\
\hspace{1em}Imprisonment & -0.087 & 0.055 & -0.178 & 0.003\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}(Baseline = Federal government) &  &  &  & \\
\hspace{1em}NGOs & -0.037 & 0.054 & -0.125 & 0.051\\
\hspace{1em}Local community members & 0.063 & 0.055 & -0.027 & 0.153\\
\hspace{1em}Local government & 0.043 & 0.061 & -0.057 & 0.142\\
\hspace{1em}International organisations & 0.051 & 0.055 & -0.04 & 0.141\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(results, }\DataTypeTok{ci=}\FloatTok{0.9}\NormalTok{, }\DataTypeTok{point.size =} \FloatTok{.8}\NormalTok{, }\DataTypeTok{dodge.size =} \DecValTok{1}\NormalTok{,}
     \DataTypeTok{text.size =} \DecValTok{10}\NormalTok{, }\DataTypeTok{main =} \StringTok{'AMCE -- Ecuador Only'}\NormalTok{,}
     \DataTypeTok{attribute.names =}\NormalTok{ attrs)}
\end{Highlighting}
\end{Shaded}

\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/unnamed-chunk-31-1.pdf}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## Mexico}
\NormalTok{results <-}\StringTok{ }\NormalTok{cjoint}\OperatorTok{::}\KeywordTok{amce}\NormalTok{(fm, }\DataTypeTok{data =}\NormalTok{ cj, }\DataTypeTok{cluster =} \OtherTok{TRUE}\NormalTok{,}
                        \DataTypeTok{respondent.id =} \StringTok{"Response.ID"}\NormalTok{, }
                        \DataTypeTok{design =}\NormalTok{ conjDesign, }\DataTypeTok{baselines =}\NormalTok{ baselines,}
                        \DataTypeTok{subset =}\NormalTok{ cj}\OperatorTok{$}\NormalTok{countryOrigin }\OperatorTok{==}\StringTok{ 'Mexico'}\NormalTok{)}
\CommentTok{# Table AMCE -- Full Model }
\KeywordTok{tableAMCE}\NormalTok{(results, }\DataTypeTok{capt =} \StringTok{'AMCE -- Mexico Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-32}AMCE -- Mexico Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lcccc}
\toprule
Feature & Estimate & Std.Error & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}(Baseline = Government bureaucracy) &  &  &  & \\
\hspace{1em}Informal norms & -0.048 & 0.057 & -0.142 & 0.047\\
\hspace{1em}Private arbitration & 0.066 & 0.054 & -0.023 & 0.154\\
\hspace{1em}Local courts & 0.036 & 0.056 & -0.056 & 0.127\\
\hspace{1em}United Nations & 0.137 & 0.063 & 0.033 & 0.242\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}(Baseline = Rich countries pay more than poor countries) &  &  &  & \\
\hspace{1em}Only rich countries pay & -0.059 & 0.053 & -0.146 & 0.028\\
\hspace{1em}Proportional to current emissions & 0.039 & 0.052 & -0.046 & 0.123\\
\hspace{1em}Proportional to history of emissions & 0.132 & 0.057 & 0.038 & 0.227\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}(Baseline = Less penalty) &  &  &  & \\
\hspace{1em}Same & 0.017 & 0.055 & -0.073 & 0.107\\
\hspace{1em}More penalty & 0.011 & 0.052 & -0.074 & 0.096\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}(Baseline = One year) &  &  &  & \\
\hspace{1em}Five years & 0.107 & 0.059 & 0.011 & 0.204\\
\hspace{1em}Twenty years & -0.023 & 0.06 & -0.122 & 0.077\\
\hspace{1em}Fifty years & -0.032 & 0.058 & -0.127 & 0.063\\
\hspace{1em}Never & -0.137 & 0.062 & -0.24 & -0.035\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}(Baseline = None) &  &  &  & \\
\hspace{1em}Blacklist & 0.033 & 0.054 & -0.056 & 0.122\\
\hspace{1em}Fines & 0.109 & 0.056 & 0.018 & 0.201\\
\hspace{1em}Imprisonment & -0.016 & 0.053 & -0.103 & 0.071\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}(Baseline = Federal government) &  &  &  & \\
\hspace{1em}NGOs & -0.038 & 0.068 & -0.15 & 0.074\\
\hspace{1em}Local community members & -0.038 & 0.062 & -0.141 & 0.064\\
\hspace{1em}Local government & -0.063 & 0.065 & -0.17 & 0.043\\
\hspace{1em}International organisations & 0.136 & 0.056 & 0.044 & 0.229\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(results, }\DataTypeTok{ci=}\FloatTok{0.9}\NormalTok{, }\DataTypeTok{point.size =} \FloatTok{.8}\NormalTok{, }\DataTypeTok{dodge.size =} \DecValTok{1}\NormalTok{,}
     \DataTypeTok{text.size =} \DecValTok{10}\NormalTok{, }\DataTypeTok{main =} \StringTok{'AMCE -- Mexico Only'}\NormalTok{,}
     \DataTypeTok{attribute.names =}\NormalTok{ attrs)}
\end{Highlighting}
\end{Shaded}

\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/unnamed-chunk-33-1.pdf}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## Panama}
\NormalTok{results <-}\StringTok{ }\NormalTok{cjoint}\OperatorTok{::}\KeywordTok{amce}\NormalTok{(fm, }\DataTypeTok{data =}\NormalTok{ cj, }\DataTypeTok{cluster =} \OtherTok{TRUE}\NormalTok{,}
                        \DataTypeTok{respondent.id =} \StringTok{"Response.ID"}\NormalTok{, }
                        \DataTypeTok{design =}\NormalTok{ conjDesign, }\DataTypeTok{baselines =}\NormalTok{ baselines,}
                        \DataTypeTok{subset =}\NormalTok{ cj}\OperatorTok{$}\NormalTok{countryOrigin }\OperatorTok{==}\StringTok{ 'Panama'}\NormalTok{)}
\CommentTok{# Table AMCE -- Panama Only}
\KeywordTok{tableAMCE}\NormalTok{(results, }\DataTypeTok{capt =} \StringTok{'AMCE -- Panama Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-34}AMCE -- Panama Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lcccc}
\toprule
Feature & Estimate & Std.Error & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}(Baseline = Government bureaucracy) &  &  &  & \\
\hspace{1em}Informal norms & -0.049 & 0.062 & -0.151 & 0.054\\
\hspace{1em}Private arbitration & 0.013 & 0.065 & -0.095 & 0.12\\
\hspace{1em}Local courts & 0.079 & 0.057 & -0.014 & 0.173\\
\hspace{1em}United Nations & 0.082 & 0.065 & -0.026 & 0.189\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}(Baseline = Rich countries pay more than poor countries) &  &  &  & \\
\hspace{1em}Only rich countries pay & -0.085 & 0.055 & -0.175 & 0.006\\
\hspace{1em}Proportional to current emissions & -0.043 & 0.05 & -0.124 & 0.039\\
\hspace{1em}Proportional to history of emissions & -0.038 & 0.051 & -0.122 & 0.046\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}(Baseline = Less penalty) &  &  &  & \\
\hspace{1em}Same & 0.065 & 0.05 & -0.017 & 0.147\\
\hspace{1em}More penalty & 0.189 & 0.052 & 0.104 & 0.274\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}(Baseline = One year) &  &  &  & \\
\hspace{1em}Five years & -0.067 & 0.058 & -0.162 & 0.028\\
\hspace{1em}Twenty years & -0.045 & 0.064 & -0.151 & 0.06\\
\hspace{1em}Fifty years & -0.196 & 0.064 & -0.301 & -0.092\\
\hspace{1em}Never & -0.269 & 0.057 & -0.363 & -0.175\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}(Baseline = None) &  &  &  & \\
\hspace{1em}Blacklist & 0.112 & 0.057 & 0.018 & 0.205\\
\hspace{1em}Fines & 0.129 & 0.051 & 0.045 & 0.214\\
\hspace{1em}Imprisonment & 0.072 & 0.061 & -0.028 & 0.172\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}(Baseline = Federal government) &  &  &  & \\
\hspace{1em}NGOs & 0.002 & 0.057 & -0.092 & 0.096\\
\hspace{1em}Local community members & 0.039 & 0.06 & -0.06 & 0.137\\
\hspace{1em}Local government & -0.001 & 0.059 & -0.099 & 0.097\\
\hspace{1em}International organisations & 0.034 & 0.059 & -0.064 & 0.131\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(results, }\DataTypeTok{ci=}\FloatTok{0.9}\NormalTok{, }\DataTypeTok{point.size =} \FloatTok{.8}\NormalTok{, }\DataTypeTok{dodge.size =} \DecValTok{1}\NormalTok{,}
     \DataTypeTok{text.size =} \DecValTok{10}\NormalTok{, }\DataTypeTok{main =} \StringTok{'AMCE -- Panama Only'}\NormalTok{,}
     \DataTypeTok{attribute.names =}\NormalTok{ attrs)}
\end{Highlighting}
\end{Shaded}

\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/unnamed-chunk-35-1.pdf}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## Peru}
\NormalTok{results <-}\StringTok{ }\NormalTok{cjoint}\OperatorTok{::}\KeywordTok{amce}\NormalTok{(fm, }\DataTypeTok{data =}\NormalTok{ cj, }\DataTypeTok{cluster =} \OtherTok{TRUE}\NormalTok{,}
                        \DataTypeTok{respondent.id =} \StringTok{"Response.ID"}\NormalTok{, }
                        \DataTypeTok{design =}\NormalTok{ conjDesign, }\DataTypeTok{baselines =}\NormalTok{ baselines,}
                        \DataTypeTok{subset =}\NormalTok{ cj}\OperatorTok{$}\NormalTok{countryOrigin }\OperatorTok{==}\StringTok{ 'Peru'}\NormalTok{)}
\CommentTok{# Table AMCE -- Peru Only}
\KeywordTok{tableAMCE}\NormalTok{(results, }\DataTypeTok{capt =} \StringTok{'AMCE -- Peru Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-36}AMCE -- Peru Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lcccc}
\toprule
Feature & Estimate & Std.Error & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}(Baseline = Government bureaucracy) &  &  &  & \\
\hspace{1em}Informal norms & -0.121 & 0.065 & -0.228 & -0.014\\
\hspace{1em}Private arbitration & 0.123 & 0.069 & 0.01 & 0.236\\
\hspace{1em}Local courts & 0.033 & 0.065 & -0.074 & 0.14\\
\hspace{1em}United Nations & 0.05 & 0.066 & -0.058 & 0.158\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}(Baseline = Rich countries pay more than poor countries) &  &  &  & \\
\hspace{1em}Only rich countries pay & -0.026 & 0.047 & -0.103 & 0.052\\
\hspace{1em}Proportional to current emissions & 0.013 & 0.055 & -0.077 & 0.103\\
\hspace{1em}Proportional to history of emissions & 0.086 & 0.057 & -0.008 & 0.18\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}(Baseline = Less penalty) &  &  &  & \\
\hspace{1em}Same & 0.079 & 0.052 & -0.006 & 0.164\\
\hspace{1em}More penalty & 0.044 & 0.05 & -0.037 & 0.126\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}(Baseline = One year) &  &  &  & \\
\hspace{1em}Five years & 0.014 & 0.071 & -0.104 & 0.131\\
\hspace{1em}Twenty years & -0.023 & 0.077 & -0.15 & 0.104\\
\hspace{1em}Fifty years & -0.12 & 0.07 & -0.235 & -0.006\\
\hspace{1em}Never & -0.12 & 0.062 & -0.222 & -0.018\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}(Baseline = None) &  &  &  & \\
\hspace{1em}Blacklist & 0.093 & 0.075 & -0.03 & 0.215\\
\hspace{1em}Fines & 0.128 & 0.069 & 0.014 & 0.243\\
\hspace{1em}Imprisonment & 0.075 & 0.071 & -0.042 & 0.191\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}(Baseline = Federal government) &  &  &  & \\
\hspace{1em}NGOs & -0.099 & 0.063 & -0.202 & 0.004\\
\hspace{1em}Local community members & -0.119 & 0.068 & -0.232 & -0.007\\
\hspace{1em}Local government & 0.029 & 0.075 & -0.095 & 0.153\\
\hspace{1em}International organisations & 0.004 & 0.067 & -0.106 & 0.114\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(results, }\DataTypeTok{ci=}\FloatTok{0.9}\NormalTok{, }\DataTypeTok{point.size =} \FloatTok{.8}\NormalTok{, }\DataTypeTok{dodge.size =} \DecValTok{1}\NormalTok{,}
     \DataTypeTok{text.size =} \DecValTok{10}\NormalTok{, }\DataTypeTok{main =} \StringTok{'AMCE -- Peru Only'}\NormalTok{,}
     \DataTypeTok{attribute.names =}\NormalTok{ attrs)}
\end{Highlighting}
\end{Shaded}

\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/unnamed-chunk-37-1.pdf}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## Elite Subsample}

\CommentTok{# Executive branch}
\NormalTok{results <-}\StringTok{ }\NormalTok{cjoint}\OperatorTok{::}\KeywordTok{amce}\NormalTok{(fm, }\DataTypeTok{data =}\NormalTok{ cj, }\DataTypeTok{cluster =} \OtherTok{TRUE}\NormalTok{,}
                        \DataTypeTok{respondent.id =} \StringTok{"Response.ID"}\NormalTok{, }
                        \DataTypeTok{design =}\NormalTok{ conjDesign, }\DataTypeTok{baselines =}\NormalTok{ baselines,}
                        \DataTypeTok{subset =}\NormalTok{ cj}\OperatorTok{$}\NormalTok{groupOrigin }\OperatorTok{==}\StringTok{ 'Executive'}\NormalTok{)}
\CommentTok{# Table AMCE -- Executive branch only}
\KeywordTok{tableAMCE}\NormalTok{(results, }\DataTypeTok{capt =} \StringTok{'AMCE -- Executive Branch Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-38}AMCE -- Executive Branch Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lcccc}
\toprule
Feature & Estimate & Std.Error & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}(Baseline = Government bureaucracy) &  &  &  & \\
\hspace{1em}Informal norms & -0.034 & 0.05 & -0.116 & 0.049\\
\hspace{1em}Private arbitration & 0.057 & 0.039 & -0.008 & 0.121\\
\hspace{1em}Local courts & 0.067 & 0.043 & -0.003 & 0.137\\
\hspace{1em}United Nations & 0.096 & 0.047 & 0.019 & 0.174\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}(Baseline = Rich countries pay more than poor countries) &  &  &  & \\
\hspace{1em}Only rich countries pay & -0.036 & 0.039 & -0.101 & 0.028\\
\hspace{1em}Proportional to current emissions & 0.08 & 0.04 & 0.014 & 0.146\\
\hspace{1em}Proportional to history of emissions & 0.055 & 0.044 & -0.017 & 0.127\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}(Baseline = Less penalty) &  &  &  & \\
\hspace{1em}Same & 0.081 & 0.038 & 0.018 & 0.144\\
\hspace{1em}More penalty & 0.125 & 0.039 & 0.061 & 0.188\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}(Baseline = One year) &  &  &  & \\
\hspace{1em}Five years & 0.06 & 0.049 & -0.02 & 0.14\\
\hspace{1em}Twenty years & -0.023 & 0.048 & -0.102 & 0.056\\
\hspace{1em}Fifty years & -0.096 & 0.045 & -0.171 & -0.022\\
\hspace{1em}Never & -0.119 & 0.045 & -0.193 & -0.045\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}(Baseline = None) &  &  &  & \\
\hspace{1em}Blacklist & 0.103 & 0.037 & 0.043 & 0.163\\
\hspace{1em}Fines & 0.125 & 0.044 & 0.053 & 0.197\\
\hspace{1em}Imprisonment & -0.054 & 0.04 & -0.12 & 0.013\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}(Baseline = Federal government) &  &  &  & \\
\hspace{1em}NGOs & -0.159 & 0.051 & -0.243 & -0.076\\
\hspace{1em}Local community members & -0.027 & 0.052 & -0.113 & 0.058\\
\hspace{1em}Local government & -0.03 & 0.045 & -0.104 & 0.044\\
\hspace{1em}International organisations & 0.021 & 0.043 & -0.048 & 0.091\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(results, }\DataTypeTok{ci=}\FloatTok{0.9}\NormalTok{, }\DataTypeTok{point.size =} \FloatTok{.8}\NormalTok{, }\DataTypeTok{dodge.size =} \DecValTok{1}\NormalTok{,}
     \DataTypeTok{text.size =} \DecValTok{10}\NormalTok{, }\DataTypeTok{main =} \StringTok{'AMCE -- Executive Branch Only'}\NormalTok{,}
     \DataTypeTok{attribute.names =}\NormalTok{ attrs)}
\end{Highlighting}
\end{Shaded}

\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/unnamed-chunk-39-1.pdf}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## Legislative branch}
\NormalTok{results <-}\StringTok{ }\NormalTok{cjoint}\OperatorTok{::}\KeywordTok{amce}\NormalTok{(fm, }\DataTypeTok{data =}\NormalTok{ cj, }\DataTypeTok{cluster =} \OtherTok{TRUE}\NormalTok{,}
                        \DataTypeTok{respondent.id =} \StringTok{"Response.ID"}\NormalTok{, }
                        \DataTypeTok{design =}\NormalTok{ conjDesign, }\DataTypeTok{baselines =}\NormalTok{ baselines,}
                        \DataTypeTok{subset =}\NormalTok{ cj}\OperatorTok{$}\NormalTok{groupOrigin }\OperatorTok{==}\StringTok{ 'Legislative'}\NormalTok{)}
\CommentTok{# Table AMCE -- Legislative branch only}
\KeywordTok{tableAMCE}\NormalTok{(results, }\DataTypeTok{capt =} \StringTok{'AMCE -- Legislative Branch Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-40}AMCE -- Legislative Branch Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lcccc}
\toprule
Feature & Estimate & Std.Error & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}(Baseline = Government bureaucracy) &  &  &  & \\
\hspace{1em}Informal norms & -0.069 & 0.054 & -0.157 & 0.02\\
\hspace{1em}Private arbitration & 0.018 & 0.065 & -0.09 & 0.126\\
\hspace{1em}Local courts & 0.044 & 0.062 & -0.058 & 0.147\\
\hspace{1em}United Nations & 0.065 & 0.06 & -0.034 & 0.164\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}(Baseline = Rich countries pay more than poor countries) &  &  &  & \\
\hspace{1em}Only rich countries pay & -0.158 & 0.052 & -0.243 & -0.073\\
\hspace{1em}Proportional to current emissions & -0.08 & 0.052 & -0.166 & 0.005\\
\hspace{1em}Proportional to history of emissions & -0.019 & 0.049 & -0.099 & 0.062\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}(Baseline = Less penalty) &  &  &  & \\
\hspace{1em}Same & 0.074 & 0.044 & 0.002 & 0.147\\
\hspace{1em}More penalty & 0.071 & 0.044 & -0.001 & 0.143\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}(Baseline = One year) &  &  &  & \\
\hspace{1em}Five years & 0.003 & 0.059 & -0.095 & 0.101\\
\hspace{1em}Twenty years & 0.002 & 0.058 & -0.093 & 0.097\\
\hspace{1em}Fifty years & -0.043 & 0.061 & -0.143 & 0.057\\
\hspace{1em}Never & -0.129 & 0.066 & -0.237 & -0.02\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}(Baseline = None) &  &  &  & \\
\hspace{1em}Blacklist & 0.113 & 0.055 & 0.022 & 0.204\\
\hspace{1em}Fines & 0.145 & 0.054 & 0.056 & 0.233\\
\hspace{1em}Imprisonment & -0.021 & 0.058 & -0.116 & 0.075\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}(Baseline = Federal government) &  &  &  & \\
\hspace{1em}NGOs & -0.116 & 0.064 & -0.22 & -0.011\\
\hspace{1em}Local community members & -0.003 & 0.068 & -0.115 & 0.108\\
\hspace{1em}Local government & 0.024 & 0.063 & -0.08 & 0.128\\
\hspace{1em}International organisations & -0.001 & 0.065 & -0.108 & 0.105\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(results, }\DataTypeTok{ci=}\FloatTok{0.9}\NormalTok{, }\DataTypeTok{point.size =} \FloatTok{.8}\NormalTok{, }\DataTypeTok{dodge.size =} \DecValTok{1}\NormalTok{,}
     \DataTypeTok{text.size =} \DecValTok{10}\NormalTok{, }\DataTypeTok{main =} \StringTok{'AMCE -- Legislative Branch Only'}\NormalTok{,}
     \DataTypeTok{attribute.names =}\NormalTok{ attrs)}
\end{Highlighting}
\end{Shaded}

\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/unnamed-chunk-41-1.pdf}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## Civil society}
\NormalTok{results <-}\StringTok{ }\NormalTok{cjoint}\OperatorTok{::}\KeywordTok{amce}\NormalTok{(fm, }\DataTypeTok{data =}\NormalTok{ cj, }\DataTypeTok{cluster =} \OtherTok{TRUE}\NormalTok{,}
                        \DataTypeTok{respondent.id =} \StringTok{"Response.ID"}\NormalTok{, }
                        \DataTypeTok{design =}\NormalTok{ conjDesign, }\DataTypeTok{baselines =}\NormalTok{ baselines,}
                        \DataTypeTok{subset =}\NormalTok{ cj}\OperatorTok{$}\NormalTok{groupOrigin }\OperatorTok{==}\StringTok{ 'Civil Society'}\NormalTok{)}
\CommentTok{# Table AMCE -- Civil Society Only }
\KeywordTok{tableAMCE}\NormalTok{(results, }\DataTypeTok{capt =} \StringTok{'AMCE -- Civil Society Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-42}AMCE -- Civil Society Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lcccc}
\toprule
Feature & Estimate & Std.Error & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}(Baseline = Government bureaucracy) &  &  &  & \\
\hspace{1em}Informal norms & -0.02 & 0.031 & -0.071 & 0.032\\
\hspace{1em}Private arbitration & 0.073 & 0.032 & 0.021 & 0.126\\
\hspace{1em}Local courts & 0.127 & 0.031 & 0.076 & 0.179\\
\hspace{1em}United Nations & 0.166 & 0.031 & 0.115 & 0.218\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}(Baseline = Rich countries pay more than poor countries) &  &  &  & \\
\hspace{1em}Only rich countries pay & -0.043 & 0.031 & -0.093 & 0.007\\
\hspace{1em}Proportional to current emissions & 0.038 & 0.03 & -0.011 & 0.087\\
\hspace{1em}Proportional to history of emissions & 0.069 & 0.029 & 0.022 & 0.117\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}(Baseline = Less penalty) &  &  &  & \\
\hspace{1em}Same & 0.035 & 0.025 & -0.006 & 0.077\\
\hspace{1em}More penalty & 0.101 & 0.026 & 0.059 & 0.143\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}(Baseline = One year) &  &  &  & \\
\hspace{1em}Five years & 0.044 & 0.031 & -0.007 & 0.095\\
\hspace{1em}Twenty years & -0.053 & 0.031 & -0.103 & -0.003\\
\hspace{1em}Fifty years & -0.102 & 0.034 & -0.158 & -0.047\\
\hspace{1em}Never & -0.147 & 0.034 & -0.203 & -0.09\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}(Baseline = None) &  &  &  & \\
\hspace{1em}Blacklist & 0.097 & 0.03 & 0.047 & 0.146\\
\hspace{1em}Fines & 0.13 & 0.028 & 0.085 & 0.176\\
\hspace{1em}Imprisonment & 0.049 & 0.03 & -0.002 & 0.099\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}(Baseline = Federal government) &  &  &  & \\
\hspace{1em}NGOs & 0.026 & 0.031 & -0.025 & 0.077\\
\hspace{1em}Local community members & 0.037 & 0.034 & -0.019 & 0.094\\
\hspace{1em}Local government & -0.008 & 0.033 & -0.063 & 0.047\\
\hspace{1em}International organisations & 0.066 & 0.029 & 0.019 & 0.113\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(results, }\DataTypeTok{ci=}\FloatTok{0.9}\NormalTok{, }\DataTypeTok{point.size =} \FloatTok{.8}\NormalTok{, }\DataTypeTok{dodge.size =} \DecValTok{1}\NormalTok{,}
     \DataTypeTok{text.size =} \DecValTok{10}\NormalTok{, }\DataTypeTok{main =} \StringTok{'AMCE -- Civil Society Only'}\NormalTok{,}
     \DataTypeTok{attribute.names =}\NormalTok{ attrs)}
\end{Highlighting}
\end{Shaded}

\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/unnamed-chunk-43-1.pdf}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## Academia Only}
\NormalTok{results <-}\StringTok{ }\NormalTok{cjoint}\OperatorTok{::}\KeywordTok{amce}\NormalTok{(fm, }\DataTypeTok{data =}\NormalTok{ cj, }\DataTypeTok{cluster =} \OtherTok{TRUE}\NormalTok{,}
                        \DataTypeTok{respondent.id =} \StringTok{"Response.ID"}\NormalTok{, }
                        \DataTypeTok{design =}\NormalTok{ conjDesign, }\DataTypeTok{baselines =}\NormalTok{ baselines,}
                        \DataTypeTok{subset =}\NormalTok{ cj}\OperatorTok{$}\NormalTok{groupOrigin }\OperatorTok{==}\StringTok{ 'Academia'}\NormalTok{)}
\CommentTok{# Table AMCE -- Full Model }
\KeywordTok{tableAMCE}\NormalTok{(results, }\DataTypeTok{capt =} \StringTok{'AMCE -- Academia Only'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{\label{tab:unnamed-chunk-44}AMCE -- Academia Only}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{lcccc}
\toprule
Feature & Estimate & Std.Error & Lower & Upper\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are conflicts resolved?}}\\
\hspace{1em}(Baseline = Government bureaucracy) &  &  &  & \\
\hspace{1em}Informal norms & -0.089 & 0.031 & -0.139 & -0.039\\
\hspace{1em}Private arbitration & 0.053 & 0.03 & 0.003 & 0.103\\
\hspace{1em}Local courts & 0.082 & 0.028 & 0.036 & 0.129\\
\hspace{1em}United Nations & 0.09 & 0.034 & 0.034 & 0.146\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are costs distributed?}}\\
\hspace{1em}(Baseline = Rich countries pay more than poor countries) &  &  &  & \\
\hspace{1em}Only rich countries pay & -0.07 & 0.029 & -0.117 & -0.023\\
\hspace{1em}Proportional to current emissions & 0.029 & 0.025 & -0.012 & 0.07\\
\hspace{1em}Proportional to history of emissions & 0.013 & 0.03 & -0.037 & 0.062\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How are repeated violations punished?}}\\
\hspace{1em}(Baseline = Less penalty) &  &  &  & \\
\hspace{1em}Same & 0.068 & 0.025 & 0.027 & 0.11\\
\hspace{1em}More penalty & 0.112 & 0.027 & 0.067 & 0.156\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{How often will the agreement be renegotiated?}}\\
\hspace{1em}(Baseline = One year) &  &  &  & \\
\hspace{1em}Five years & 0.075 & 0.03 & 0.026 & 0.124\\
\hspace{1em}Twenty years & 0.039 & 0.03 & -0.009 & 0.088\\
\hspace{1em}Fifty years & -0.092 & 0.032 & -0.145 & -0.039\\
\hspace{1em}Never & -0.101 & 0.03 & -0.15 & -0.052\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{What punishments do they use?}}\\
\hspace{1em}(Baseline = None) &  &  &  & \\
\hspace{1em}Blacklist & 0.121 & 0.031 & 0.07 & 0.172\\
\hspace{1em}Fines & 0.158 & 0.028 & 0.112 & 0.204\\
\hspace{1em}Imprisonment & 0.087 & 0.032 & 0.034 & 0.14\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Who makes the rules?}}\\
\hspace{1em}(Baseline = Federal government) &  &  &  & \\
\hspace{1em}NGOs & 0 & 0.031 & -0.051 & 0.051\\
\hspace{1em}Local community members & 0.055 & 0.033 & 0.001 & 0.108\\
\hspace{1em}Local government & 0.062 & 0.031 & 0.011 & 0.112\\
\hspace{1em}International organisations & 0.105 & 0.032 & 0.053 & 0.157\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(results, }\DataTypeTok{ci=}\FloatTok{0.9}\NormalTok{, }\DataTypeTok{point.size =} \FloatTok{.8}\NormalTok{, }\DataTypeTok{dodge.size =} \DecValTok{1}\NormalTok{,}
     \DataTypeTok{text.size =} \DecValTok{10}\NormalTok{, }\DataTypeTok{main =} \StringTok{'AMCE -- Academia Only'}\NormalTok{,}
     \DataTypeTok{attribute.names =}\NormalTok{ attrs)}
\end{Highlighting}
\end{Shaded}

\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/unnamed-chunk-45-1.pdf}

\newpage

\hypertarget{apsa-experimental-section-standard-report-for-experiments}{%
\section{APSA Experimental Section Standard Report for
Experiments}\label{apsa-experimental-section-standard-report-for-experiments}}

\hypertarget{hypothesis}{%
\subsection{Hypothesis}\label{hypothesis}}

The experiment was designed to study the characteristics of the climate
change mitigation treaties favored by Latin American elites.

\hypertarget{subjects-and-context}{%
\subsection{Subjects and Context}\label{subjects-and-context}}

The eligibility criteria for the research was to belong to one of the
four elite profiles listed below.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Members of the Executive branch}: members of regulatory
  agencies, politicians (mayors, governors, presidents), members of
  ministries and secretaries at the federal and state levels.
\item
  \textbf{Members of the Legislative branch}: legislators and staff
  officers from the federal and state level in the researched countries.
\item
  \textbf{Civil society}: environmental and energy-related NGO members;
  oil, renewables, and environmental solutions firm owners in the
  researched countries.
\item
  \textbf{Academics}: professors from energy and engineering departments
  of the most prestigious universities in the countries researched.
\end{enumerate}

We decided to include these profiles based on the stakes and influence
that such elites have on climate change policies. According to the IR
constructivist theories, academics have notable influence on our topic
of interest as they form the epistemic community around climate change
policies. Civil society members that have stakes on climate policies are
also influential, such as oil-extracting companies, lobby groups, and
environmental NGOs. Members of the Executive and Legislative branches
handle the decisions themselves, so their importance.

We selected ten countries to our sample: Argentina, Bolivia, Brazil,
Chile, Colombia, Costa Rica, Ecuador, Mexico, Panama, and Peru. Our
selection criteria were the impact of climate change on each nation,
their size and regional importance, and natural resources availability.
For each country, we built a dataset of potential respondents with at
least:

\begin{itemize}
\tightlist
\item
  150 potential legislative respondents
\item
  150 potential executive respondents
\item
  200 potential civil society respondents
\item
  200 potential academic respondents
\end{itemize}

These numbers represent ten times the number of respondents we were
aiming to interview by the end of the survey. The dataset was build from
October 5th to November 12th, 2018. We hired a team of enumerators that
searched the internet for people in the countries that fulfilled our
pre-determined elite profiles.

We ranked the possible respondents in two tiers. In the first tier we
placed people that we had complete information about their profiles:
office telephone numbers, emails, social media contacts, and so on. In
the second tier we included participants that had incomplete profiles,
such as those with only online information available. The dataset
including the first-tier respondents was randomly divided between
enumerators in Rio de Janeiro and São Paulo to avoid any eventual biases
from the collecting stage. Both groups conducted the interviews by phone
from November 12th to December 5th, 2018. The second tier dataset was
used for an online only version of the survey.

The response rate in each of the enumeration sites and in the online
only survey follows in the table below.

\vspace{.5cm}

\begin{longtable}[]{@{}lllrrr@{}}
\caption{Survey response rates.}\tabularnewline
\toprule
Country & Enumeration Place & Type & Total & Phone &
Online\tabularnewline
\midrule
\endfirsthead
\toprule
Country & Enumeration Place & Type & Total & Phone &
Online\tabularnewline
\midrule
\endhead
Argentina & Rio de Janeiro & Online and Phone & 300 & 33 &
17\tabularnewline
Bolivia & Rio de Janeiro & Online and Phone & 372 & 31 &
11\tabularnewline
Brazil & Rio de Janeiro & Online and Phone & 346 & 34 &
12\tabularnewline
Chile & Rio de Janeiro & Online and Phone & 417 & 35 & 12\tabularnewline
Colombia & Rio de Janeiro & Online and Phone & 445 & 35 &
21\tabularnewline
Costa Rica & Rio de Janeiro & Online and Phone & 409 & 32 &
11\tabularnewline
Ecuador & Rio de Janeiro & Online and Phone & 512 & 30 &
12\tabularnewline
Mexico & Rio de Janeiro & Online and Phone & 468 & 41 &
17\tabularnewline
Panama & Rio de Janeiro & Online and Phone & 291 & 30 &
14\tabularnewline
Peru & Rio de Janeiro & Online and Phone & 243 & 32 & 11\tabularnewline
\emph{TOTAL} & & Online and Phone & 3803 & 333 & 138\tabularnewline
& & & & &\tabularnewline
Argentina & Sao Paulo & Online and Phone & 292 & 48 & 16\tabularnewline
Bolivia & Sao Paulo & Online and Phone & 373 & 50 & 13\tabularnewline
Brazil & Sao Paulo & Online and Phone & 344 & 43 & 15\tabularnewline
Chile & Sao Paulo & Online and Phone & 416 & 41 & 15\tabularnewline
Colombia & Sao Paulo & Online and Phone & 448 & 42 & 19\tabularnewline
Costa Rica & Sao Paulo & Online and Phone & 412 & 44 & 20\tabularnewline
Ecuador & Sao Paulo & Online and Phone & 507 & 43 & 18\tabularnewline
Mexico & Sao Paulo & Online and Phone & 464 & 41 & 17\tabularnewline
Panama & Sao Paulo & Online and Phone & 291 & 50 & 19\tabularnewline
Peru & Sao Paulo & Online and Phone & 248 & 44 & 9\tabularnewline
\emph{TOTAL} & & Online and Phone & 3803 & 446 & 161\tabularnewline
& & & & &\tabularnewline
Argentina & & Online Only & 517 & & 19\tabularnewline
Bolivia & & Online Only & 132 & & 12\tabularnewline
Brazil & & Online Only & 1183 & & 52\tabularnewline
Chile & & Online Only & 470 & & 31\tabularnewline
Colombia & & Online Only & 522 & & 29\tabularnewline
Costa Rica & & Online Only & 325 & & 20\tabularnewline
Ecuador & & Online Only & 460 & & 37\tabularnewline
Mexico & & Online Only & 955 & & 29\tabularnewline
Panama & & Online Only & 319 & & 26\tabularnewline
Peru & & Online Only & 540 & & 25\tabularnewline
\emph{TOTAL} & & Online Only & 5569 & & 293\tabularnewline
\bottomrule
\end{longtable}

The response rate was 10.24 percent of the original population in the
telephone survey (779 of 7606 possible respondents). We had an attrition
of 61.62 percent from the telephone (779) to the online survey (299).
The response rate for the online only survey was 5.26 percent (293 of
5569 invited by email).

The conjoint experiment analyzed in this paper was in the online and
online-only dataset. Therefore, the relevant column for our data is the
last one.

\hypertarget{allocation-methods}{%
\subsection{Allocation Methods}\label{allocation-methods}}

\textbf{Random Assignment:} We programmed the experiment using a Python
application provided by \citet{strezhnev2013conjoint}. We translated the
original survey questions (as shown in the text) into Portuguese and
Spanish, and then embedded the PHP file with the randomization
parameters in a Qualtrics survey. The PHP code we used to randomize the
values of the attributes is available at the project's GitHub
repository: \url{http://github.com/danilofreire/climate-governance}.

For each attribute, the probability of selecting a given component
follows a uniform distribution function. The table below describes the
results.

\vspace{.5cm}

\begin{longtable}[]{@{}llr@{}}
\caption{Conjoint experiment attributes and their respective
probabilities.}\tabularnewline
\toprule
\textbf{Attribute} & \textbf{Values} &
\textbf{Probabilities}\tabularnewline
\midrule
\endfirsthead
\toprule
\textbf{Attribute} & \textbf{Values} &
\textbf{Probabilities}\tabularnewline
\midrule
\endhead
\textbf{Who makes the rules?} & International organizations &
1/5\tabularnewline
& Federal government & 1/5\tabularnewline
& Local government & 1/5\tabularnewline
& Local community members & 1/5\tabularnewline
& Non-governmental organizations & 1/5\tabularnewline
& &\tabularnewline
\textbf{Conflict resolution mechanism} & United Nations &
1/5\tabularnewline
& Government bureaucracy & 1/5\tabularnewline
& Local courts & 1/5\tabularnewline
& Private arbitration & 1/5\tabularnewline
& Informal norms & 1/5\tabularnewline
& &\tabularnewline
\textbf{Punishment} & Imprisonment & 1/4\tabularnewline
& Fines & 1/4\tabularnewline
& Blacklist & 1/4\tabularnewline
& None & 1/4\tabularnewline
& &\tabularnewline
\textbf{Punishment for repeated violations} & More penalty &
1/3\tabularnewline
& Same & 1/3\tabularnewline
& Less penalty & 1/3\tabularnewline
& &\tabularnewline
\textbf{Agreement costs} & Rich countries pay more than poor countries &
1/4\tabularnewline
& Proportional to history of emissions & 1/4\tabularnewline
& Proportional to current emissions & 1/4\tabularnewline
& Only rich countries pay & 1/4\tabularnewline
& &\tabularnewline
\textbf{Renegotiation} & Never & 1/5\tabularnewline
& Fifty years & 1/5\tabularnewline
& Twenty years & 1/5\tabularnewline
& Five years & 1/5\tabularnewline
& One year & 1/5\tabularnewline
\bottomrule
\end{longtable}

We added one logical restriction to the set climate change treaties:
\textbf{Punishment = None} can never appear together with
\textbf{Punishment for repeated violations = Less penalty}. The number
of possible treaties are the product of the attributes
(\(5\times 5 \times 4 \times 3 \times 4 \times 5 = 6,000\)) minus the
number of removed possibilities (\(5\times 5 \times 4 \times 5 = 500\)).
The total number of possible treaties is \(5,500\).

\hypertarget{treatments}{%
\subsection{Treatments}\label{treatments}}

\textbf{Descriptions of the intervention:} Our treatment was the
randomization of components in the conjoined climate mitigation
treaties. We provided the following description before the conjoint
experiment starts, in Spanish and Portuguese:

\vspace{.5cm}

\begin{verbatim}
For the following questions, read carefully the instructions below.

Imagine that your country would sign an international treaty to mitigate climate change. 
A climate mitigation treaty has, in general, the following attributes:

1. Who defines the rules? Which group will define the parameters for the treaty?
2. How are the conflicts that might happen when the treaty begin be resolved?
3. What punishments should be applied to ensure compliance?
4. How will repeated violations be punished?
5. How are costs for implementing the treaty be distributed?
6. How often the treaty has to be renegotiated?

In the following questions, you will see variations in treaties on these six attributes.
Please select the treaty that has the characteristics that you believe are best for your country.
\end{verbatim}

\vspace{.5cm}

The conjoint experiment consisted in comparing two treaties that vary in
these six attributes, repeating the application for seven times for each
respondent.

\noindent \textbf{Software:} To program the PHP randomization device we
used the Python application provided by \citet{strezhnev2013conjoint}.
We used Qualtrics to administer the survey.

\noindent \textbf{Delivery:} We repeated the conjoint experiment seven
times, to improve test power.

\noindent \textbf{Dataset description:} Each line in the dataset
represents a given treaty, varying for the selection status (selected
versus non-selected) for a given respondent. When a respondent finished
an entire response set, this adds 14 rows to the dataset. Seven rows
with the selected treaties and other seven rows with the non-selected
treaties.

\noindent \textbf{Deception:} We used no deception in this survey
experiment.

\hypertarget{results}{%
\subsection{Results}\label{results}}

\noindent \textbf{Outcome measures and covariates:} The main outcome is
a binary indicator for the selection of a given conjoined package. We
added two covariates to measure heterogeneous effects: \emph{Country}
and \emph{Elite type}.

\noindent \textbf{Questionnaire:} For the English translation of the
entire survey instrument, with both the telephone and the online
components, see the attachment below.

\noindent \textbf{Statistical analysis:} We fit a marginal means
estimator and an AMCE estimator. The estimation conditions the results
within subjects and the presented conjoined packages. The results are
available in sections \ref{sec:features} and \ref{sec:amce}.

\noindent \textbf{Missing data and attrition:} Conditional on having
started the conjoint experiment, for each task we had the following
completion rates:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{barplot}\NormalTok{(}\DecValTok{100}\OperatorTok{*}\KeywordTok{table}\NormalTok{(cj}\OperatorTok{$}\NormalTok{task)}\OperatorTok{/}\KeywordTok{max}\NormalTok{(}\KeywordTok{table}\NormalTok{(cj}\OperatorTok{$}\NormalTok{task)), }\DataTypeTok{xlab =} \StringTok{'Conjoint task'}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{'Percentage completed'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{appendix-institutional-design-climate-policy_files/figure-latex/unnamed-chunk-46-1.pdf}

\hypertarget{other-information}{%
\subsection{Other information}\label{other-information}}

\noindent \textbf{IRB:} This research received IRB approval from Brown
University (Protocol 2195/2018) and Fundação Getulio Vargas (Protocol
83/2018).

\noindent \textbf{Pre-registration:} The result was not pre-registered.

\noindent \textbf{Funding:} The research was funded by Konrad Adenauer
Stiftung (KAS). KAS provided EUR\$ 55,000.00 for this research. They
have not interfered in the research design and in the question choices.

\noindent \textbf{Replication materials:} Available at
\url{http://github.com/danilofreire/climate-governance}.

% More bibliography
\renewcommand\refname{Bibliography}
\bibliography{references.bib}

\end{document}